### 2021-03-14 oscar_figures.Rmd
Riboseq volcano plots

---
title: "oscar_figures"
author: "Oscar Wilkins"
date: "14/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DESeq2)
library(here)
library(ggrepel)
```

## Ribosome profiling quality control

Perform quality control to check that reads are periodic and enriched across the llllllllllllllll

## Ribosome profiling

Use DESeq to detect differential ribosome footprints and create volcano plots

```{r message = FALSE}
gene_names <- read_csv(paste0(here(), "/data/gencode_v29_names.csv"))
counts_df <- read_csv(paste0(here(), "/data/riboseq/cds_feature_counts.csv"))
counts <- as.matrix(counts_df %>% column_to_rownames("gene_id"))
ipsc_splicing <- read_csv(paste0(here(), "/data/ipsc_splicing_results.csv"))
ipsc_de = read_csv(paste0(here(), "/data/ipsc_differential_expression.csv"))

col_data <- data.frame(sample = colnames(counts)) %>%
  mutate(condition = ifelse(str_detect(sample, "tdp"), "tdp_ko", "control"))

dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = col_data,
                              design= ~condition)
dds <- DESeq(dds)
results_name <- resultsNames(dds)[2]
res <- results(dds, name=results_name)

res_df <- data.frame(res) %>%
  rownames_to_column("gene_id") %>%
  left_join(gene_names, by = "gene_id")

cryptic_df <- ipsc_splicing %>%
  dplyr::select(gene_name, cryptic_junction) %>%
  unique() %>%
  filter(cryptic_junction)

res_df2 <- res_df %>%
  mutate(cryptic_junction = gene_name %in% cryptic_df$gene_name) %>%
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2","TARDBP") ~ gene_name)) %>%
  mutate(colour = ifelse(cryptic_junction, "#fe6101",  "#648FFF")) %>%
  arrange(colour)
```

```{r} 
p1 <- ggplot(res_df2, aes(x = log2FoldChange, y = -log10(pvalue), label = label_junction)) +
  geom_point(data = res_df2 %>% filter(!cryptic_junction),
             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#648FFF") +
  geom_point(data = res_df2 %>% filter(cryptic_junction),
             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#fe6101") +
  ggpubr::theme_pubr() +
  ylab("-Log10(p-value)") +
  xlab("Log2 fold change") +
  geom_text_repel(box.padding = 1.2, max.overlaps = Inf)

p2 <- p1 + ylim(0,10)
p1
p2
```