### 2021-03-14 oscar_figures.Rmd
Add iCLIP peak plots

---
title: "oscar_figures"
author: "Oscar Wilkins"
date: "14/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DESeq2)
library(here)
library(ggrepel)
```

## Ribosome profiling quality control

Perform quality control to check that reads are periodic.

```{r message=FALSE, warning=FALSE}
read_length_filter <- 29  # length of reads of interest
gencode_v29_info <- read_tsv(paste0(here(), "/data/longest_proteincoding_transcript_hs_details.txt"))
ribo_df <- read_tsv(paste0(here(), "/data/riboseq/sh_tdpb.Aligned.toTranscriptome.out.bam.bed.gz"), 
                    col_names = c("transcript_id", "start0", "end", "strand")) %>%
    inner_join(gencode_v29_info, by = "transcript_id") %>%
    dplyr::filter(end-start0 == read_length_filter, strand == "+") %>%
    mutate(distance_from_start = start0-cds_start) %>%
    group_by(distance_from_start) %>%
    mutate(n_this_dist = n()) %>%
    dplyr::select(distance_from_start, n_this_dist) %>%
    unique()

```

```{r, warning=FALSE}
ggplot(ribo_df, aes(x=distance_from_start, y=n_this_dist)) +
    geom_bar(stat="identity") +
    xlim(-50, 50) +
    ylab("Counts") + 
    xlab("Distance of 5' end from annotated start codon") +
    ggtitle(paste0("Periodicity of ", read_length_filter, "nt reads from TDP-43 KD replicate B"))
```

## Ribosome profiling volcano plot

Use DESeq to detect differential ribosome footprints and create volcano plots

```{r message = FALSE, warning=FALSE}
counts_df <- read_csv(paste0(here(), "/data/riboseq/cds_feature_counts.csv"))
counts <- as.matrix(counts_df %>% column_to_rownames("gene_id"))
ipsc_splicing <- read_csv(paste0(here(), "/data/ipsc_splicing_results.csv"))
ipsc_de = read_csv(paste0(here(), "/data/ipsc_differential_expression.csv"))

col_data <- data.frame(sample = colnames(counts)) %>%
  mutate(condition = ifelse(str_detect(sample, "tdp"), "tdp_ko", "control"))

dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = col_data,
                              design= ~condition)
dds <- DESeq(dds)
results_name <- resultsNames(dds)[2]
res <- results(dds, name=results_name)

res_df <- data.frame(res) %>%
  rownames_to_column("gene_id") %>%
  left_join(gencode_v29_info, by = "gene_id")

cryptic_df <- ipsc_splicing %>%
  dplyr::select(gene_name, cryptic_junction) %>%
  unique() %>%
  filter(cryptic_junction)

res_df2 <- res_df %>%
  mutate(cryptic_junction = gene_name %in% cryptic_df$gene_name) %>%
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2","TARDBP") ~ gene_name)) %>%
  mutate(colour = ifelse(cryptic_junction, "#fe6101",  "#648FFF")) %>%
  arrange(colour)
```

```{r, warning=FALSE} 
p1 <- ggplot(res_df2, aes(x = log2FoldChange, y = -log10(pvalue), label = label_junction)) +
  geom_point(data = res_df2 %>% filter(!cryptic_junction),
             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#648FFF") +
  geom_point(data = res_df2 %>% filter(cryptic_junction),
             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#fe6101") +
  ggpubr::theme_pubr() +
  ylab("-Log10(p-value)") +
  xlab("Log2 fold change") +
  geom_text_repel(box.padding = 1.2, max.overlaps = Inf)

p2 <- p1 + ylim(0,10)
p1
p2
```

## Targeted RNA-seq

Read in data from processed bam files

TODO: add scripts which deduplicate UMIs for each.

```{r message=FALSE, warning=FALSE}
specific_rt <- read_csv(paste0(here(), "/data/targeted_RNA_seq/specific_RT_primer/for_plot.csv")) %>%
    select(risk, sample_name, n)

random_hexamer <- read_csv(paste0(here(), "/data/targeted_RNA_seq/random_hexamer/values_for_plot.csv")) %>%
    select(-sample_name) %>%
    select(sample_name = actual_sample_name, risk, n)

tg_rnaseq_df <- bind_rows(specific_rt, random_hexamer) %>%
  group_by(sample_name) %>%
  mutate(n_healthy = ifelse(risk == "Non-Risk", n, 0)) %>%
  mutate(combined_healthy = sum(n_healthy)) %>%
  select(-n_healthy) %>%
  mutate(combined = sum(n)) %>%
  group_by(sample_name, risk) %>%
  mutate(combined_n = sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  unique() %>%
  group_by(sample_name) %>%
  mutate(p_value = pbinom(combined_healthy, sum(combined_n), 0.5)) %>%  # single-tailed bionomial test
  mutate(colour = ifelse(risk == "Non-Risk", "lightblue", "red")) %>%
  ungroup() %>%
  mutate(sample_name_ordered = fct_reorder(sample_name, dplyr::desc(combined_n)))

ggplot(tg_rnaseq_df %>%
         filter(combined_n > 0), aes(x = sample_name_ordered, y = combined_n, fill = risk)) +
  geom_bar(stat="identity", position="dodge") +
  geom_text(aes(x=sample_name, y=-2.5, label=signif(p_value,2)), size = 3.5) +
  xlab("") +
  ylab("Unique cDNAs") +
  ggeasy::easy_add_legend_title("Allele") +
  theme_minimal() +
  ggpubr::theme_pubr() +
  ggeasy::easy_rotate_x_labels()
```

## iCLIP of minigenes

Read in the processed data (following alignment with Bowtie2, and filtering for high-confidence alignments), and generate
a plot of the overall distribution of crosslinks, and the rolling average difference between the 2x Risk and 2x Healthy

TODO: add script which does this filtering

```{r message=FALSE, warning=FALSE}
library(patchwork)

exon_snp <- 556 
intron_snp <- 1106
intron_start <- 141 
intron_end <- 1429
acceptor1 <- 394
acceptor2 <- 444
donor <- 572
mg_l <- 1624 
cpoor_start <- 490
cpoor_end <- 1325
risk_colour <- "#00BFC4"
healthy_colour <- "#F8766D"

filtered_df <- read_csv(paste0(here(), "/data/iCLIP/bowtie2_aligned_2_filtered_df.csv"))

rolling_window <- 20

df2 <- filtered_df %>%
  ungroup() %>%
  group_by(condition, pos) %>%
  mutate(mean_counts = sum(normalised_counts)/2) %>%
  select(pos, condition, mean_counts) %>%
  unique() %>%
  pivot_wider(names_from = "condition", values_from ="mean_counts")

df2[is.na(df2)] <- 0

df2 <- df2 %>% 
  mutate(diff = R-H) %>%
  mutate(mean_4 = (H+R)/2) %>%
  arrange(pos)

top <- ggplot(df2, aes(x=pos, y=diff)) +
  geom_area(aes(x=pos, y = mean_4), colour="grey50", fill = "grey50", alpha=0.2) +
  geom_vline(xintercept = exon_snp, linetype="dashed") +
  geom_vline(xintercept = intron_snp, linetype="dashed") +
  xlim(1,1600) +
  #ylim(-1,5) +
  theme_minimal() +
  ylab("XLs per 1000") +
  xlab("") +
  geom_vline(xintercept = cpoor_start) +
  geom_vline(xintercept = cpoor_end)

bottom <- ggplot(df2, aes(x=pos, y=zoo::rollmean(diff, k = rolling_window, na.pad=T))) +
  geom_area(colour = "grey50", fill="grey50", alpha = 1) +
  geom_vline(xintercept =  exon_snp, linetype="dashed") +
  geom_vline(xintercept =  intron_snp, linetype="dashed") +
  xlim(1,1600) +
  ylim(-0.5,0.75) +
  theme_minimal() +
  ylab("Change in XL density") +
  xlab("Distance from start of minigene") +
  geom_vline(xintercept =  intron_start) +
  geom_vline(xintercept =  intron_end) +
  geom_vline(xintercept =  acceptor1) +
  geom_vline(xintercept =  acceptor2) +
  geom_vline(xintercept =  donor)

combined <- top/bottom
combined
```

Next, filter for peaks in the iCLIP signal, and see how they change in 2x Risk versus 2x Healthy. Highlight those which are close to the SNPs.

```{r message=FALSE, warning=FALSE}

find_ratio_bound <- function(e_x, sem_x, e_y, sem_y, quantile_value, randoms = 10000){
  # monte carlo confidence interval estimate for a ratio. Assumes the two variables
  # x and y are normally distributed. Finds confidence interval for (x-y)/y

  # randomly generate possible global means for x and y

  xs = rnorm(randoms, mean = e_x, sd=sem_x)  # risk
  ys = rnorm(randoms, mean = e_y, sd=sem_y)  # healthy

  ratios = (xs-ys)/ys

  return(quantile(ratios, quantile_value))
}

## Plot big changes

confidence_interval <- 0.75  # fracion, i.e. 0.75 = 75%.

peak_threshold <- 5 # how many times above the mean to be a "peak"
region_length <- 50

peak_t <- peak_threshold*1000/mg_l

peak_df <- filtered_df %>%
  ungroup() %>%
  group_by(pos) %>%
  mutate(max_at_pos = max(normalised_counts)) %>%
  ungroup() %>%
  filter(max_at_pos >= peak_t) %>%
  group_by(condition, pos) %>%
  mutate(peak_mean = mean(normalised_counts), 
         peak_sd = sd(normalised_counts)) %>%
  ungroup() %>%
  select(condition, pos, peak_mean, peak_sd) %>%
  unique() %>%
  pivot_wider(names_from = "condition", values_from = c("peak_mean", "peak_sd")) %>%
  mutate(frac_increase = (peak_mean_R-peak_mean_H)/peak_mean_H) %>%
  mutate(peak_rank = rank(frac_increase)) %>%
  mutate(colour = ifelse(abs(pos-exon_snp)<region_length, "gold", 
                         ifelse(abs(pos-intron_snp)<region_length, "purple", "grey50")))

peak_df$upper_bound <- mapply(find_ratio_bound, 
                        peak_df$peak_mean_R,
                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
                        peak_df$peak_mean_H,
                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
                        0.5*(1+confidence_interval))

peak_df$lower_bound <- mapply(find_ratio_bound, 
                        peak_df$peak_mean_R,
                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
                        peak_df$peak_mean_H,
                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
                        0.5*(1-confidence_interval))

ggplot(peak_df, aes(x = peak_rank, y = 100*frac_increase, fill=colour)) +
  geom_bar(stat="identity") +
  scale_fill_identity() +
  theme_minimal() +
  ggpubr::theme_pubr()+
  ylab("% change in 2x Risk") +
  xlab("Rank") +
  geom_errorbar(aes(ymin = 100*lower_bound, ymax = 100*upper_bound, x = peak_rank), 
                width = 0.3)

```

##