### 2021-03-19 paper_figures.Rmd
Update paper_figures.Rmd
fix git merge flags

---
title: "Figures for Paper"
author: "Anna-Leigh Brown + Matteo Zanovello"
date: "06/10/2020"
output: 
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
---

```{css, echo = FALSE}
h1, #TOC>ul>li {
  color: #000000;
    font-weight: bold;
}

h2, #TOC>ul>ul>li {
  color: #41494D;
}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
knitr::opts_chunk$set(
  comment = '', fig.width = 9, fig.height = 9
)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load("data.table",
       "tidyverse",
       "ggplot2",
       "rcompanion",
       "lme4",
       "ggpubr",
       "forcats",
       "scales",
       "here",
       "ggthemes",
       "ggrepel")
```

# iPSC splicing

Read in the splicing results returned by MAJIQ and make a volcano plot, only highlight 
genes of interest with a label.

```{r volcano_plots}

ipsc_splicing = fread(file.path(here::here(),"data","ipsc_splicing_results.csv"))
ipsc_de = fread(file.path(here::here(),"data","ipsc_differential_expression.csv"))

splicing_dots_tables <- ipsc_splicing %>% 
  mutate(junction_name = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ gene_name,
                                    T ~ "")) %>% 
  mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
  mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>% 
  mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>% 
  mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ junction_name,
                                    T ~ ""))

fig1a = ggplot() +
  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions != 0),
             aes(x = deltaPSI, y =log10_test_stat,
                 alpha = graph_alpha,,fill = "Annotated Junction"), pch = 21, size = 10) + 
  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions == 0),
             aes(x = deltaPSI, y =log10_test_stat,
                 alpha = graph_alpha,fill = "Novel Junction"), pch = 21, size = 10) + 
  geom_text_repel(data = splicing_dots_tables[junction_name != ""],
                  aes(x = deltaPSI, y =log10_test_stat,
                      label = label_junction,
                      color = as.character(de_novo_junctions)), point.padding = 0.3,
                    nudge_y = 0.2,
                  min.segment.length = 0,
                  box.padding  = 2,
                  size=6,show.legend = F) +
  geom_hline(yintercept = -log10(1 - .9)) + 
  geom_vline(xintercept = -0,linetype="dotted") + 
     scale_fill_manual(name = "",
                      breaks = c("Annotated Junction", "Novel Junction"),
                      values = c("Annotated Junction" = "#648FFF", "Novel Junction" = "#fe6101") ) +
       scale_color_manual(name = "",
                      breaks = c("0", "1"),
                      values = c("1" = "#648FFF", "0" = "#fe6101") ) +
  guides(alpha = FALSE, size = FALSE) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab(expression(paste("-Lo", g[10], " Test Statistic"))) +
  theme(text = element_text(size = 24)) +
  theme(legend.text=element_text(size=22)) +
  xlim(-1,1) +
  scale_x_continuous(labels = scales::percent)

de_table = ipsc_de %>% 
  mutate(contains_cryptic = gene_name %in% splicing_dots_tables[cryptic_junction == T,unique(gene_name)]) %>% 
  mutate(contains_cryptic = as.character(as.numeric(contains_cryptic))) %>% 
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2","TARDBP") ~ gene_name,
                                    T ~ "")) %>% 
  mutate(graph_alpha = ifelse(padj < 0.1, 1, 0.2)) %>% 
  mutate(y_data = -log10(padj))

fig1b = ggplot(data = de_table) + 
  geom_point(data = de_table %>% filter(contains_cryptic == "0"),
         aes(x = log2FoldChange, y = -log10(padj),
             alpha = graph_alpha,fill = "No Cryptic"), pch = 21, size = 10) + 
    geom_point(data = de_table %>% filter(contains_cryptic == "1"),
         aes(x = log2FoldChange, y = -log10(padj),
             alpha = graph_alpha,fill = "Contains Cryptic"), pch = 21, size = 10)+ 
    geom_text_repel(data = de_table[label_junction != ""],max.overlaps = 500,
                  aes(x = log2FoldChange, 
                      y = -log10(padj),
                      label = label_junction,
                      color = as.character(contains_cryptic)),
                   nudge_y = 5,
                  min.segment.length = 0,
                  box.padding  = 4,
                  size=6,show.legend = F) +
       scale_fill_manual(name = "",
                      breaks = c("No Cryptic", "Contains Cryptic"),
                      values = c("No Cryptic" = "#648FFF", "Contains Cryptic" = "#fe6101") ) +
         scale_color_manual(name = "",
                      breaks = c("1", "0"),
                      values = c("1" = "#fe6101", "0" = "#648FFF") ) +
  xlim(-7.5,7.5) + 
  ylab(expression(paste("-Lo", g[10], " P-value"))) +
    guides(alpha = FALSE, size = FALSE) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab(expression(paste("Lo", g[2], " Fold Change"))) +
  theme(text = element_text(size = 24)) +
  theme(legend.text=element_text(size=22)) +
    geom_hline(yintercept = -log10(0.1))  +
  geom_vline(xintercept=c(0), linetype="dotted")

print(fig1a)
print(fig1b)
```

# UNC13A CE PSI across TDP-43 KD experiments

The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192. 
The SK-N-DZ cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele. 
There was variability on the Klim hMN set on allelic expression.

```{r read in }

rel_rna_cryptic_amount = data.table::fread(file.path(here::here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
```

## barplot UNC13A CE PSI - full five

```{r}
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A CE PSI") + 
    xlab("") + 
    guides(color = FALSE) +
    scale_y_continuous(labels = scales::percent) +
    theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))
```

## barplot UNC13B frameshift PSI

```{r}
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "condition",
              add = c("mean_se","jitter"),
              y = "unc13b_nmd_exon_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8),facet.by = 'source') + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B \nNMD Exon PSI") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    facet_wrap(~source) + 
  stat_compare_means() +
  scale_y_continuous(labels = scales::percent)
```

## TARDBP RNA and UNC13A Cryptic

```{r}

rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12,method = 's',cor.coef.name = 'rho') +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
    scale_y_continuous(labels = scales::percent)
```

## UNC13A RNA and UNC13A Cryptic

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12,
             method = 'spearman',
           cor.coef.name = 'rho') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
      scale_y_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) + 
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
      theme(text = element_text(size = 18,family = 'sans'), 
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
```
## UNC13A RNA and UNC13A IR
```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12,cor.coef.name = 'rho',method = 's') +
    geom_smooth(aes(x = UNC13A, y = normalized_unc13a_ir),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
```
## TARDBP RNA and UNC13A IR

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13a_ir),size = 12) +
    geom_smooth(aes(x = TARDBP, y = normalized_unc13a_ir),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
```

## barplot UNC13A cryptic PSI

```{r}
rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "cryptic_psi_full",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("UNC13A CE PSI") + 
  xlab("") + 
  guides(color = FALSE) 
```

## barplot STMN2 cryptic PSI

```{r}
stmn2 = rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "stmn_2_cryptic_psi",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("STMN2 Cryptic PSI") + 
  xlab("") + 
  guides(color = FALSE)
print(stmn2)
```

## barplot UNC13B normalized IR
```{r}
unc13b_ir = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13b_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13B \nIntron Retention Ratio") + 
    xlab("") + 
    guides(color = FALSE) + 
      theme(legend.position = "none") +
  theme(text = element_text(size = 18)) +
      guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

print(unc13b_ir)
```

## barplot UNC13A normalized IR
```{r}
unc13a_ir = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13a_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13A \nIntron Retention Ratio") + 
    xlab("") + 
    theme(legend.position = "none") +
    guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

```

## barplot UNC13A normalized RNA Level

```{r}
text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "UNC13A",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A") + 
    xlab("") + 
    theme(legend.position = "none") +
    guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent) +
    geom_text(data = text_table[gene_name == "UNC13A"],aes(x = source, y= 1.35,label = padj_plot),size = 8) + 
    geom_text(data = text_table[gene_name == "UNC13A"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

```

## barplot UNC13B normalized RNA Level

```{r}
text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "UNC13A",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B") + 
    xlab("") + 
    theme(legend.position = "none") +
    guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent) +
    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.35,label = padj_plot),size = 8) + 
    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

```

## barplot TARDBP normalized RNA Level

```{r}
text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "TARDBP",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("TARDBP") + 
    xlab("") + 
    theme(legend.position = "none") +
    guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent,expand = expansion(mult = c(0, .1))) +
    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.35,label = padj_plot),size = 8) + 
    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

```
## Scatterplot stmn2 normalized TARDBP
```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = stmn_2_cryptic_psi, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = stmn_2_cryptic_psi)) +
    theme(legend.position = 'bottom')
```

## Scatterplot UNC13BIR normalized TARDBP

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')
```

## Scatterplot UNC13A-IR and UNC13A Crptic
```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, 
                   y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, 
                 y = normalized_unc13a_ir)) +
    theme(legend.position = 'bottom')
```

## Scatterplot UNC13B-IR and UNC13A Cryptic

```{r}

rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)
```

## scatterplot UNC13BIR and UNC13B NMD

```{r}
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)

```

# NMD experiment on SH-SY5Y cells

```{r}

for_bar_plot_sy <- read_csv(file.path(here::here(),"data","results_17022021_sy5y.csv"))

for_bar_plot_sy <- data.frame(for_bar_plot_sy)
for_bar_plot_sy$condition <- as.factor(for_bar_plot_sy$condition)
for_bar_plot_sy$condition <- factor(for_bar_plot_sy$condition,levels(for_bar_plot_sy$condition)[c(2,1)])

nmd_plot = ggbarplot(for_bar_plot_sy,
                        x = 'condition',
                        add = c("mean_se","jitter"),
                        y = 'value',
                        fill = 'condition',
                        color = 'condition',
                        position = position_dodge(0.8),
                        dot.size = 10,facet.by = 'gene') + 
  scale_y_continuous(labels = scales::percent) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
    values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
    values = c("#1C2617","#262114")
  ) +
  ylab("Percent of transcript after CHX treatment") + 
  stat_compare_means()

plot(nmd_plot)
```

```{r}

for_bar_plot_dz <- read_csv(file.path(here::here(),"data","results_17022021_dz.csv"))

for_bar_plot_dz <- data.frame(for_bar_plot_dz)
for_bar_plot_dz$condition <- as.factor(for_bar_plot_dz$condition)
for_bar_plot_dz$condition <- factor(for_bar_plot_dz$condition,levels(for_bar_plot_dz$condition)[c(2,1)])

nmd_plot2 = ggbarplot(for_bar_plot_dz,
                        x = 'gene',
                        add = c("mean_se","jitter"),
                        y = 'value',
                        fill = 'condition',
                        color = 'condition',
                        position = position_dodge(0.8),
                        dot.size = 10) + 
  scale_y_continuous(labels = scales::percent) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
    values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
    values = c("#1C2617","#262114")
  )

plot(nmd_plot2)
```

# iPSC protein abundance mass spec

```{r}
peptides = fread(file.path(here::here(),"data","peptide_abdundance.csv"))

peptides %>% 
  mutate(gene = fct_relevel(gene, "UNC13A","UNC13B")) %>% 
  ggbarplot(,
            x = "condition",
            add = c("mean_se"),
            y = "protein",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8),
            facet.by = 'gene') + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("Protein abundance") + 
  xlab("") + 
  geom_jitter(pch = 21,height = 0,aes(fill = condition),size = 2) + 
  guides(color = FALSE) +
  facet_wrap(~gene, scales = 'free') +
  theme(legend.position = 'none') + 
  scale_y_continuous(labels = scientific) + 
  stat_compare_means(comparisons = list(c("Control","TDP-43 KD")),
                     label = 'p.signif',tip.length = 0,
                     size = 10)

```

# Patient summary statistics

```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(rs12973192 = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(rs12973192) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>% 
    mutate(junction_reads_stmn2 = STMN2_annotated_leaf + STMN2_cryptic_leaf) %>% 
  mutate(junction_reads_unc13a = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf + UNC13A_annotated_leaf) %>% 
  unique()

print(glue::glue("Number of unique patients: {clean_data_table[,length(unique(participant_id))]}"))
print(glue::glue("Number of unique tissue samples: {clean_data_table[,length(unique(sample))]}"))
print("Patients Per Disease Category")
clean_data_table[,length(unique(participant_id)),by = disease]
print("Tissues Per Disease Category")
clean_data_table[,length(unique(sample)),by = disease]
print("Number of patients per UNC13A SNP genotype")
unique(clean_data_table[,.(participant_id,rs12608932,rs12973192)]) %>% select(-participant_id) %>% table()
print("Number of tissues per disease")
clean_data_table[,.N,by = c("disease","tissue_clean")]
print("Number of partcipants by mutation and  disease")
clean_data_table[,length(unique(participant_id)),by = c("mutations","disease")]

print(glue::glue("Number of patients per pathology:"))
clean_data_table[,length(unique(participant_id)),by = .(pathology)]
```
# UNC13A cryptic is an event that is specific to TDP-43 proteinopathy

FTLD-non-TDP are those with TAU and FUS aggregates

Non-tdp ALS are those with SOD1 or FUS mutations. The n's are quite low on this 
unfortunately, only 8 ALS with SOD1 and 2 with FUS mutations.

First we look at detection rate in tissues affected by TDP-43 proteinopathy,
For FTLD this is frontal and temporal Cortices, and
for ALS this is lumbar, cervical, and thoracic spinal cord samples as well as 
motor cortex. For controls we also take all 6 tissues, 
frontal,temporal,motor cortices and the lumbar, cervical, and thoracic spinal cords.

(As a side note, once we do this the number of ALS-non-TDP drops down to 6 (2 FUS) because
the ALS sample tissues are not balanced and not every participant has samples in every tissue)

```{r}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_1",variable)) %>% 
    group_by(sample) %>% 
    mutate(inclusion_reads = sum(value)) %>% 
    ungroup() %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>% 
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```
## Pecent of patients UNC13A detected
Looking at disease tissue only, so just taking the cord and motor cortex in ALS and the 
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.

```{r}

boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected) %>% 
  unique() %>% 
  group_by(disease_group2) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 24)) + 
  xlab("N individuals")
```

## Pecent of samples UNC13A detected by tissue

```{r}
boxplot_table %>% 
  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(sample,disease_group2,detected,tissue_clean) %>% 
  unique() %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  ungroup() %>% 
  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>% 
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Motor Cortex",
                                    "Thoracic Spinal Cord")) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate,fill = detection_name),position = position_dodge2(width = 0.8, preserve = "single")) + 
  ggpubr::theme_pubr() + 
  ylab("Percent of Tissues \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N tissues") + 
  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) + 
  theme(axis.text.x =  element_text(size = 14)) + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  +
  scale_fill_manual(values = c("#4C97B5","#4C97B5","#4C97B5","#4C97B5","#4C97B5","red","#C87156","#C87156")) +
  theme(legend.position = "none") 
```

## Pecent of samples UNC13A detected by tissue and genotype

```{r}

boxplot_table %>% 
  left_join(clean_data_table %>% dplyr::select(sample, rs12973192)) %>% 
  unique() %>% 
  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(sample,disease_group2,detected,tissue_clean,rs12973192) %>% 
  unique() %>% 
  group_by(disease_group2,tissue_clean,rs12973192) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  ungroup() %>% 
  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>% 
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Motor Cortex",
                                    "Thoracic Spinal Cord")) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate,fill = rs12973192),position = position_dodge2(width = 0.8, preserve = "single")) + 
  ggpubr::theme_pubr() + 
  ylab("Percent of Tissues \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N tissues") + 
  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) + 
  theme(axis.text.x =  element_text(size = 14)) + 
      theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  +
  theme(legend.position = "bottom") 
```

## UNC13A CE level across tissue

This only plots the short novel acceptor and novel donor

```{r onlyshort}
boxplot_table %>% 
    filter(tissue_clean %in% 
               c("Frontal Cortex", 
                 "Lumbar Spinal Cord", 
                 "Cervical Spinal Cord",
                 "Motor Cortex", 
                 "Temporal Cortex",
                 "Cerebellum") )%>% 
    group_by(disease_group2,tissue_clean) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    ungroup() %>% 
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>% 
    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                      "Frontal Cortex",
                                      "Lumbar Spinal Cord",
                                      "Temporal Cortex",
                                      "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, 
               y = inclusion_reads * 10^6,
               fill ="red")) + 
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1),outlier.colour = NA)) + 
    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+ 
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) + 
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) + 
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
    theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    theme(text = element_text(size = 18))  +
    theme(legend.position = "bottom") 
```

## UNC13A CE level across tissue - split by junction type

```{r}
clean_data_table %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_1|_5prime_2",variable)) %>% 
      # filter(grepl("_5prime_2",variable)) %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  dplyr::select(sample,disease_group2,
                tissue_clean,
                junction_name,
                value) %>% 
    unique() %>% 
    filter(tissue_clean %in% 
               c("Frontal Cortex", 
                 "Lumbar Spinal Cord", 
                 "Cervical Spinal Cord",
                 "Motor Cortex", 
                 "Temporal Cortex",
                 "Cerebellum"))%>% 
    group_by(disease_group2,tissue_clean) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    ungroup() %>% 
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>% 
    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                      "Frontal Cortex",
                                      "Lumbar Spinal Cord",
                                      "Temporal Cortex",
                                      "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, 
               y = value * 10^6,
               fill = junction_name)) + 
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single")) + 
    geom_point(size = 2,pch = 21, alpha = 0.7, position = position_jitterdodge(jitter.width = 0.2))+ 
    scale_y_log10() + 
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) + 
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) + 
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
        theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  +
  theme(legend.position = "bottom") 
```

## UNC13A CE level across disease types

```{r}

disease_comparisons = list( c("Control","ALS-TDP"),
                           c("Control","ALS \n non-TDP"),
                           c("Control","FTLD-TDP"),
                           c("Control","FTLD \n non-TDP" ))

table_to_test = boxplot_table %>% 
    filter(disease_tissue == T) %>% 
    group_by(disease_group2) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
    dplyr::select(detection_name,
                  participant_id,
                  inclusion_reads,
                  disease_group2,
                  tissue_clean,
                  disease_tissue) %>% 
    unique()

test_pair = pairwise.wilcox.test(table_to_test$inclusion_reads, table_to_test$detection_name,
                     p.adjust.method = "BH") %>% broom::tidy()

test_pair = test_pair %>% 
  mutate(p_value_draw = case_when(p.value < 0.0001~ "***",
                                  p.value < 0.01 ~ "**",
                                   p.value < 0.05 ~ "*",
                          TRUE ~ paste0("Adj. p-value \n",as.character(round(p.value,2))))) %>% 
  mutate(y.position = seq(0.25,by = 0.1,length.out = 7))

table_to_test %>% 
    ggplot(aes(x = detection_name, y = inclusion_reads * 10^6)) + 
    geom_boxplot() + 
    geom_jitter(height = 0) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    ylab("UNC13A cryptic inclusion \n reads per million") + 
    theme(text = element_text(size = 24)) + 
    xlab("N samples") + 
    stat_pvalue_manual(test_pair %>% filter(p.value < 0.05),
                       label = "p_value_draw",size = 8) +
   stat_compare_means(size = 8)

```

# UNC13A RNA expression (TPM)

## UNC13A TPM by disease and tissue

```{r}
comps = clean_data_table %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  ungroup() %>% 
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% pull(detection_name) %>% 
  unique() %>% 
  combn(.,2,simplify = F)

clean_data_table %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  ungroup() %>% 
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
 filter(tissue_clean %in% 
           c("Frontal Cortex", 
             "Lumbar Spinal Cord", 
             "Cervical Spinal Cord",
             "Motor Cortex", 
             "Temporal Cortex",
             "Cerebellum") )%>% 
mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                              "Frontal Cortex",
                              "Lumbar Spinal Cord",
                              "Temporal Cortex",
                              "Motor Cortex")) %>% 
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) + 
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none') + 
  stat_compare_means(comparisons = comps)
```

```{r make_those_plots_function}
plot_by_tissue = function(tissue_name,dt){

only_tissue = dt %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  ungroup() %>% 
  filter(tissue_clean == tissue_name) %>% 
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean))

unique_x_values = only_tissue %>% 
  pull(detection_name) %>% 
  unique()

comparisons = combn(unique_x_values,2,simplify = F)

the_plot = only_tissue %>% 
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) + 
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none') + 
  stat_compare_means() + 
  stat_compare_means(label = "p.signif",comparisons = comparisons,)

  return(the_plot)
}

dis_tis = clean_data_table[disease_tissue == T,unique(tissue_clean)]
dis_tis = c("Cerebellum",dis_tis)
my_plots = purrr::map(dis_tis,plot_by_tissue, clean_data_table)
ggarrange(plotlist=my_plots)

```

## Correlation STMN2 Cryptic PSI and UNC13A TPM

only in ALS/FTLD-TDP

```{r}
clean_data_table %>% 
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A TPM") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  ylim(0,50)
```

## Correlation STMN2 Cryptic PSI and UNC13A TPM - tissue separations

```{r}

clean_data_table %>% 
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A TPM") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = "free_y")
```

## Correlation between UNC13A CE PSI and UNC13A TPM

```{r}
clean_data_table %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(UNC13A_annotated_leaf > 15) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(unc13a_cryptic_leaf_psi > 0) %>% 
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13B_TPM)) + 
  geom_point() + 
  stat_cor(label.y = 20,size = 8,method = 's',cor.coef.name = 'rho') + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 22)) +
  scale_x_continuous(labels = scales::percent)

```

## Correlation between UNC13A CE PSI and UNC13A TPM - tissue separated

```{r}
clean_data_table %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(UNC13A_annotated_leaf > 15) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM, color = disease_group2)) + 
  geom_point() + 
  stat_cor(size = 8) + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = 'free') +
  scale_x_continuous(labels = scales::percent)

```

# Relationship between STMN2 and UNC13A CE PSI 
## Scatter plot showing the correlation in non-log space in cortex UNC13A & STMN2 CE

```{r}
clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(junction_reads_stmn2 >= 30) %>% 
  filter(junction_reads_unc13a >= 30) %>%
  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>% 
  # filter(STMN2_cryptic_leaf >= 2) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(unc13a_cryptic_leaf_psi > 0) %>% 
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) + 
  geom_point() + 
  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") + 
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A CE PSI") + 
  ggpubr::theme_pubr() + 
  theme(text = element_text(size = 18)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))  
```

## Side by side comparison UNC13A & STMN2 CE PSI
```{r}
clean_data_table %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
    filter(tissue_clean %in% 
               c("Frontal Cortex", 
                 "Lumbar Spinal Cord", 
                 "Cervical Spinal Cord",
                 "Motor Cortex", 
                 "Temporal Cortex",
                 "Cerebellum") )%>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(disease_tissue == T) %>% 
  filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 1) %>% 
  filter(STMN2_cryptic_leaf >= 1) %>% 
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  melt() %>% 
    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Temporal Cortex")) %>% 
  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(aes(color = variable),outlier.shape = NA) +
  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
  scale_fill_manual(values = c("#008574","#cd5582")) +
  scale_color_manual(values = c("#008574","#cd5582")) +
  facet_wrap(~tissue_clean,nrow = 3, scales = 'free') +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("CE PSI")  +
  theme(legend.position = 'none')

```

## Side by side comparison UNC13A & STMN2 CE PSI v2 
```{r}
clean_data_table %>% 
  filter(UNC13A_annotated_leaf > 10) %>% 
  filter(STMN2_annotated_leaf > 10) %>% 
  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  # filter(unc13a_cryptic_leaf_psi > 0) %>%
  # filter(disease_tissue == T) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  filter(tissue_clean %in% c("Motor Cortex", "Cervical Spinal Cord", "Frontal Cortex", "Lumbar Spinal Cord", "Temporal Cortex")) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_rat = log2((unc13a_cryptic_leaf_psi + 0.1) / (stmn_2_cryptic_psi_leaf + 0.1))) %>% 
  group_by(tissue_clean, disease_group2) %>% 
  mutate(average_log2Fold = mean(log2_rat)) %>% 
  ungroup() %>% 
  dplyr::select(-log2_rat) %>% 
  pivot_longer(cols = c("stmn_2_cryptic_psi_leaf","unc13a_cryptic_leaf_psi")) %>% 
  mutate(tissue_clean = fct_reorder(tissue_clean,average_log2Fold)) %>% 
  mutate(variable = ifelse(grepl("stmn_2",name), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = tissue_clean, y = value,fill = name)) +
  geom_boxplot() +
  geom_point(pch = 21,height = 0,width = 0.2,position = position_jitterdodge()) +
  scale_fill_manual(values = c("#004D40","#882255")) +
  facet_wrap(~tissue_clean+disease_group2,nrow = 1,scales = 'free')  +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position = 'none')
```

## Side by side comparison UNC13A & STMN2 CE PSI - v3
```{r}
clean_data_table %>% 
  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  # filter(unc13a_cryptic_leaf_psi > 0) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  melt() %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Temporal Cortex")) %>% 
  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(aes(color = variable)) +
  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
  scale_fill_manual(values = c("#004D40","#882255")) +
  scale_color_manual(values = c("#004D40","#882255")) +
  facet_wrap(~tissue_clean+disease_group2,nrow = 3,scales = 'free') +
    scale_y_continuous(labels = scales::percent)  +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position = 'none')

```

## Log2FC UNC13A & STMN2 CE PSI
```{r}
clean_data_table %>% 
  filter(UNC13A_annotated_leaf > 10) %>% 
  filter(STMN2_annotated_leaf > 10) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(unc13a_cryptic_leaf_psi > 0) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf)))%>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%

  # mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>% 
  ggplot(aes(x = tissue_clean,y = log2_unc, fill = tissue_clean)) + 
  geom_boxplot(position = 'dodge') +
  # geom_jitter(height = 0,width = 0.2,pch = 21) +
  facet_wrap(~disease_group2,scales = "free_x") + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
```

# Effect of UNC13A genotype

## UNC13A psi by genotype - rs12973192 - RE

```{r}
rs12973192_comparisons = clean_data_table %>% 
  filter(junction_reads_unc13a >=30) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )"))

comp_for_gg = rs12973192_comparisons %>% pull(detection_name) %>% 
  unique() %>% 
  combn(.,2,simplify = F)

ce_psi = rs12973192_comparisons %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  ylab("UNC13A PSI") +
  xlab(element_blank()) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means() +
  theme(legend.position = "none") +
  stat_compare_means(comparisons = comp_for_gg,label = "p.signif") + 
  scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
  ggtitle("CE SNP (rs12973192)")

```

## UNC13A psi by genotype - rs12608932 - RI
```{r}
rs12608932_comparisons = clean_data_table %>% 
  filter(junction_reads_unc13a >=30) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12608932) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )"))

comp_for_gg = rs12608932_comparisons %>% pull(detection_name) %>% 
  unique() %>% 
  combn(.,2,simplify = F)

in_psi = rs12608932_comparisons %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12608932)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  ylab(element_blank()) +
  xlab(element_blank()) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(legend.position = "none") +
  theme(text = element_text(size = 24)) +
  stat_compare_means() +
  stat_compare_means(comparisons = comp_for_gg,label = "p.signif") + 
    scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
  ggtitle("intronic SNP (rs12973192)")
```

```{r}
ggpubr::ggarrange(ce_psi,in_psi)
```

## UNC13A psi by genotype - rs12973192
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## UNC13A psi by genotype - tissue split
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## UNC13A psi in detected tissues by genotype - tissue split
```{r}

clean_data_table %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi by genotype 
```{r}

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi by genotype - tissue split
```{r}

clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi in detected tissues by genotype - tissue split
```{r}

clean_data_table %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% 
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) + 
  geom_boxplot() + 
  geom_jitter( height = 0) + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## Correlation UNC13A PSI and STMN2 PSI with genotype effect - rs12973192

```{r}

ce_snp = clean_data_table %>% 
  filter(disease_tissue == T) %>% 
  filter(junction_reads_stmn2 >= 30) %>% 
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(rs12973192 = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>%
  filter(!grepl("Cord",tissue_clean)) %>% 
  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn2",
                      y = "log10_unc13a",
                      color = 'rs12973192',                      
                      add = "reg.line", size = 3
                      ) +
    ylab("UNC13A CE PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                      y = unc13a_cryptic_leaf_psi,
                      color = rs12973192),
           method = 'spearman',
           cor.coef.name = 'rho',
           show.legend = FALSE,
           size = 14) + 
    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  theme(legend.title = element_blank()) +
  theme(text = element_text(size = 24),legend.text = element_text(size = 32)) +
    ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI"))) +
  ggtitle("CE SNP (rs12973192) genotype")

```

## Correlation UNC13A PSI and STMN2 PSI with genotype effect - rs12608932

```{r}

in_snp = clean_data_table %>% 
  filter(disease_tissue == T) %>% 
    filter(junction_reads_stmn2 >= 30) %>% 
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
    mutate(rs12608932 = fct_relevel(rs12608932,
                              "A/A", "A/C", "C/C")) %>%
  filter(!grepl("Cord",tissue_clean)) %>% 
  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>% 
    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>% 
    ggpubr::ggscatter(., 
                      x = "log10_stmn2",
                      y = "log10_unc13a",
                      color = 'rs12608932',                      
                      add = "reg.line", size = 3
                      ) +
    ylab("UNC13A CE PSI ") + 
    xlab("STMN2 Cryptic PSI ") +
  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                      y = unc13a_cryptic_leaf_psi,
                      color = rs12608932),
           method = 'spearman',
           cor.coef.name = 'rho',
           show.legend = FALSE,
           size = 14) + 
    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  theme(legend.title = element_blank()) +
  theme(text = element_text(size = 24),legend.text = element_text(size = 32)) +
      ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI"))) +
   ggtitle("intronic SNP (rs12608932) genotype")

```

```{r}
ggpubr::ggarrange(ce_snp,in_snp)
```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex only

```{r}
detected_table_for_graph = clean_data_table %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>% 
  filter(junction_reads_unc13a >= 30) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))

comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F)

detected_table_for_graph %>% 
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
  geom_boxplot(show.legend = F) + 
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
  stat_compare_means() + 
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5) 
```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex and cord

```{r}
detected_table_for_graph = clean_data_table %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(junction_reads_stmn2 >= 30) %>% 
  filter(junction_reads_unc13a >= 30) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))

comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F)

detected_table_for_graph %>% 
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
  geom_boxplot(show.legend = F) + 
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
  stat_compare_means( label = "p.format") + 
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5) 
```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932

```{r}
detected_table_for_graph_rs12608932 = clean_data_table %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
  filter(junction_reads_stmn2 >= 30) %>% 
  filter(junction_reads_unc13a >= 30) %>% 
  filter(grepl("Cortex",tissue_clean)) %>% 
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>% 
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  group_by(rs12608932) %>% 
  mutate(n_sample = length(unique(sample))) %>% 
  ungroup() %>% 
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )"))

comparisons = combn(unique(detected_table_for_graph_rs12608932$detection_name),2,simplify = F)

detected_table_for_graph_rs12608932 %>% 
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) + 
  geom_boxplot(show.legend = F) + 
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) + 
  stat_compare_means( label = "p.format") + 
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") + 
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12608932 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5) 
```

## Detection rate by genotype

```{r}
overall_fisher = clean_data_table %>%
    filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>% 
  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
  dplyr::select(participant_id,unc13a_detected,rs12973192) %>% 
  unique() %>% 
  group_by(rs12973192) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(unc13a_detected)) %>% 
  dplyr::select(rs12973192,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(n_non_detected = n_sample - n_sample_detected) %>% 
  dplyr::select(-n_sample)

over_p = overall_fisher %>% 
  column_to_rownames('rs12973192') %>% 
  chisq.test() %>% 
  broom::tidy() %>% 
  .$p.value

overall_fisher %>% 
  mutate(n_sample = n_non_detected + n_sample_detected) %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>% 
  ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
  geom_col(show.legend = F) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(labels = scales::percent) + 
  ylab("% of TDP-43 Proteionopathy Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals") + 
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))

```

Although this difference is not significant, with the Fisher's exact giving
a p-value of `r round(over_p,2)`.

# Variant allele expression in TDP-43 KD experiments

Across our KD's, we observed a clear allelic bias in the SK-DZ-N cells

```{r}
kd_vafs = fread(file.path(here::here(),"data","all_kds_unc13.snp.out.tsv"))

kd_vafs %>% 
    left_join(rel_rna_cryptic_amount) %>% 
    filter(!is.na(condition)) %>% 
    mutate(VAF = alt_reads / total) %>% 
    filter(condition != "control" ) %>% 
    filter(alt_reads > 5) %>%  
    ggplot(aes(x = stmn_2_cryptic_psi,y = cryptic_psi, fill = -(VAF - 0.5),size = total)) + 
    geom_point(pch = 21)  + 
    coord_fixed() +
  ylim(0,0.8) + 
  xlim(0,0.8) +
  geom_abline() + 
    scale_fill_gradient2()

```

# Liu Facs Neurons

Also look at the coverage in the Liu nuclear facs sorted neurons

```{r}
liu_vafs = fread(file.path(here::here(),"data","liu_facs_neurons_unc13.snp.out.tsv"))
liu_stmn2 = fread(file.path(here::here(),"data","liu_stmn2_and_unc13aaggregated.clean.annotated.bed"))
liu_stmn2[,sample := gsub(".SJ.out","",V4)]
liu_stmn2 = liu_stmn2 %>% 
  unique() %>% 
    pivot_wider(id_cols = 'sample',
                names_from = "V7",
                values_from = 'V5',
                values_fill = 0)

liu_stmn2 <- liu_stmn2 %>% 
   mutate(stmn2_psi = STMN2_cryptic / (STMN2_cryptic + STMN2_annotated)) %>% 
   mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2))
```

```{r}
liu_vafs %>% 
    mutate(VAF = alt_reads / total) %>% 
    data.table::melt(id.vars = c("sample",'VAF','total')) %>% 
    mutate(allele = ifelse(variable == "ref_reads", "C","G")) %>% 
    mutate(sample_name = sub("_TDP_.*", "", sample)) %>% 
    mutate(sample_name = gsub("_unsorted", "", sample_name)) %>% 
    mutate(condition = sub(".*_", "", sample)) %>% 
    ggplot(aes(x = condition,y = value, fill = allele)) + 
    geom_col(pch = 21, size = 4) +
    facet_wrap(~sample_name,scales = "free")

liu_stmn2 %>% 
  dplyr::select(stmn2_psi,sample,unc13a_psi) %>% 
  data.table::melt() %>% 
   mutate(sample_name = sub("_TDP_.*", "", sample)) %>% 
    filter(!grepl("unsorted",sample_name)) %>% 
    mutate(condition = sub(".*_", "", sample)) %>% 
    mutate(condition = ifelse(grepl("negative",condition), "TDP-43 \nNegative", "TDP-43 \nPositive")) %>% 
    mutate(variable = ifelse(grepl("stmn2",variable), "STMN2", "UNC13A")) %>% 
    ggplot(aes(x = condition,y = value, fill = variable)) + 
    geom_col(size = 4,position = 'dodge') +
    facet_wrap(~sample_name,scales = "free_x") + 
    scale_fill_manual(values = c("#004D40","#882255")) +
    theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    theme(text = element_text(size = 24))  +
      xlab("") +
    ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
```

```{r}
liu_stmn2 %>% 
  dplyr::select(stmn2_psi,sample,unc13a_psi) %>% 
  data.table::melt() %>% 
   mutate(sample_name = sub("_TDP_.*", "", sample)) %>% 
    filter(!grepl("unsorted",sample_name)) %>% 
    mutate(condition = sub(".*_", "", sample)) %>% 
    mutate(condition = ifelse(grepl("negative",condition), "TDP-43 \nNegative", "TDP-43 \nPositive")) %>% 
    mutate(variable = ifelse(grepl("stmn2",variable), "STMN2", "UNC13A")) %>% 
  ggbarplot(,
            x = "condition",
            add = c("mean_se",'jitter'),
            y = "value",
            fill = 'variable',
            color = 'variable',
            position = position_dodge(0.8),dot.size = 6) + 
  ggpubr::theme_pubr() +
  scale_color_manual(values = c("#000000","#000000")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 24))  +
  scale_y_continuous(labels = scales::percent) + 
  xlab("") +
  ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
```

```{r}
liu_stmn2 %>% 
   mutate(stmn2_psi = STMN2_cryptic / (STMN2_cryptic + STMN2_annotated)) %>% 
   mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime)) %>% 
   left_join(liu_vafs) %>% 
    mutate(alt_vaf = (alt_reads)/total) %>% 
    mutate(condition = ifelse(grepl("negative",sample),"negative", "positive")) %>% 
             filter(!grepl("unsorted",sample)) %>%

    ggplot(aes(y = unc13a_psi,x = stmn2_psi, fill = alt_vaf)) + 
    geom_point(pch = 21,size = 5) + 
    ggtitle("STMN2 PSI and the VAF of the G allele \n in the Liu Nuclei ") +
  facet_grid(~condition) + 
  scale_fill_viridis_c(option = "plasma") + 
  coord_fixed() + 
  geom_abline()
```

## Variant allele expression in patients

```{r}
nycg_vafs = fread(file.path(here::here(),"data","NYGC_all_samples_UNC13A_snp_coverage.tsv"),fill = T)
nycg_vafs %>% 
    left_join(clean_data_table) %>% 
    filter(rs12973192 == "C/G") %>% 
    mutate(fraction_risk = alt_reads / total) %>% 
    filter(!is.na(fraction_risk)) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
    geom_point(size = 4, pch = 21, aes(fill = fraction_risk - 0.5)) + 
    xlab("STMN2 Cryptic PSI") +
    ylab("UNC13A CE PSI") +
    geom_abline() + 
    scale_fill_gradient2()

nycg_vafs %>% 
    left_join(clean_data_table) %>% 
    filter(rs12973192 == "C/G") %>% 
    mutate(fraction_risk = alt_reads / total) %>% 
    filter(!is.na(fraction_risk)) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>% 
    filter(unc13a_cryptic_leaf_psi > 0 ) %>% 
    mutate(stmn2_unc_fract = ((unc13a_cryptic_leaf_psi * 100) / (stmn_2_cryptic_psi_leaf * 100))) %>% 
    ggplot(aes(x = stmn2_unc_fract, y = fraction_risk)) +
    geom_point(size = 4, pch = 21, aes(fill = fraction_risk - 0.5)) + 
    xlab("STMN2 Cryptic PSI") +
    ylab("UNC13A CE PSI") +
    scale_fill_gradient2()
```