### 2021-06-28 paper_figures.Rmd
alsod
Code for scrapping ALS mutations from ALSoD website + technical variates

---
title: "Figures for Paper"
author: "Anna-Leigh Brown + Matteo Zanovello + Oscar Wilkins"
date: "06/10/2020"
output:
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'docs/index.html')) })
---

```{css, echo = FALSE}
h1, #TOC>ul>li {
  color: #000000;
    font-weight: bold;
}

h2, #TOC>ul>ul>li {
  color: #41494D;
}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
knitr::opts_chunk$set(
  comment = '', fig.width = 9, fig.height = 9
)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load("data.table",
       "tidyverse",
       "ggplot2",
       "rcompanion",
       "lme4",
       "ggpubr",
       "forcats",
       "scales",
       "here",
       "ggthemes",
       "ggrepel",
       "DESeq2",
       "patchwork",
       "rstatix",
       "ggeasy")
```

# iPSC splicing

Read in the splicing results returned by MAJIQ and make a volcano plot, only highlight
genes of interest with a label.

```{r volcano_plots}

ipsc_splicing = fread(file.path(here::here(),"data","ipsc_splicing_results.csv"))
ipsc_de = fread(file.path(here::here(),"data","ipsc_differential_expression.csv"))

splicing_dots_tables <- ipsc_splicing %>%
    mutate(junction_name = case_when(gene_name %in% c("UNC13A","AGRN",
                                                      "UNC13B","PFKP","SETD5",
                                                      "ATG4B","STMN2") &
                                         p_d_psi_0_10_per_lsv_junction > 0.9  &
                                         deltaPSI > 0 ~ gene_name,
                                     T ~ "")) %>%
    mutate(`Novel Junction` = de_novo_junctions == 0) %>%
    mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>%
    mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>%
    mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>%
    mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                       "UNC13B","PFKP","SETD5",
                                                       "ATG4B","STMN2") &
                                          p_d_psi_0_10_per_lsv_junction > 0.9  &
                                          deltaPSI > 0 ~ junction_name,
                                      T ~ ""))

fig1a = ggplot() +
    geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions != 0),
               aes(x = deltaPSI, y =log10_test_stat,
                   alpha = graph_alpha,,fill = "Annotated Junction"), pch = 21, size = 10) +
    geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions == 0),
               aes(x = deltaPSI, y =log10_test_stat,
                   alpha = graph_alpha,fill = "Novel Junction"), pch = 21, size = 10) +
    geom_text_repel(data = splicing_dots_tables[junction_name != ""],
                    aes(x = deltaPSI, y =log10_test_stat,
                        label = label_junction,
                        color = as.character(de_novo_junctions)), point.padding = 0.3,
                    nudge_y = 0.2,
                    min.segment.length = 0,
                    box.padding  = 2,
                    size=6,show.legend = F) +
    geom_hline(yintercept = -log10(1 - .9)) +
    geom_vline(xintercept = -0,linetype="dotted") +
    scale_fill_manual(name = "",
                      breaks = c("Annotated Junction", "Novel Junction"),
                      values = c("Annotated Junction" = "#648FFF", "Novel Junction" = "#fe6101") ) +
    scale_color_manual(name = "",
                       breaks = c("0", "1"),
                       values = c("1" = "#648FFF", "0" = "#fe6101") ) +
    guides(alpha = FALSE, size = FALSE) +
    theme(legend.position = 'top') +
    ggpubr::theme_pubr() +
    xlab("delta PSI") +
    ylab(expression(paste("-Lo", g[10], " Test Statistic"))) +
    theme(text = element_text(size = 24)) +
    theme(legend.text=element_text(size=22)) +
    xlim(-1,1) +
    scale_x_continuous(labels = scales::percent)

de_table = ipsc_de %>%
    mutate(contains_cryptic = gene_name %in% splicing_dots_tables[cryptic_junction == T,unique(gene_name)]) %>%
    mutate(contains_cryptic = as.character(as.numeric(contains_cryptic))) %>%
    mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                       "UNC13B","PFKP","SETD5",
                                                       "ATG4B","STMN2","TARDBP") ~ gene_name,
                                      T ~ "")) %>%
    mutate(graph_alpha = ifelse(padj < 0.1, 1, 0.2)) %>%
    mutate(y_data = -log10(padj))

fig1b = ggplot(data = de_table) +
    geom_point(data = de_table %>% filter(contains_cryptic == "0"),
               aes(x = log2FoldChange, y = -log10(padj),
                   alpha = graph_alpha,fill = "No Cryptic"), pch = 21, size = 10) +
    geom_point(data = de_table %>% filter(contains_cryptic == "1"),
               aes(x = log2FoldChange, y = -log10(padj),
                   alpha = graph_alpha,fill = "Contains Cryptic"), pch = 21, size = 10)+
    geom_text_repel(data = de_table[label_junction != ""],max.overlaps = 500,
                    aes(x = log2FoldChange,
                        y = -log10(padj),
                        label = label_junction,
                        color = as.character(contains_cryptic)),
                    nudge_y = 5,
                    min.segment.length = 0,
                    box.padding  = 4,
                    size=6,show.legend = F) +
    scale_fill_manual(name = "",
                      breaks = c("No Cryptic", "Contains Cryptic"),
                      values = c("No Cryptic" = "#648FFF", "Contains Cryptic" = "#fe6101") ) +
    scale_color_manual(name = "",
                       breaks = c("1", "0"),
                       values = c("1" = "#fe6101", "0" = "#648FFF") ) +
    xlim(-7.5,7.5) +
    ylab(expression(paste("-Lo", g[10], " P-value"))) +
    guides(alpha = FALSE, size = FALSE) +
    theme(legend.position = 'top') +
    ggpubr::theme_pubr() +
    xlab(expression(paste("Lo", g[2], " Fold Change"))) +
    theme(text = element_text(size = 24)) +
    theme(legend.text=element_text(size=22)) +
    geom_hline(yintercept = -log10(0.1))  +
    geom_vline(xintercept=c(0), linetype="dotted")

print(fig1a)
print(fig1b)
```

<!-- # iPSC WTC11 parent line genotype -->

<!-- ```{r confirm_no_als_ftd_wtc11 mutations} -->
<!-- library(vcf) -->
<!-- wtc11_vcf = "/Users/annaleigh/Downloads/3546dc62_AH77TTBBXX_DS-229105_GCCAAT_recalibrated.vcf.gz" -->
<!-- vcf <- read.vcfR( wtc11_vcf, verbose = FALSE ) -->
<!-- ``` -->

# UNC13A CE PSI across TDP-43 KD experiments

The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192.
The SK-N-DZ cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele.
There was variability on the Klim hMN set on allelic expression.

```{r read in }

rel_rna_cryptic_amount = data.table::fread(file.path(here::here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime +
                                                  UNC13A_5prime +  UNC13A_5prime_2 +
                                                  UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime +
                                                                          UNC13A_5prime +  UNC13A_5prime_2 +
                                                                          UNC13A_5prime_3)]
```

## barplot UNC13A CE PSI - full five

```{r}
rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
                      values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(name = "",
                       values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A CE PSI") +
    xlab("") +
    guides(color = FALSE) +
    scale_y_continuous(labels = scales::percent) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))
```

## barplot UNC13B frameshift PSI

```{r}
rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "condition",
              add = c("mean_se","jitter"),
              y = "unc13b_nmd_exon_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8),facet.by = 'source') +
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
                      values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(name = "",
                       values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B \nNMD Exon PSI") +
    xlab("") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    facet_wrap(~source) +
    stat_compare_means() +
    scale_y_continuous(labels = scales::percent)
```

## TARDBP RNA and UNC13A Cryptic

```{r}

rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12,method = 's',cor.coef.name = 'rho') +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
    scale_y_continuous(labels = scales::percent)
```

## UNC13A RNA and UNC13A Cryptic

```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) +
    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12,
             method = 'spearman',
             cor.coef.name = 'rho') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    scale_y_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
    theme(text = element_text(size = 18,family = 'sans'),
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
```
## UNC13A RNA and UNC13A IR
```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12,cor.coef.name = 'rho',method = 's') +
    geom_smooth(aes(x = UNC13A, y = normalized_unc13a_ir),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom')
```
## TARDBP RNA and UNC13A IR

```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = TARDBP, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = TARDBP, y = normalized_unc13a_ir),size = 12) +
    geom_smooth(aes(x = TARDBP, y = normalized_unc13a_ir),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom')
```

## barplot UNC13A cryptic PSI

```{r}
rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A CE PSI") +
    xlab("") +
    guides(color = FALSE)
```

## barplot STMN2 cryptic PSI

```{r}
stmn2 = rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "stmn_2_cryptic_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("STMN2 Cryptic PSI") +
    xlab("") +
    guides(color = FALSE)
print(stmn2)
```

## barplot UNC13B normalized IR
```{r}
unc13b_ir = rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13b_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13B \nIntron Retention Ratio") +
    xlab("") +
    guides(color = FALSE) +
    theme(legend.position = "none") +
    theme(text = element_text(size = 18)) +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

print(unc13b_ir)
```

## barplot UNC13A normalized IR
```{r}
unc13a_ir = rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13a_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13A \nIntron Retention Ratio") +
    xlab("") +
    theme(legend.position = "none") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

```

## barplot UNC13A normalized RNA Level

```{r}
text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))

rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "UNC13A",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A") +
    xlab("") +
    theme(legend.position = "none") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent) +
    geom_text(data = text_table[gene_name == "UNC13A"],aes(x = source, y= 1.35,label = padj_plot),size = 8) +
    geom_text(data = text_table[gene_name == "UNC13A"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

```

## barplot UNC13B normalized RNA Level

```{r}
text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))

rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "UNC13A",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B") +
    xlab("") +
    theme(legend.position = "none") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent) +
    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.35,label = padj_plot),size = 8) +
    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

```

## barplot TARDBP normalized RNA Level

```{r}
text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))

rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "TARDBP",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("TARDBP") +
    xlab("") +
    theme(legend.position = "none") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent,expand = expansion(mult = c(0, .1))) +
    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.35,label = padj_plot),size = 8) +
    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

```
## Scatterplot stmn2 normalized TARDBP
```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = TARDBP, y = stmn_2_cryptic_psi, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = TARDBP, y = stmn_2_cryptic_psi)) +
    theme(legend.position = 'bottom')
```

## Scatterplot UNC13BIR normalized TARDBP

```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = TARDBP, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = TARDBP, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')
```

## Scatterplot UNC13A-IR and UNC13A Crptic
```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = cryptic_psi_full,
                   y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = cryptic_psi_full,
                 y = normalized_unc13a_ir)) +
    theme(legend.position = 'bottom')
```

## Scatterplot UNC13B-IR and UNC13A Cryptic

```{r}

rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+
    facet_wrap(~source)
```

## scatterplot UNC13BIR and UNC13B NMD

```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+
    facet_wrap(~source)

```

## scatterplot UNC13A CE and UNC13B NMD

```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = cryptic_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+
    facet_wrap(~source)

```
# NMD experiment on SH-SY5Y cells

```{r}

fbp_sy <- read_csv(file.path(here::here(),"data","results_17022021_sy5y.csv"))
fbp_sy$condition <- as.factor(fbp_sy$condition)
fbp_sy$condition <- factor(fbp_sy$condition,levels(fbp_sy$condition)[c(2,1)])

stattest_sy <- fbp_sy %>%
    group_by(gene) %>%
    t_test(value ~ 1, mu = 1) %>%
    add_significance() %>%
    add_xy_position(x="gene")
stattest_sy

nmd_plot_sy <- ggbarplot(fbp_sy,
                         x = 'gene',
                         y = 'value',
                         add = c("mean_se","jitter"),
                         fill = 'condition',
                         color = 'condition',
                         position = position_dodge(0.8),
                         dot.size = 10) +
    scale_y_continuous(labels = scales::percent) +
    ggpubr::theme_pubr() +
    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) +
    scale_color_manual(values = c("#1C2617","#262114")) +
    ylab("Percent of transcript after CHX treatment") +
    xlab("") +
    stat_pvalue_manual(stattest_sy, label="p.signif", y.position = 9)
nmd_plot_sy

```

```{r}

fbp_dz <- read_csv(file.path(here::here(),"data","results_17022021_dz.csv"))
fbp_dz$condition <- as.factor(fbp_dz$condition)
fbp_dz$condition <- factor(fbp_dz$condition,levels(fbp_dz$condition)[c(2,1)])

stattest_dz <- fbp_dz %>%
    group_by(gene) %>%
    t_test(value ~ 1, mu = 1) %>%
    add_significance() %>%
    add_xy_position(x="gene")
stattest_dz

nmd_plot_dz <- ggbarplot(fbp_dz,
                         x = 'gene',
                         y = 'value',
                         add = c("mean_se","jitter"),
                         fill = 'condition',
                         color = 'condition',
                         position = position_dodge(0.8),
                         dot.size = 10) +
    scale_y_continuous(labels = scales::percent) +
    ggpubr::theme_pubr() +
    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) +
    scale_color_manual(values = c("#1C2617","#262114")) +
    ylab("Percent of transcript after CHX treatment") +
    xlab("") +
    stat_pvalue_manual(stattest_dz, label="p.signif", y.position = 26)
nmd_plot_dz

```

# iPSC protein abundance mass spec

```{r}
peptides = fread(file.path(here::here(),"data","peptide_abdundance.csv"))

peptides %>%
    mutate(gene = fct_relevel(gene, "UNC13A","UNC13B")) %>%
    ggbarplot(,
              x = "condition",
              add = c("mean_se"),
              y = "protein",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8),
              facet.by = 'gene') +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Protein abundance") +
    xlab("") +
    geom_jitter(pch = 21,height = 0,aes(fill = condition),size = 2) +
    guides(color = FALSE) +
    facet_wrap(~gene, scales = 'free') +
    theme(legend.position = 'none') +
    scale_y_continuous(labels = scientific) +
    stat_compare_means(comparisons = list(c("Control","TDP-43 KD")),
                       label = 'p.signif',tip.length = 0,
                       size = 10)

```

# Patient summary statistics

```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%
    mutate(rs12973192 = fct_relevel(rs12973192,
                                    "C/C", "C/G", "G/G")) %>%
    mutate(number_g_alleles = as.numeric(rs12973192) - 1) %>%
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>%
    mutate(junction_reads_stmn2 = STMN2_annotated_leaf + STMN2_cryptic_leaf) %>%
    mutate(junction_reads_unc13a = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf + UNC13A_annotated_leaf) %>%
    unique()

print(glue::glue("Number of unique patients: {clean_data_table[,length(unique(participant_id))]}"))
print(glue::glue("Number of unique tissue samples: {clean_data_table[,length(unique(sample))]}"))
print("Patients Per Disease Category")
clean_data_table[,length(unique(participant_id)),by = disease]
print("Tissues Per Disease Category")
clean_data_table[,length(unique(sample)),by = disease]
print("Number of patients per UNC13A SNP genotype")
unique(clean_data_table[,.(participant_id,rs12608932,rs12973192)]) %>% select(-participant_id) %>% table()
print("Number of tissues per disease")
clean_data_table[,.N,by = c("disease","tissue_clean")]
print("Number of partcipants by mutation and  disease")
clean_data_table[,length(unique(participant_id)),by = c("mutations","disease")]

print(glue::glue("Number of patients per pathology:"))
clean_data_table[,length(unique(participant_id)),by = .(pathology)]
```
# UNC13A cryptic is an event that is specific to TDP-43 proteinopathy

FTLD-non-TDP are those with TAU and FUS aggregates

Non-tdp ALS are those with SOD1 or FUS mutations. The n's are quite low on this
unfortunately, only 8 ALS with SOD1 and 2 with FUS mutations.

First we look at detection rate in tissues affected by TDP-43 proteinopathy,
For FTLD this is frontal and temporal Cortices, and
for ALS this is lumbar, cervical, and thoracic spinal cord samples as well as
motor cortex. For controls we also take all 6 tissues,
frontal,temporal,motor cortices and the lumbar, cervical, and thoracic spinal cords.

(As a side note, once we do this the number of ALS-non-TDP drops down to 6 (2 FUS) because
the ALS sample tissues are not balanced and not every participant has samples in every tissue)

```{r Inclusion reads by if TDP-potential}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>%
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>%
    filter(!tissue_clean %in% c("Choroid","Liver")) %>%
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>%
    melt() %>%
    filter(grepl("_3prime|_5prime_1",variable)) %>%
    group_by(sample) %>%
    mutate(inclusion_reads = sum(value)) %>%
    ungroup() %>%
    unique() %>%
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor",
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor",
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>%
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>%
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```
## Pecent of patients UNC13A detected
Looking at disease tissue only, so just taking the cord and motor cortex in ALS and the
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.

```{r}

clean_data_table %>%
  filter(disease_tissue == T) %>%
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(detected = inclusion_reads > 0) %>%
  dplyr::select(participant_id,disease_group2,detected) %>%
  unique() %>% 
  group_by(disease_group2) %>%
  mutate(n_sample = n_distinct(participant_id)) %>%
  mutate(n_sample_detected = sum(detected)) %>%
  dplyr::select(disease_group2,n_sample,n_sample_detected) %>%
  unique() %>%
  mutate(detection_rate = n_sample_detected / n_sample) %>%
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>%
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>%
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) +
  ggpubr::theme_pubr() +
  scale_y_continuous(lim = c(0,1),labels = scales::percent) +
  ylab("Percent of Patients \n UNC13A Cryptic Detected") +
  theme(text = element_text(size = 24)) +
  xlab("N individuals")
```

## Effect of library size on CE discovery

```{r unc detection by library size  }
# library size

clean_data_table %>%
   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  filter(disease_tissue == T) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
  ggplot(aes(x = unc13a_detected,y = log10(library_size) )) + 
  geom_point(colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
  geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
  labs(title = "Library size and UNC13A CE detection ",subtitle = "Wilcox test") + 
  facet_wrap(~ tissue + disease_group2) + 
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
  labs(y = "log10 library size", x = "UNC13A CE detected") +
  theme(axis.text.x = element_text(size = 15)) + 
  ggpubr::theme_pubr()
```

## Effect of UNC13A TPM onCE discovery

```{r unc detection by UNC13A tpm }
# library size

clean_data_table %>%
   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  filter(disease_tissue == T) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
  ggplot(aes(x = unc13a_detected,y = UNC13A_TPM )) + 
  geom_point( colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
  geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
  labs(title = "UNC13A TPM and UNC13A CE detection ",subtitle = "Wilcox test") + 
  facet_wrap(~ tissue + disease_group2, scales = 'free_y') + 
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
  labs(y = "UNC13A expression (TPM)", x = "UNC13A CE detection") +
  theme(axis.text.x = element_text(size = 15)) +
  ggpubr::theme_pubr()

```

## Effect of RIN and UNC13A discovery

```{r rin discovery}
## effect of RIN?
clean_data_table %>%
  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  filter(disease_tissue == T) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  filter(!is.na(rin)) %>%
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
  ggplot(aes(x = unc13a_detected,y = rin )) + 
  geom_point( colour = "firebrick", size = 0.5, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
   geom_boxplot(fill = NA, colour = "black", outlier.colour = NA) +
  labs(title = "Sample RIN and UNC13A CE detection ",subtitle = "Wilcox test") + 
  facet_wrap(~disease_group2 + tissue) + 
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center" ) +
  labs(y = "RIN", x = "UNC13A CE detection") +
  theme(axis.text.x = element_text(size = 15)) +
  ggpubr::theme_pubr()

```

```{r unc13a tpm and ce detection}
## effect of RIN?
clean_data_table %>%
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  add_count(tissue_clean,disease_group2) %>% 
  mutate(plot_name = glue::glue("{disease_group2} \n ({n})")) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
  ggplot(aes(x = plot_name, fill = unc13a_detected,y = UNC13B_TPM )) + 
  geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
  geom_boxplot(outlier.colour = NA) +
  labs(title = "UNC13A CE detection and UNC13B TPM",subtitle = "Wilcox test") + 
  facet_wrap(~ tissue_clean,scales = 'free') + 
  ggpubr::stat_compare_means(label = "p.signif",) +
  labs(y = "UNC13B TPM", x = "UNC13A CE detection") +
  theme(axis.text.x = element_text(size = 15)) +
  ggpubr::theme_pubr()
```

## Effect of celltype composition on UNC13A CE detection

```{r}
conversion_table <- data.frame(short = c("Ast","End", "Mic", "Ex", "Oli", "Per" ),
                               long = c("Astrocytes", "Endothelial", "Microglia", "Neurons", "Oligos", "Pericytes")
  )

cell_types = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/MuSiC_residual_results.tsv")

cell_types %>% 
  left_join(clean_data_table %>%
   filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  filter(disease_tissue == T) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ))

```

## Pecent of samples UNC13A detected by tissue

```{r}
boxplot_table %>%
  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>%
  mutate(detected = inclusion_reads > 0) %>%
  dplyr::select(sample,disease_group2,detected,tissue_clean) %>%
  unique() %>%
  group_by(disease_group2,tissue_clean) %>%
  mutate(n_sample = n_distinct(sample)) %>%
  mutate(n_sample_detected = sum(detected)) %>%
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>%
  unique() %>%
  mutate(detection_rate = n_sample_detected / n_sample) %>%
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>%
  ungroup() %>%
  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>%
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Motor Cortex",
                                    "Thoracic Spinal Cord")) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>%
  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>%
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate,fill = detection_name),position = position_dodge2(width = 0.8, preserve = "single")) +
  ggpubr::theme_pubr() +
  ylab("Percent of Tissues \n UNC13A Cryptic Detected") +
  theme(text = element_text(size = 18)) +
  xlab("N tissues") +
  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) +
  theme(axis.text.x =  element_text(size = 14)) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size = 18))  +
  scale_fill_manual(values = c("#4C97B5","#4C97B5","#4C97B5","#4C97B5","#4C97B5","red","#C87156","#C87156")) +
  theme(legend.position = "none")
```

## Pecent of samples UNC13A detected by tissue and genotype

```{r}

boxplot_table %>%
  left_join(clean_data_table %>% dplyr::select(sample, rs12973192)) %>%
  unique() %>%
  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>%
  mutate(detected = inclusion_reads > 0) %>%
  dplyr::select(sample,disease_group2,detected,tissue_clean,rs12973192) %>%
  unique() %>%
  group_by(disease_group2,tissue_clean,rs12973192) %>%
  mutate(n_sample = n_distinct(sample)) %>%
  mutate(n_sample_detected = sum(detected)) %>%
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>%
  unique() %>%
  mutate(detection_rate = n_sample_detected / n_sample) %>%
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>%
  ungroup() %>%
  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>%
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Motor Cortex",
                                    "Thoracic Spinal Cord")) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>%
  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>%
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate,fill = rs12973192),position = position_dodge2(width = 0.8, preserve = "single")) +
  ggpubr::theme_pubr() +
  ylab("Percent of Tissues \n UNC13A Cryptic Detected") +
  theme(text = element_text(size = 18)) +
  xlab("N tissues") +
  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) +
  theme(axis.text.x =  element_text(size = 14)) +
      theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size = 18))  +
  theme(legend.position = "bottom")
```

## UNC13A CE level across tissue

This only plots the short novel acceptor and novel donor

```{r onlyshort}
boxplot_table %>%
    filter(tissue_clean %in%
               c("Frontal Cortex",
                 "Lumbar Spinal Cord",
                 "Cervical Spinal Cord",
                 "Motor Cortex",
                 "Temporal Cortex",
                 "Cerebellum") )%>%
    group_by(disease_group2,tissue_clean) %>%
    mutate(n_sample = n_distinct(sample)) %>%
    ungroup() %>%
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                      "Frontal Cortex",
                                      "Lumbar Spinal Cord",
                                      "Temporal Cortex",
                                      "Motor Cortex")) %>%
    ggplot(aes(x = detection_name,
               y = inclusion_reads * 10^6,
               fill ="red")) +
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1),outlier.colour = NA) +
    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) +
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) +
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(text = element_text(size = 18))  +
    theme(legend.position = "bottom")
```

## UNC13A CE level across tissue - split by junction type

```{r}
clean_data_table %>%
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>%
    filter(!tissue_clean %in% c("Choroid","Liver")) %>%
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>%
    melt() %>%
    filter(grepl("_3prime|_5prime_1|_5prime_2",variable)) %>%
      # filter(grepl("_5prime_2",variable)) %>%
    unique() %>%
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor",
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor",
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>%
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  dplyr::select(sample,disease_group2,
                tissue_clean,
                junction_name,
                value) %>%
    unique() %>%
    filter(tissue_clean %in%
               c("Frontal Cortex",
                 "Lumbar Spinal Cord",
                 "Cervical Spinal Cord",
                 "Motor Cortex",
                 "Temporal Cortex",
                 "Cerebellum"))%>%
    group_by(disease_group2,tissue_clean) %>%
    mutate(n_sample = n_distinct(sample)) %>%
    ungroup() %>%
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                      "Frontal Cortex",
                                      "Lumbar Spinal Cord",
                                      "Temporal Cortex",
                                      "Motor Cortex")) %>%
    ggplot(aes(x = detection_name,
               y = value * 10^6,
               fill = junction_name)) +
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single")) +
    geom_point(size = 2,pch = 21, alpha = 0.7, position = position_jitterdodge(jitter.width = 0.2))+
    scale_y_log10() +
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) +
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) +
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
        theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size = 18))  +
  theme(legend.position = "bottom")
```

## UNC13A CE level across disease types

```{r}

disease_comparisons = list( c("Control","ALS-TDP"),
                           c("Control","ALS \n non-TDP"),
                           c("Control","FTLD-TDP"),
                           c("Control","FTLD \n non-TDP" ))

table_to_test = boxplot_table %>%
    filter(disease_tissue == T) %>%
    group_by(disease_group2) %>%
    mutate(n_sample = n_distinct(sample)) %>%
    mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>%
    dplyr::select(detection_name,
                  participant_id,
                  inclusion_reads,
                  disease_group2,
                  tissue_clean,
                  disease_tissue) %>%
    unique()

test_pair = pairwise.wilcox.test(table_to_test$inclusion_reads, table_to_test$detection_name,
                     p.adjust.method = "BH") %>% broom::tidy()

test_pair = test_pair %>%
  mutate(p_value_draw = case_when(p.value < 0.0001~ "***",
                                  p.value < 0.01 ~ "**",
                                   p.value < 0.05 ~ "*",
                          TRUE ~ paste0("Adj. p-value \n",as.character(round(p.value,2))))) %>%
  mutate(y.position = seq(0.25,by = 0.1,length.out = 7))

table_to_test %>%
    ggplot(aes(x = detection_name, y = inclusion_reads * 10^6)) +
    geom_boxplot() +
    geom_jitter(height = 0) +
    scale_y_log10() +
    ggpubr::theme_pubr() +
    ylab("UNC13A cryptic inclusion \n reads per million") +
    theme(text = element_text(size = 24)) +
    xlab("N samples") +
    stat_pvalue_manual(test_pair %>% filter(p.value < 0.05),
                       label = "p_value_draw",size = 8) +
   stat_compare_means(size = 8)

```

# UNC13A RNA expression (TPM)

## UNC13A TPM by disease and tissue

```{r}
comps = clean_data_table %>%
  group_by(disease_group2,tissue_clean) %>%
  mutate(n_sample = n_distinct(sample)) %>%
  ungroup() %>%
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% pull(detection_name) %>%
  unique() %>%
  combn(.,2,simplify = F)

clean_data_table %>%
  group_by(disease_group2,tissue_clean) %>%
  mutate(n_sample = n_distinct(sample)) %>%
  ungroup() %>%
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
 filter(tissue_clean %in%
           c("Frontal Cortex",
             "Lumbar Spinal Cord",
             "Cervical Spinal Cord",
             "Motor Cortex",
             "Temporal Cortex",
             "Cerebellum") )%>%
mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                              "Frontal Cortex",
                              "Lumbar Spinal Cord",
                              "Temporal Cortex",
                              "Motor Cortex")) %>%
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) +
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") +
    xlab("") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none') +
  stat_compare_means(comparisons = comps)
```

```{r make_those_plots_function}
plot_by_tissue = function(tissue_name,dt){

only_tissue = dt %>%
  group_by(disease_group2,tissue_clean) %>%
  mutate(n_sample = n_distinct(sample)) %>%
  ungroup() %>%
  filter(tissue_clean == tissue_name) %>%
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean))

unique_x_values = only_tissue %>%
  pull(detection_name) %>%
  unique()

comparisons = combn(unique_x_values,2,simplify = F)

the_plot = only_tissue %>%
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) +
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") +
    xlab("") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none') +
  stat_compare_means() +
  stat_compare_means(label = "p.signif",comparisons = comparisons,)

  return(the_plot)
}

dis_tis = clean_data_table[disease_tissue == T,unique(tissue_clean)]
dis_tis = c("Cerebellum",dis_tis)
my_plots = purrr::map(dis_tis,plot_by_tissue, clean_data_table)
ggarrange(plotlist=my_plots)

```

## Correlation STMN2 Cryptic PSI and UNC13A TPM

only in ALS/FTLD-TDP

```{r}
clean_data_table %>%
    filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) +
  geom_point() +
  stat_cor(size = 8) +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A TPM") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  ylim(0,50)
```

## Correlation STMN2 Cryptic PSI and UNC13A TPM - tissue separations

```{r}

clean_data_table %>%
    filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) +
  geom_point() +
  stat_cor(size = 8) +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A TPM") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = "free_y")
```

## Correlation between UNC13A CE PSI and UNC13A TPM

```{r}
clean_data_table %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(UNC13A_annotated_leaf > 15) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(unc13a_cryptic_leaf_psi > 0) %>%
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13B_TPM)) +
  geom_point() +
  stat_cor(label.y = 20,size = 8,method = 's',cor.coef.name = 'rho') +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 22)) +
  scale_x_continuous(labels = scales::percent)

```

## Correlation between UNC13A CE PSI and UNC13A TPM - tissue separated

```{r}
clean_data_table %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(UNC13A_annotated_leaf > 15) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM, color = disease_group2)) +
  geom_point() +
  stat_cor(size = 8) +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = 'free') +
  scale_x_continuous(labels = scales::percent)

```

# Relationship between STMN2 and UNC13A CE PSI
## Scatter plot showing the correlation in non-log space in cortex UNC13A & STMN2 CE

```{r}
clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>%
  # filter(STMN2_cryptic_leaf >= 2) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(unc13a_cryptic_leaf_psi > 0) %>%
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
  geom_point() +
  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A CE PSI") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
```

# Relationship between STMN2 and UNC13B fsE PSI
## Scatter plot showing the correlation in non-log space in cortex UNC13B fsE & STMN2 CE

```{r}
clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  mutate(junction_reads_unc13b = UNC13B_nmd_junction_5prime + UNC13B_annotated + UNC13B_nmd_junction_3prime) %>% 
  filter(junction_reads_unc13b >= 30) %>%
  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>%
  # filter(STMN2_cryptic_leaf >= 2) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(unc13a_cryptic_leaf_psi > 0) %>%
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = unc13b_nmd_psi)) +
  geom_point() +
  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("UNC13A CE PSI") +
  ylab("UNC13B fsE PSI") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
```

## Side by side comparison UNC13A & STMN2 CE PSI
```{r}
clean_data_table %>%
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
    filter(tissue_clean %in%
               c("Frontal Cortex",
                 "Lumbar Spinal Cord",
                 "Cervical Spinal Cord",
                 "Motor Cortex",
                 "Temporal Cortex",
                 "Cerebellum") )%>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(disease_tissue == T) %>%
  filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 1) %>%
  filter(STMN2_cryptic_leaf >= 1) %>%
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  melt() %>%
    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Temporal Cortex")) %>%
  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(aes(color = variable),outlier.shape = NA) +
  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
  scale_fill_manual(values = c("#008574","#cd5582")) +
  scale_color_manual(values = c("#008574","#cd5582")) +
  facet_wrap(~tissue_clean,nrow = 3, scales = 'free') +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("CE PSI")  +
  theme(legend.position = 'none')

```

## Side by side comparison UNC13A & STMN2 CE PSI v2
```{r}
clean_data_table %>%
  filter(UNC13A_annotated_leaf > 10) %>%
  filter(STMN2_annotated_leaf > 10) %>%
  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  # filter(unc13a_cryptic_leaf_psi > 0) %>%
  # filter(disease_tissue == T) %>%
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  filter(tissue_clean %in% c("Motor Cortex", "Cervical Spinal Cord", "Frontal Cortex", "Lumbar Spinal Cord", "Temporal Cortex")) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_rat = log2((unc13a_cryptic_leaf_psi + 0.1) / (stmn_2_cryptic_psi_leaf + 0.1))) %>%
  group_by(tissue_clean, disease_group2) %>%
  mutate(average_log2Fold = mean(log2_rat)) %>%
  ungroup() %>%
  dplyr::select(-log2_rat) %>%
  pivot_longer(cols = c("stmn_2_cryptic_psi_leaf","unc13a_cryptic_leaf_psi")) %>%
  mutate(tissue_clean = fct_reorder(tissue_clean,average_log2Fold)) %>%
  mutate(variable = ifelse(grepl("stmn_2",name), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = tissue_clean, y = value,fill = name)) +
  geom_boxplot() +
  geom_point(pch = 21,height = 0,width = 0.2,position = position_jitterdodge()) +
  scale_fill_manual(values = c("#004D40","#882255")) +
  facet_wrap(~tissue_clean+disease_group2,nrow = 1,scales = 'free')  +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position = 'none')
```

## Side by side comparison UNC13A & STMN2 CE PSI - v3
```{r}
clean_data_table %>%
  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  # filter(unc13a_cryptic_leaf_psi > 0) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  melt() %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Temporal Cortex")) %>%
  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(aes(color = variable)) +
  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
  scale_fill_manual(values = c("#004D40","#882255")) +
  scale_color_manual(values = c("#004D40","#882255")) +
  facet_wrap(~tissue_clean+disease_group2,nrow = 3,scales = 'free') +
    scale_y_continuous(labels = scales::percent)  +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position = 'none')

```

## Log2FC UNC13A & STMN2 CE PSI
```{r}
clean_data_table %>%
  filter(UNC13A_annotated_leaf > 10) %>%
  filter(STMN2_annotated_leaf > 10) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(unc13a_cryptic_leaf_psi > 0) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf)))%>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%

  # mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = tissue_clean,y = log2_unc, fill = tissue_clean)) +
  geom_boxplot(position = 'dodge') +
  # geom_jitter(height = 0,width = 0.2,pch = 21) +
  facet_wrap(~disease_group2,scales = "free_x") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
```

# Effect of UNC13A genotype

## UNC13A psi by genotype - rs12973192 - RE

```{r}
rs12973192_comparisons = clean_data_table %>%
  filter(junction_reads_unc13a >=30) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )"))

comp_for_gg = rs12973192_comparisons %>% pull(detection_name) %>%
  unique() %>%
  combn(.,2,simplify = F)

ce_psi = rs12973192_comparisons %>%
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  ylab("UNC13A PSI") +
  xlab(element_blank()) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means() +
  theme(legend.position = "none") +
  stat_compare_means(comparisons = comp_for_gg,label = "p.signif") +
  scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
  ggtitle("CE SNP (rs12973192)")

```

## UNC13A psi by genotype - rs12608932 - RI
```{r}
rs12608932_comparisons = clean_data_table %>%
  filter(junction_reads_unc13a >=30) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  group_by(rs12608932) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )"))

comp_for_gg = rs12608932_comparisons %>% pull(detection_name) %>%
  unique() %>%
  combn(.,2,simplify = F)

in_psi = rs12608932_comparisons %>%
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12608932)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  ylab(element_blank()) +
  xlab(element_blank()) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(legend.position = "none") +
  theme(text = element_text(size = 24)) +
  stat_compare_means() +
  stat_compare_means(comparisons = comp_for_gg,label = "p.signif") +
    scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
  ggtitle("intronic SNP (rs12973192)")
```

```{r}
ggpubr::ggarrange(ce_psi,in_psi)
```

## UNC13A psi by genotype - rs12973192
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## UNC13A psi by genotype - tissue split
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## UNC13A psi in detected tissues by genotype - tissue split
```{r}

clean_data_table %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi by genotype
```{r}

clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi by genotype - tissue split
```{r}

clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi in detected tissues by genotype - tissue split
```{r}

clean_data_table %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## Correlation UNC13A PSI and STMN2 PSI with genotype effect - rs12973192

```{r}

ce_snp = clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
    mutate(rs12973192 = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>%
  filter(!grepl("Cord",tissue_clean)) %>%
  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>%
    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>%
    ggpubr::ggscatter(.,
                      x = "log10_stmn2",
                      y = "log10_unc13a",
                      color = 'rs12973192',
                      add = "reg.line", size = 3
                      ) +
    ylab("UNC13A CE PSI ") +
    xlab("STMN2 Cryptic PSI ") +
  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                      y = unc13a_cryptic_leaf_psi,
                      color = rs12973192),
           method = 'spearman',
           cor.coef.name = 'rho',
           show.legend = FALSE,
           size = 14) +
    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  theme(legend.title = element_blank()) +
  theme(text = element_text(size = 24),legend.text = element_text(size = 32)) +
    ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI"))) +
  ggtitle("CE SNP (rs12973192) genotype")

```

## Correlation UNC13A PSI and STMN2 PSI with genotype effect - rs12608932

```{r}

in_snp = clean_data_table %>%
  filter(disease_tissue == T) %>%
    filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
    mutate(rs12608932 = fct_relevel(rs12608932,
                              "A/A", "A/C", "C/C")) %>%
  filter(!grepl("Cord",tissue_clean)) %>%
  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>%
    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>%
    ggpubr::ggscatter(.,
                      x = "log10_stmn2",
                      y = "log10_unc13a",
                      color = 'rs12608932',
                      add = "reg.line", size = 3
                      ) +
    ylab("UNC13A CE PSI ") +
    xlab("STMN2 Cryptic PSI ") +
  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                      y = unc13a_cryptic_leaf_psi,
                      color = rs12608932),
           method = 'spearman',
           cor.coef.name = 'rho',
           show.legend = FALSE,
           size = 14) +
    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  theme(legend.title = element_blank()) +
  theme(text = element_text(size = 24),legend.text = element_text(size = 32)) +
      ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI"))) +
   ggtitle("intronic SNP (rs12608932) genotype")

```

```{r}
ggpubr::ggarrange(ce_snp,in_snp)
```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex only

```{r}
detected_table_for_graph = clean_data_table %>%
  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3)) %>% as.data.table()

#get out that person who is discongruous

comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F)

detected_table_for_graph %>%
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) +
  geom_boxplot(show.legend = F) +
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
  stat_compare_means() +
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") +
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5)
```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex and cord

```{r}
detected_table_for_graph = clean_data_table %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))

comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F)

detected_table_for_graph %>%
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) +
  geom_boxplot(show.legend = F) +
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
  stat_compare_means( label = "p.format") +
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") +
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5)
```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932

```{r}
detected_table_for_graph_rs12608932 = clean_data_table %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
  group_by(rs12608932) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )"))

comparisons = combn(unique(detected_table_for_graph_rs12608932$detection_name),2,simplify = F)

detected_table_for_graph_rs12608932 %>%
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) +
  geom_boxplot(show.legend = F) +
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
  stat_compare_means( label = "p.format") +
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") +
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12608932 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5)
```

## Detection rate by genotype

```{r}
overall_fisher = clean_data_table %>%
    filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>%
  dplyr::select(participant_id,unc13a_detected,rs12973192) %>%
  unique() %>%
  group_by(rs12973192) %>%
  mutate(n_sample = n_distinct(participant_id)) %>%
  mutate(n_sample_detected = sum(unc13a_detected)) %>%
  dplyr::select(rs12973192,n_sample,n_sample_detected) %>%
  unique() %>%
  mutate(n_non_detected = n_sample - n_sample_detected) %>%
  dplyr::select(-n_sample)

over_p = overall_fisher %>%
  column_to_rownames('rs12973192') %>%
  chisq.test() %>%
  broom::tidy() %>%
  .$p.value

overall_fisher %>%
  mutate(n_sample = n_non_detected + n_sample_detected) %>%
  mutate(detection_rate = n_sample_detected / n_sample) %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
  geom_col(show.legend = F) +
  ggpubr::theme_pubr() +
  scale_y_continuous(labels = scales::percent) +
  ylab("% of TDP-43 Proteionopathy Patients \n UNC13A Cryptic Detected") +
  theme(text = element_text(size = 18)) +
  xlab("N individuals") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))

```

Although this difference is not significant, with the Fisher's exact giving
a p-value of `r round(over_p,2)`.

# Variant allele expression in TDP-43 KD experiments

Across our KD's, we observed a clear allelic bias in the SK-DZ-N cells

```{r}
kd_vafs = fread(file.path(here::here(),"data","all_kds_unc13.snp.out.tsv"))

kd_vafs %>%
    left_join(rel_rna_cryptic_amount) %>%
    filter(!is.na(condition)) %>%
    mutate(VAF = alt_reads / total) %>%
    filter(condition != "control" ) %>%
    filter(alt_reads > 5) %>%
    ggplot(aes(x = stmn_2_cryptic_psi,y = cryptic_psi, fill = -(VAF - 0.5),size = total)) +
    geom_point(pch = 21)  +
    coord_fixed() +
  ylim(0,0.8) +
  xlim(0,0.8) +
  geom_abline() +
    scale_fill_gradient2()

```

# Liu Facs Neurons

Also look at the coverage in the Liu nuclear facs sorted neurons

```{r}
liu_vafs = fread(file.path(here::here(),"data","liu_facs_neurons_unc13.snp.out.tsv"))
liu_stmn2 = fread(file.path(here::here(),"data","liu_stmn2_and_unc13aaggregated.clean.annotated.bed"))
liu_stmn2[,sample := gsub(".SJ.out","",V4)]
liu_stmn2 = liu_stmn2 %>%
  unique() %>%
    pivot_wider(id_cols = 'sample',
                names_from = "V7",
                values_from = 'V5',
                values_fill = 0)

liu_stmn2 <- liu_stmn2 %>%
   mutate(stmn2_psi = STMN2_cryptic / (STMN2_cryptic + STMN2_annotated)) %>%
   mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2))
```

```{r}
liu_vafs %>%
    mutate(VAF = alt_reads / total) %>%
    data.table::melt(id.vars = c("sample",'VAF','total')) %>%
    mutate(allele = ifelse(variable == "ref_reads", "C","G")) %>%
    mutate(sample_name = sub("_TDP_.*", "", sample)) %>%
    mutate(sample_name = gsub("_unsorted", "", sample_name)) %>%
    mutate(condition = sub(".*_", "", sample)) %>%
    ggplot(aes(x = condition,y = value, fill = allele)) +
    geom_col(pch = 21, size = 4) +
    facet_wrap(~sample_name,scales = "free")

liu_stmn2 %>%
  dplyr::select(stmn2_psi,sample,unc13a_psi) %>%
  data.table::melt() %>%
   mutate(sample_name = sub("_TDP_.*", "", sample)) %>%
    filter(!grepl("unsorted",sample_name)) %>%
    mutate(condition = sub(".*_", "", sample)) %>%
    mutate(condition = ifelse(grepl("negative",condition), "TDP-43 \nNegative", "TDP-43 \nPositive")) %>%
    mutate(variable = ifelse(grepl("stmn2",variable), "STMN2", "UNC13A")) %>%
    ggplot(aes(x = condition,y = value, fill = variable)) +
    geom_col(size = 4,position = 'dodge') +
    facet_wrap(~sample_name,scales = "free_x") +
    scale_fill_manual(values = c("#004D40","#882255")) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(text = element_text(size = 24))  +
      xlab("") +
    ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
```

```{r}
liu_stmn2 %>%
  dplyr::select(stmn2_psi,sample,unc13a_psi) %>%
  data.table::melt() %>%
   mutate(sample_name = sub("_TDP_.*", "", sample)) %>%
    filter(!grepl("unsorted",sample_name)) %>%
    mutate(condition = sub(".*_", "", sample)) %>%
    mutate(condition = ifelse(grepl("negative",condition), "TDP-43 \nNegative", "TDP-43 \nPositive")) %>%
    mutate(variable = ifelse(grepl("stmn2",variable), "STMN2", "UNC13A")) %>%
  ggbarplot(,
            x = "condition",
            add = c("mean_se",'jitter'),
            y = "value",
            fill = 'variable',
            color = 'variable',
            position = position_dodge(0.8),dot.size = 6) +
  ggpubr::theme_pubr() +
  scale_color_manual(values = c("#000000","#000000")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size = 24))  +
  scale_y_continuous(labels = scales::percent) +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
```

```{r}
liu_stmn2 %>%
   mutate(stmn2_psi = STMN2_cryptic / (STMN2_cryptic + STMN2_annotated)) %>%
   mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime)) %>%
   left_join(liu_vafs) %>%
    mutate(alt_vaf = (alt_reads)/total) %>%
    mutate(condition = ifelse(grepl("negative",sample),"negative", "positive")) %>%
             filter(!grepl("unsorted",sample)) %>%

    ggplot(aes(y = unc13a_psi,x = stmn2_psi, fill = alt_vaf)) +
    geom_point(pch = 21,size = 5) +
    ggtitle("STMN2 PSI and the VAF of the G allele \n in the Liu Nuclei ") +
  facet_grid(~condition) +
  scale_fill_viridis_c(option = "plasma") +
  coord_fixed() +
  geom_abline()
```

## Variant allele expression in patients

```{r}
nycg_vafs = fread(file.path(here::here(),"data","NYGC_all_samples_UNC13A_snp_coverage.tsv"),fill = T)
nycg_vafs %>%
    left_join(clean_data_table) %>%
    filter(rs12973192 == "C/G") %>%
    mutate(fraction_risk = alt_reads / total) %>%
    filter(!is.na(fraction_risk)) %>%
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
    geom_point(size = 4, pch = 21, aes(fill = fraction_risk - 0.5)) +
    xlab("STMN2 Cryptic PSI") +
    ylab("UNC13A CE PSI") +
    geom_abline() +
    scale_fill_gradient2()

nycg_vafs %>%
    left_join(clean_data_table) %>%
    filter(rs12973192 == "C/G") %>%
    mutate(fraction_risk = alt_reads / total) %>%
    filter(!is.na(fraction_risk)) %>%
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
    mutate(stmn2_unc_fract = ((unc13a_cryptic_leaf_psi * 100) / (stmn_2_cryptic_psi_leaf * 100))) %>%
    ggplot(aes(x = stmn2_unc_fract, y = fraction_risk)) +
    geom_point(size = 4, pch = 21, aes(fill = fraction_risk - 0.5)) +
    xlab("STMN2 Cryptic PSI") +
    ylab("UNC13A CE PSI") +
    scale_fill_gradient2()
```

# Ribosome profiling
## Ribosome profiling quality control

TODO: convert all libraries to p_load.

Perform quality control to check that reads are periodic.

```{r message=FALSE, warning=FALSE}
read_length_filter <- 29  # length of reads of interest
gencode_v29_info <- read_tsv(paste0(here(), "/data/longest_proteincoding_transcript_hs_details.txt"))
ribo_df <- read_tsv(paste0(here(), "/data/riboseq/sh_tdpb.Aligned.toTranscriptome.out.bam.bed.gz"),
                    col_names = c("transcript_id", "start0", "end", "strand")) %>%
    inner_join(gencode_v29_info, by = "transcript_id") %>%
    dplyr::filter(end-start0 == read_length_filter, strand == "+") %>%
    mutate(distance_from_start = start0-cds_start) %>%
    group_by(distance_from_start) %>%
    mutate(n_this_dist = n()) %>%
    dplyr::select(distance_from_start, n_this_dist) %>%
    unique()

```

```{r, warning=FALSE}
ggplot(ribo_df, aes(x=distance_from_start, y=n_this_dist)) +
    geom_line() +
    geom_point() +
    xlim(-50, 50) +
    ylab("Counts") +
    xlab("Distance of 5' end from annotated start codon") +
    ggtitle(paste0("Periodicity of ", read_length_filter, "nt reads from TDP-43 KD replicate B"))
```

## Ribosome profiling volcano plot

Use DESeq to detect differential ribosome footprints and create volcano plots

```{r message = FALSE, warning=FALSE}
counts_df <- read_csv(paste0(here(), "/data/riboseq/cds_feature_counts.csv"))
counts <- as.matrix(counts_df %>% column_to_rownames("gene_id"))
ipsc_splicing <- read_csv(paste0(here(), "/data/ipsc_splicing_results.csv"))
ipsc_de = read_csv(paste0(here(), "/data/ipsc_differential_expression.csv"))

col_data <- data.frame(sample = colnames(counts)) %>%
  mutate(condition = ifelse(str_detect(sample, "tdp"), "tdp_ko", "control"))

dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = col_data,
                              design= ~condition)
dds <- DESeq(dds)
results_name <- resultsNames(dds)[2]
res <- results(dds, name=results_name)

res_df <- data.frame(res) %>%
  rownames_to_column("gene_id") %>%
  left_join(gencode_v29_info, by = "gene_id")

cryptic_df <- ipsc_splicing %>%
  dplyr::select(gene_name, cryptic_junction) %>%
  unique() %>%
  filter(cryptic_junction)

res_df2 <- res_df %>%
  mutate(cryptic_junction = gene_name %in% cryptic_df$gene_name) %>%
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2","TARDBP") ~ gene_name)) %>%
  mutate(colour = ifelse(cryptic_junction, "#fe6101",  "#648FFF")) %>%
  arrange(colour)
```

```{r, warning=FALSE}
p1 <- ggplot(res_df2, aes(x = log2FoldChange, y = -log10(pvalue), label = label_junction)) +
  geom_point(data = res_df2 %>% filter(!cryptic_junction),
             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#648FFF") +
  geom_point(data = res_df2 %>% filter(cryptic_junction),
             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#fe6101") +
  ggpubr::theme_pubr() +
  ylab("-Log10(p-value)") +
  xlab("Log2 fold change") +
  geom_text_repel(box.padding = 1.2, max.overlaps = Inf)

p2 <- p1 + ylim(0,10)
p1
p2
```

# Targeted RNA-seq

Read in data from processed bam files

TODO: add scripts which deduplicate UMIs for each.

```{r message=FALSE, warning=FALSE}
specific_rt <- read_csv(paste0(here(), "/data/targeted_RNA_seq/specific_RT_primer/for_plot.csv")) %>%
    select(risk, sample_name, n)

random_hexamer <- read_csv(paste0(here(), "/data/targeted_RNA_seq/random_hexamer/values_for_plot.csv")) %>%
    select(-sample_name) %>%
    select(sample_name = actual_sample_name, risk, n)

tg_rnaseq_df <- bind_rows(specific_rt, random_hexamer) %>%
  group_by(sample_name) %>%
  mutate(n_healthy = ifelse(risk == "Non-Risk", n, 0)) %>%
  mutate(combined_healthy = sum(n_healthy)) %>%
  select(-n_healthy) %>%
  mutate(combined = sum(n)) %>%
  group_by(sample_name, risk) %>%
  mutate(combined_n = sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  unique() %>%
  group_by(sample_name) %>%
  mutate(p_value = pbinom(combined_healthy, sum(combined_n), 0.5)) %>%  # single-tailed bionomial test
  ungroup() %>%
  mutate(sample_name_ordered = fct_reorder(sample_name, dplyr::desc(combined_n)))

ggplot(tg_rnaseq_df %>%
         filter(combined_n > 0), aes(x = sample_name_ordered, y = combined_n, fill = risk)) +
  geom_bar(stat="identity", position="dodge") +
  geom_text(aes(x=sample_name, y=-2.5, label=signif(p_value,2)), size = 3.5) +
  xlab("") +
  ylab("Unique cDNAs") +
  ggeasy::easy_add_legend_title("Allele") +
  theme_minimal() +
  ggpubr::theme_pubr() +
  ggeasy::easy_rotate_x_labels()
```

# iCLIP of minigenes

Read in the processed data (following alignment with Bowtie2, and filtering for high-confidence alignments), and generate
a plot of the overall distribution of crosslinks, and the rolling average difference between the 2x Risk and 2x Healthy.

Next, filter for peaks in the iCLIP signal, and see how they change in 2x Risk versus 2x Healthy. Highlight those which are close to the SNPs.

TODO: add script which does this filtering
TODO: add fasta files for the minigenes

```{r message=FALSE, warning=FALSE}
exon_snp <- 556
intron_snp <- 1106
intron_start <- 141
intron_end <- 1429
acceptor1 <- 394
acceptor2 <- 444
donor <- 572
mg_l <- 1624
cpoor_start <- 490
cpoor_end <- 1325
risk_colour <- "#00BFC4"
healthy_colour <- "#F8766D"
confidence_interval <- 0.75  # fracion, i.e. 0.75 = 75%.
peak_threshold <- 5 # how many times above the mean to be a "peak"
region_length <- 50

filtered_df <- read_csv(paste0(here(), "/data/iCLIP/bowtie2_aligned_2_filtered_df.csv"))

rolling_window <- 20
peak_t <- peak_threshold*1000/mg_l  # by definition...

df2 <- filtered_df %>%
  ungroup() %>%
  group_by(condition, pos) %>%
  mutate(mean_counts = sum(normalised_counts)/2) %>%
  select(pos, condition, mean_counts) %>%
  unique() %>%
  pivot_wider(names_from = "condition", values_from ="mean_counts")

df2[is.na(df2)] <- 0

df2 <- df2 %>%
  mutate(diff = R-H) %>%
  mutate(mean_4 = (H+R)/2) %>%
  arrange(pos)

top <- ggplot(df2, aes(x=pos, y=diff)) +
  geom_area(aes(x=pos, y = mean_4), colour="grey50", fill = "grey50", alpha=0.2) +
  geom_vline(xintercept = exon_snp, linetype="dashed") +
  geom_vline(xintercept = intron_snp, linetype="dashed") +
  xlim(1,1600) +
  theme_minimal() +
  ylab("XLs per 1000") +
  xlab("") +
  geom_vline(xintercept = cpoor_start) +
  geom_vline(xintercept = cpoor_end)

bottom <- ggplot(df2, aes(x=pos, y=zoo::rollmean(diff, k = rolling_window, na.pad=T))) +
  geom_area(colour = "grey50", fill="grey50", alpha = 1) +
  geom_vline(xintercept =  exon_snp, linetype="dashed") +
  geom_vline(xintercept =  intron_snp, linetype="dashed") +
  xlim(1,1600) +
  ylim(-0.5,0.75) +
  theme_minimal() +
  ylab("Change in XL density") +
  xlab("Distance from start of minigene") +
  geom_vline(xintercept =  intron_start) +
  geom_vline(xintercept =  intron_end) +
  geom_vline(xintercept =  acceptor1) +
  geom_vline(xintercept =  acceptor2) +
  geom_vline(xintercept =  donor)

combined <- top/bottom
combined

find_ratio_bound <- function(e_x, sem_x, e_y, sem_y, quantile_value, randoms = 10000){
  # monte carlo confidence interval estimate for a ratio. Assumes the two variables
  # x and y are normally distributed. Finds confidence interval for (x-y)/y

  # randomly generate possible global means for x and y

  xs = rnorm(randoms, mean = e_x, sd=sem_x)  # risk
  ys = rnorm(randoms, mean = e_y, sd=sem_y)  # healthy

  ratios = (xs-ys)/ys

  return(quantile(ratios, quantile_value))
}

peak_df <- filtered_df %>%
  ungroup() %>%
  group_by(pos) %>%
  mutate(max_at_pos = max(normalised_counts)) %>%  # at least 1 of the 4 samples above the threshold
  ungroup() %>%
  filter(max_at_pos >= peak_t) %>%  # at least 1 of the 4 samples above the threshold
  group_by(condition, pos) %>%
  mutate(peak_mean = mean(normalised_counts),
         peak_sd = sd(normalised_counts)) %>%
  ungroup() %>%
  select(condition, pos, peak_mean, peak_sd) %>%
  unique() %>%
  pivot_wider(names_from = "condition", values_from = c("peak_mean", "peak_sd")) %>%
  mutate(frac_increase = (peak_mean_R-peak_mean_H)/peak_mean_H) %>%
  mutate(peak_rank = rank(frac_increase)) %>%
  mutate(colour = ifelse(abs(pos-exon_snp)<region_length, "gold",
                         ifelse(abs(pos-intron_snp)<region_length, "purple", "grey50")))

peak_df$upper_bound <- mapply(find_ratio_bound,
                        peak_df$peak_mean_R,
                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
                        peak_df$peak_mean_H,
                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
                        0.5*(1+confidence_interval))

peak_df$lower_bound <- mapply(find_ratio_bound,
                        peak_df$peak_mean_R,
                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
                        peak_df$peak_mean_H,
                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
                        0.5*(1-confidence_interval))

ggplot(peak_df, aes(x = peak_rank, y = 100*frac_increase, fill=colour)) +
  geom_bar(stat="identity") +
  scale_fill_identity() +
  theme_minimal() +
  ggpubr::theme_pubr()+
  ylab("% change in 2x Risk") +
  xlab("Rank") +
  geom_errorbar(aes(ymin = 100*lower_bound, ymax = 100*upper_bound, x = peak_rank),
                width = 0.3)

```

Supplementary iCLIP Figures

```{r message=FALSE, warning=FALSE}

bars <- filtered_df %>%
  ungroup() %>%
  group_by(condition, pos) %>%
  mutate(mean_normalised = mean(normalised_counts),
         sd_normalised = sd(normalised_counts)) %>%
  mutate(colour = ifelse(condition == "H", healthy_colour, risk_colour))

bar_top <- ggplot(bars %>% filter(abs(pos-exon_snp) < 35), aes(x = pos, y = mean_normalised,
                                                       fill = colour)) +
  geom_bar(stat="identity", position="dodge") +
  scale_fill_identity() +
  geom_errorbar(aes(ymin = mean_normalised-sd_normalised, ymax=mean_normalised+sd_normalised),
                position="dodge") +
  geom_vline(xintercept = exon_snp, linetype="dashed") +
  ylab("XLs per 1000") +
  ggpubr::theme_pubr() +
  ggtitle("iCLIP signal around CE SNP") +
  xlab("")

bar_bottom <- ggplot(bars %>% filter(abs(pos-intron_snp) < 35), aes(x = pos, y = mean_normalised,
                                                       fill = colour)) +
  geom_bar(stat="identity", position="dodge") +
  scale_fill_identity() +
  geom_errorbar(aes(ymin = mean_normalised-sd_normalised, ymax=mean_normalised+sd_normalised),
                position="dodge") +
  geom_vline(xintercept = intron_snp, linetype="dashed") +
  ggpubr::theme_pubr() +
  ylab("XLs per 1000") +
  ggtitle("iCLIP signal around intronic SNP") +
  xlab("")

bar_combined <- bar_top / bar_bottom
bar_combined

```

# Heptameric analysis

Analyse E-values (higher E value = stronger binding enrichment) for the motifs present at the exon SNP.

```{r message=FALSE, warning=FALSE}

# Read in supplementary 4 from Ray et al 2013

data <- read_tsv(paste0(here(), "/data/heptamers/41586_2013_BFnature12311_MOESM334_ESM.txt"))

for(type in c("exon", "intron")){

    if(type == "exon"){
        snp <-     "TAAAAGCATGGAT"
        healthy <- "TAAAAGGATGGAT"
    } else {
        snp <-     "GGATGGGTGGATA"
        healthy <- "GGATGGTTGGATA"
    }

    all_healthy <- c()
    all_risk <- c()

    # Make dataframe containing E scores for each relevant 7mer

    for(i in 1:(str_length(snp)-6)){
        this_snp <- str_sub(snp, i, i+6)
        this_healthy <- str_sub(healthy, i, i+6)

        all_healthy <- c(all_healthy, str_replace_all(this_healthy, "T", "U"))
        all_risk <- c(all_risk, str_replace_all(this_snp, "T", "U"))

        row1 <- data %>% filter(`7mer` == str_replace_all(this_snp, "T", "U")) %>%
            mutate(snp = T, pos = i)

        row2 <- data %>% filter(`7mer` == str_replace_all(this_healthy, "T", "U")) %>%
            mutate(snp = F, pos = i)

        if(i == 1){
            full <- bind_rows(row1, row2)
        } else {
            full <- bind_rows(full, row1, row2)
        }
    }

    # Find average change in E score for the two groups of heptamers

    full2 <- full %>%
        group_by(pos) %>%
        filter(n() == 2) %>%
        ungroup() %>%
        pivot_longer(cols = contains("set")) %>%
        select(-`7mer`) %>%
        pivot_wider(names_from = "snp", values_from = "value") %>%
        mutate(abs_change = `TRUE`-`FALSE`) %>%
        filter(!is.na(`TRUE`) & !is.na(`FALSE`)) %>%
        group_by(name) %>%
        mutate(mean_abs_change = mean(abs_change)) %>%
        select(name, mean_abs_change) %>%
        unique() %>%
        arrange(mean_abs_change) %>%
        mutate(is_tdp43 = ifelse(str_detect(name, "RNCMPT00076"), "red", "grey80")) %>%
        mutate(short_name = str_sub(name, 1, 11)) %>%
        ungroup() %>% group_by(short_name) %>%
        mutate(average_delta = mean(mean_abs_change)) %>%
        select(short_name, is_tdp43, average_delta) %>%
        unique()

    # Plot ranks

    plot <- ggplot(full2, aes(x=rank(average_delta), y = average_delta,
                      fill = is_tdp43)) +
        geom_bar(stat="identity") +
        scale_fill_identity() +
        theme_minimal() +
        ggtitle(paste0("Average change in binding E value for each dataset with SNP for ", type, " SNP")) +
        xlab("") +
        ylab("Average change in E score") +
        ggpubr::theme_pubr() +
        xlab("Rank")

    print(plot)

    full_just_tdp <- full %>%
        group_by(pos) %>%
        filter(n() == 2) %>%
        ungroup() %>%
        select(`7mer`, contains("RNCMPT00076"), snp, pos) %>%
        mutate(mean_E = (RNCMPT00076_e_setA + RNCMPT00076_e_setB)/2) %>%
        pivot_longer(cols = contains("RNCMP")) %>%
        mutate(ypos = ifelse(snp, -0.35, -0.27))

    plot2 <- ggplot(full_just_tdp, aes(x = pos, y = value, colour = snp,
                              label = `7mer`)) +
        geom_point(size = 3) +
        xlab("7mer") +
        ylab("E value (similar to rank?)") +
        ggeasy::easy_add_legend_title("Contains Risk SNP") +
        ggtitle(paste0("Binding E value for TDP43 for all 7x 7mers with or without risk SNP for ", type, " SNP")) +
        geom_label(aes(y = ypos), size = 2.7) +
        xlim(0.5, 7.5) +
        theme_bw() +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  +
        ggeasy::easy_remove_x_axis()

    print(plot2)

}

```

# Isothermal Titration Calorimetry

Read in values

```{r warning=FALSE, message=FALSE}
df <- read_csv(paste0(here(),
                      "/data/itc/itc final values.csv")) %>%
    group_by(RNA) %>%
    mutate(sd =sd(kd)) %>%
    mutate(mean_kd = mean(kd))

ggplot(df, aes(x = RNA, y = kd)) +
    geom_bar(aes(x = RNA, y = mean_kd), stat="identity", position="dodge") +
    geom_point() +
    geom_errorbar(aes(x=RNA, ymin = mean_kd-sd, ymax = mean_kd+sd),
                  width = 0.3) +
    ylim(0, NA) +
    ggpubr::theme_pubr() +
    ylab("Kd / nM") +
    xlab("")

t.test(df$kd[which(df$RNA=="H")], df$kd[which(df$RNA=="R")])
```