### 2021-08-05 paper_figures.Rmd
Update paper_figures.Rmd
added plot for 2F and S2C

---
title: "Figures for Paper"
author: "Anna-Leigh Brown + Matteo Zanovello + Oscar Wilkins"
date: "06/10/2020"
output:
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'docs/index.html')) })
---

```{css, echo = FALSE}
h1, #TOC>ul>li {
  color: #000000;
    font-weight: bold;
}

h2, #TOC>ul>ul>li {
  color: #41494D;
}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
knitr::opts_chunk$set(
  comment = '', fig.width = 9, fig.height = 9
)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load("data.table",
       "tidyverse",
       "ggplot2",
       "rcompanion",
       "lme4",
       "ggpubr",
       "forcats",
       "scales",
       "here",
       "ggthemes",
       "ggrepel",
       "DESeq2",
       "patchwork",
       "rstatix",
       "ggeasy")
```

# iPSC splicing

Read in the splicing results returned by MAJIQ and make a volcano plot, only highlight
genes of interest with a label.

```{r volcano_plots}

ipsc_splicing = fread(file.path(here::here(),"data","ipsc_splicing_results.csv"))
ipsc_de = fread(file.path(here::here(),"data","ipsc_differential_expression.csv"))

splicing_dots_tables <- ipsc_splicing %>%
    mutate(junction_name = case_when(gene_name %in% c("UNC13A","AGRN",
                                                      "UNC13B","PFKP","SETD5",
                                                      "ATG4B","STMN2") &
                                         p_d_psi_0_10_per_lsv_junction > 0.9  &
                                         deltaPSI > 0 ~ gene_name,
                                     T ~ "")) %>%
    mutate(`Novel Junction` = de_novo_junctions == 0) %>%
    mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>%
    mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>%
    mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>%
    mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                       "UNC13B","PFKP","SETD5",
                                                       "ATG4B","STMN2") &
                                          p_d_psi_0_10_per_lsv_junction > 0.9  &
                                          deltaPSI > 0 ~ junction_name,
                                      T ~ ""))

fig1a = ggplot() +
    geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions != 0),
               aes(x = deltaPSI, y =log10_test_stat,
                   alpha = graph_alpha,,fill = "Annotated Junction"), pch = 21, size = 10) +
    geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions == 0),
               aes(x = deltaPSI, y =log10_test_stat,
                   alpha = graph_alpha,fill = "Novel Junction"), pch = 21, size = 10) +
    geom_text_repel(data = splicing_dots_tables[junction_name != ""],
                    aes(x = deltaPSI, y =log10_test_stat,
                        label = label_junction,
                        color = as.character(de_novo_junctions)), point.padding = 0.3,
                    nudge_y = 0.2,
                    min.segment.length = 0,
                    box.padding  = 2,
                    size=6,show.legend = F) +
    geom_hline(yintercept = -log10(1 - .9)) +
    geom_vline(xintercept = -0,linetype="dotted") +
    scale_fill_manual(name = "",
                      breaks = c("Annotated Junction", "Novel Junction"),
                      values = c("Annotated Junction" = "#648FFF", "Novel Junction" = "#fe6101") ) +
    scale_color_manual(name = "",
                       breaks = c("0", "1"),
                       values = c("1" = "#648FFF", "0" = "#fe6101") ) +
    guides(alpha = FALSE, size = FALSE) +
    theme(legend.position = 'top') +
    ggpubr::theme_pubr() +
    xlab("delta PSI") +
    ylab(expression(paste("-Lo", g[10], " Test Statistic"))) +
    theme(text = element_text(size = 24)) +
    theme(legend.text=element_text(size=22)) +
    xlim(-1,1) +
    scale_x_continuous(labels = scales::percent)

de_table = ipsc_de %>%
    mutate(contains_cryptic = gene_name %in% splicing_dots_tables[cryptic_junction == T,unique(gene_name)]) %>%
    mutate(contains_cryptic = as.character(as.numeric(contains_cryptic))) %>%
    mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                       "UNC13B","PFKP","SETD5",
                                                       "ATG4B","STMN2","TARDBP") ~ gene_name,
                                      T ~ "")) %>%
    mutate(graph_alpha = ifelse(padj < 0.1, 1, 0.2)) %>%
    mutate(y_data = -log10(padj))

fig1b = ggplot(data = de_table) +
    geom_point(data = de_table %>% filter(contains_cryptic == "0"),
               aes(x = log2FoldChange, y = -log10(padj),
                   alpha = graph_alpha,fill = "No Cryptic"), pch = 21, size = 10) +
    geom_point(data = de_table %>% filter(contains_cryptic == "1"),
               aes(x = log2FoldChange, y = -log10(padj),
                   alpha = graph_alpha,fill = "Contains Cryptic"), pch = 21, size = 10)+
    geom_text_repel(data = de_table[label_junction != ""],max.overlaps = 500,
                    aes(x = log2FoldChange,
                        y = -log10(padj),
                        label = label_junction,
                        color = as.character(contains_cryptic)),
                    nudge_y = 5,
                    min.segment.length = 0,
                    box.padding  = 4,
                    size=6,show.legend = F) +
    scale_fill_manual(name = "",
                      breaks = c("No Cryptic", "Contains Cryptic"),
                      values = c("No Cryptic" = "#648FFF", "Contains Cryptic" = "#fe6101") ) +
    scale_color_manual(name = "",
                       breaks = c("1", "0"),
                       values = c("1" = "#fe6101", "0" = "#648FFF") ) +
    xlim(-7.5,7.5) +
    ylab(expression(paste("-Lo", g[10], " P-value"))) +
    guides(alpha = FALSE, size = FALSE) +
    theme(legend.position = 'top') +
    ggpubr::theme_pubr() +
    xlab(expression(paste("Lo", g[2], " Fold Change"))) +
    theme(text = element_text(size = 24)) +
    theme(legend.text=element_text(size=22)) +
    geom_hline(yintercept = -log10(0.1))  +
    geom_vline(xintercept=c(0), linetype="dotted")

print(fig1a)
print(fig1b)
```

# UNC13A CE PSI across TDP-43 KD experiments

The 'C/G' tells which genotypes were supported by RNA-seq on rs12973192.
The SK-N-DZ cell lines are het, as is the WTC11 cell line. SH-SY5Y cells are homozygote for the major allele.
There was variability on the Klim hMN set on allelic expression.

```{r read in }

rel_rna_cryptic_amount = data.table::fread(file.path(here::here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime +
                                                  UNC13A_5prime +  UNC13A_5prime_2 +
                                                  UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime +
                                                                          UNC13A_5prime +  UNC13A_5prime_2 +
                                                                          UNC13A_5prime_3)]
```

## barplot UNC13A CE PSI - full five

```{r}
rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
                      values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(name = "",
                       values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A CE PSI") +
    xlab("") +
    guides(color = FALSE) +
    scale_y_continuous(labels = scales::percent) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))
```

## barplot UNC13B frameshift PSI

```{r}
rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "condition",
              add = c("mean_se","jitter"),
              y = "unc13b_nmd_exon_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8),facet.by = 'source') +
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
                      values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(name = "",
                       values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B \nNMD Exon PSI") +
    xlab("") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    facet_wrap(~source) +
    stat_compare_means() +
    scale_y_continuous(labels = scales::percent)
```

## TARDBP RNA and UNC13A Cryptic

```{r}

rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12,method = 's',cor.coef.name = 'rho') +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
    scale_y_continuous(labels = scales::percent)
```

## UNC13A RNA and UNC13A Cryptic

```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) +
    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12,
             method = 'spearman',
             cor.coef.name = 'rho') +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    scale_y_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
    theme(text = element_text(size = 18,family = 'sans'),
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
```
## UNC13A RNA and UNC13A IR
```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12,cor.coef.name = 'rho',method = 's') +
    geom_smooth(aes(x = UNC13A, y = normalized_unc13a_ir),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom')
```
## TARDBP RNA and UNC13A IR

```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = TARDBP, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = TARDBP, y = normalized_unc13a_ir),size = 12) +
    geom_smooth(aes(x = TARDBP, y = normalized_unc13a_ir),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom')
```

## barplot UNC13A cryptic PSI

```{r}
rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A CE PSI") +
    xlab("") +
    guides(color = FALSE)
```

## barplot STMN2 cryptic PSI

```{r}
stmn2 = rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "stmn_2_cryptic_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("STMN2 Cryptic PSI") +
    xlab("") +
    guides(color = FALSE)
print(stmn2)
```

## barplot UNC13B normalized IR
```{r}
unc13b_ir = rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13b_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13B \nIntron Retention Ratio") +
    xlab("") +
    guides(color = FALSE) +
    theme(legend.position = "none") +
    theme(text = element_text(size = 18)) +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

print(unc13b_ir)
```

## barplot UNC13A normalized IR
```{r}
unc13a_ir = rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13a_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13A \nIntron Retention Ratio") +
    xlab("") +
    theme(legend.position = "none") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

```

## barplot UNC13A normalized RNA Level

```{r}
text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))

rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "UNC13A",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A") +
    xlab("") +
    theme(legend.position = "none") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent) +
    geom_text(data = text_table[gene_name == "UNC13A"],aes(x = source, y= 1.35,label = padj_plot),size = 8) +
    geom_text(data = text_table[gene_name == "UNC13A"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

```

## barplot UNC13B normalized RNA Level

```{r}
text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))

rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "UNC13A",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B") +
    xlab("") +
    theme(legend.position = "none") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent) +
    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.35,label = padj_plot),size = 8) +
    geom_text(data = text_table[gene_name == "UNC13B"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

```

## barplot TARDBP normalized RNA Level

```{r}
text_table = fread(file.path(here::here(),"data","deseq2_cellline_fold_change.csv"))

rel_rna_cryptic_amount %>%
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "TARDBP",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("TARDBP") +
    xlab("") +
    theme(legend.position = "none") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'),
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28)) +
    scale_y_continuous(labels = scales::percent,expand = expansion(mult = c(0, .1))) +
    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.35,label = padj_plot),size = 8) +
    geom_text(data = text_table[gene_name == "TARDBP"],aes(x = source, y= 1.2,label = log2FoldChange_plot),size = 8)

```
## Scatterplot stmn2 normalized TARDBP
```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = TARDBP, y = stmn_2_cryptic_psi, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = TARDBP, y = stmn_2_cryptic_psi)) +
    theme(legend.position = 'bottom')
```

## Scatterplot UNC13BIR normalized TARDBP

```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = TARDBP, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = TARDBP, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')
```

## Scatterplot UNC13A-IR and UNC13A Crptic
```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = cryptic_psi_full,
                   y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = cryptic_psi_full,
                 y = normalized_unc13a_ir)) +
    theme(legend.position = 'bottom')
```

## Scatterplot UNC13B-IR and UNC13A Cryptic

```{r}

rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+
    facet_wrap(~source)
```

## scatterplot UNC13BIR and UNC13B NMD

```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+
    facet_wrap(~source)

```

## scatterplot UNC13A CE and UNC13B NMD

```{r}
rel_rna_cryptic_amount %>%
    filter(condition != "control") %>%
    ggplot() +
    geom_point(aes(x = cryptic_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) +
    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+
    facet_wrap(~source)

```
# NMD experiment on SH-SY5Y cells - CHX

```{r}

for_bar_plot_input <- fread(file.path(here::here(),"data","SY5Y_Tidy_4.csv"))
for_bar_plot <- data.frame(for_bar_plot_input)
for_bar_plot$Treatment <- factor(for_bar_plot$Treatment, levels = c("DMSO", "CHX"))

shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "HNRNPL")$Value)
shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "STMN2")$Value)
shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "UNC13A")$Value)
shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "UNC13B")$Value)

stattest <- for_bar_plot %>%
  group_by(Gene) %>%
  pairwise_t_test(Value ~ Treatment, ref.group = "DMSO", alternative = "greater") %>%
  add_significance() %>% 
  add_xy_position(x="Treatment")
stattest

nmd_plot_chx = ggbarplot(for_bar_plot,
                     x = 'Gene',
                     add = c("mean_se","jitter"),
                     y = 'Value',
                     color = 'Treatment',
                     position = position_dodge(0.8),
                     dot.size = 10) +
  scale_y_continuous() + 
  ggpubr::theme_pubr() + stat_pvalue_manual(stattest, label="p.signif" , x = "Gene")
plot(nmd_plot_chx)

```

# NMD experiment on SH-SY5Y cells - sTDP43+sUPF1

```{r}

for_bar_plot_sy_input <- fread(file.path(here::here(),"data","results_upf1.csv"))

for_bar_plot_sy_df <- data.frame(for_bar_plot_sy_input)
for_bar_plot_sy_df$Condition <- as.factor(for_bar_plot_sy_df$Condition)

#assess TDP-43 and UPF1 KD
for_bar_plot_sz <- for_bar_plot_sy_df %>%
  filter(Gene == "TDP43" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1" | Condition == "Non-treated"))

shapiro.test(filter(for_bar_plot_sz, Condition == "siTDP43+siCTRL" & Gene == "TDP43")$Value)
shapiro.test(filter(for_bar_plot_sz, Condition == "siTDP43+siUPF1" & Gene == "TDP43")$Value)

stattest_sz <- for_bar_plot_sz %>%
  pairwise_t_test(Value ~ Condition, ref.group = "Non-treated", alternative = "less") %>%
  add_significance() %>% 
  add_xy_position(x="Condition")
stattest_sz

for_bar_plot_sx <- for_bar_plot_sy_df %>%
  filter(Gene == "UPF1" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1"))

shapiro.test(filter(for_bar_plot_sx, Condition == "siTDP43+siUPF1" & Gene == "UPF1")$Value)

stattest_sx <- for_bar_plot_sx %>%
  pairwise_t_test(Value ~ Condition, ref.group = "siTDP43+siCTRL", alternative = "less") %>%
  add_significance() %>% 
  add_xy_position(x="Condition")
stattest_sx

for_bar_plot_szx <- for_bar_plot_sy_df %>%
  filter(Gene == "TDP43" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1" | Condition == "Non-treated")
         | Gene == "UPF1" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1"))

#fix position of stats for plot
stattest_sx$xmin[1] <- 1.8
stattest_sx$xmax[1] <- 2.2
stattest_sx$y.position[1] <- 1.3
stattest_sz$xmin[1] <- 0.75
stattest_sz$xmin[2] <- 0.75
stattest_sz$xmax[1] <- 1
stattest_sz$xmax[2] <- 1.25
stattest_sz$y.position[1] <- 1.2
stattest_sz$y.position[2] <- 1.3

nmd_plotz = ggbarplot(for_bar_plot_szx,
                     x = 'Gene',
                     add = c("mean_se","jitter"),
                     y = 'Value',
                     color = 'Condition',
                     position = position_dodge(0.8),
                     dot.size = 10) +
  scale_y_continuous() + ylab("Percent of transcript after UPF1 knockdown") + xlab("Gene") + 
  ggpubr::theme_pubr() + stat_pvalue_manual(stattest_sz) + stat_pvalue_manual(stattest_sx)
plot(nmd_plotz)

#assess rescue of NMD-sensitive transcripts after UPF1 inhibition
for_bar_plot_sy <- for_bar_plot_sy_df %>%
  filter(Gene == "hnRNPL" | Gene == "STMN2" | Gene == "UNC13A" | Gene == "UNC13B")

shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "hnRNPL")$Value)
shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "STMN2")$Value)
shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "UNC13A")$Value)
shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "UNC13B")$Value)

stattest_sy <- for_bar_plot_sy %>%
  group_by(Gene) %>%
  pairwise_t_test(Value ~ Condition, ref.group = "siTDP43+siCTRL", alternative = "greater") %>%
  add_significance() %>% 
  add_xy_position(x="Condition")
stattest_sy

nmd_plot = ggbarplot(for_bar_plot_sy, 
                     x = 'Gene', 
                     add = c("mean_se","jitter"), 
                     y = 'Value',
                     color = 'Condition', 
                     position = position_dodge(0.8), 
                     dot.size = 10) +
  scale_y_continuous() + 
  ggpubr::theme_pubr() + ylab("Percent of transcript after UPF1 knockdown") + xlab("Gene") +
  stat_pvalue_manual(stattest_sy, label="p.signif", x = "Gene")
plot(nmd_plot)

```

# iPSC protein abundance mass spec

```{r}
peptides = fread(file.path(here::here(),"data","peptide_abdundance.csv"))

peptides %>%
    mutate(gene = fct_relevel(gene, "UNC13A","UNC13B")) %>%
    ggbarplot(,
              x = "condition",
              add = c("mean_se"),
              y = "protein",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8),
              facet.by = 'gene') +
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) +
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Protein abundance") +
    xlab("") +
    geom_jitter(pch = 21,height = 0,aes(fill = condition),size = 2) +
    guides(color = FALSE) +
    facet_wrap(~gene, scales = 'free') +
    theme(legend.position = 'none') +
    scale_y_continuous(labels = scientific) +
    stat_compare_means(comparisons = list(c("Control","TDP-43 KD")),
                       label = 'p.signif',tip.length = 0,
                       size = 10)

```

# Patient summary statistics

```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%
    mutate(rs12973192 = fct_relevel(rs12973192,
                                    "C/C", "C/G", "G/G")) %>%
    mutate(number_g_alleles = as.numeric(rs12973192) - 1) %>%
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>%
    mutate(junction_reads_stmn2 = STMN2_annotated_leaf + STMN2_cryptic_leaf) %>%
    mutate(junction_reads_unc13a = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf + UNC13A_annotated_leaf) %>%
    unique()

print(glue::glue("Number of unique patients: {clean_data_table[,length(unique(participant_id))]}"))
print(glue::glue("Number of unique tissue samples: {clean_data_table[,length(unique(sample))]}"))
print("Patients Per Disease Category")
clean_data_table[,length(unique(participant_id)),by = disease]
print("Tissues Per Disease Category")
clean_data_table[,length(unique(sample)),by = disease]
print("Number of patients per UNC13A SNP genotype")
unique(clean_data_table[,.(participant_id,rs12608932,rs12973192)]) %>% select(-participant_id) %>% table()
print("Number of tissues per disease")
clean_data_table[,.N,by = c("disease","tissue_clean")]
print("Number of partcipants by mutation and  disease")
clean_data_table[,length(unique(participant_id)),by = c("mutations","disease")]

print(glue::glue("Number of patients per pathology:"))
clean_data_table[,length(unique(participant_id)),by = .(pathology)]
```
# UNC13A cryptic is an event that is specific to TDP-43 proteinopathy

FTLD-non-TDP are those with TAU and FUS aggregates

Non-tdp ALS are those with SOD1 or FUS mutations. The n's are quite low on this
unfortunately, only 8 ALS with SOD1 and 2 with FUS mutations.

First we look at detection rate in tissues affected by TDP-43 proteinopathy,
For FTLD this is frontal and temporal Cortices, and
for ALS this is lumbar, cervical, and thoracic spinal cord samples as well as
motor cortex. For controls we also take all 6 tissues,
frontal,temporal,motor cortices and the lumbar, cervical, and thoracic spinal cords.

(As a side note, once we do this the number of ALS-non-TDP drops down to 6 (2 FUS) because
the ALS sample tissues are not balanced and not every participant has samples in every tissue)

```{r Inclusion reads by if TDP-potential}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>%
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>%
    filter(!tissue_clean %in% c("Choroid","Liver")) %>%
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>%
    melt() %>%
    filter(grepl("_3prime|_5prime_1",variable)) %>%
    group_by(sample) %>%
    mutate(inclusion_reads = sum(value)) %>%
    ungroup() %>%
    unique() %>%
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor",
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor",
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>%
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>%
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```
## Pecent of patients UNC13A detected
Looking at disease tissue only, so just taking the cord and motor cortex in ALS and the
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.

```{r}

clean_data_table %>%
  filter(disease_tissue == T) %>%
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(detected = inclusion_reads > 0) %>%
  dplyr::select(participant_id,disease_group2,detected) %>%
  unique() %>% 
  group_by(disease_group2) %>%
  mutate(n_sample = n_distinct(participant_id)) %>%
  mutate(n_sample_detected = sum(detected)) %>%
  dplyr::select(disease_group2,n_sample,n_sample_detected) %>%
  unique() %>%
  mutate(detection_rate = n_sample_detected / n_sample) %>%
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>%
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>%
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) +
  ggpubr::theme_pubr() +
  scale_y_continuous(lim = c(0,1),labels = scales::percent) +
  ylab("Percent of Patients \n UNC13A Cryptic Detected") +
  theme(text = element_text(size = 24)) +
  xlab("N individuals")
```

## Read length affects detection probability in cases

```{r readlength_detection_in_cases}
# library size 
#llumina HiSeq 2500 (125 bp paired end) or an Illumina NovaSeq (100 bp paired end). 
case_control_sequencing <- clean_data_table %>%
  filter(tissue_clean %in% c("Frontal_Cortex", 
                             "Lumbar_Spinal_Cord", 
                             "Cervical_Spinal_Cord", 
                             "Motor_Cortex", 
                             "Temporal_Cortex",
                             "Cerebellum")) %>% 
  filter(sequencing_platform != "")

quick_chiq = function(tbl){

    if(length(unique(tbl$sequencing_platform)) < 2){
      return(NA)
    }
    comparison_table = table(disease = tbl$disease_group2, 
                             sequencing_platform = tbl$sequencing_platform)
    print(comparison_table)

    p_value = chisq.test(comparison_table) %>%  broom::tidy() %>% pull(p.value)

    print(p_value)

    return(p_value)
}

case_control_sequencing %>% 
  group_by(tissue_clean) %>% 
  nest()  %>% 
  mutate(chiq_pvalue = map_dbl(data, quick_chiq))

unc13a_data_fisher$data = NULL

unc13a_sequencing_platform_pct = unc13a_sequencing_platform %>%
  group_by(disease_group2,tissue_clean, sequencing_platform, unc13a_detected) %>% tally() %>%
  spread(key = unc13a_detected, value = n,fill = 0) %>%
  mutate(detection = `+` / ( `-` + `+`) ) %>% 
  mutate(total =  ( `-` + `+`) ) %>% 
  mutate(plot_name = glue::glue("{sequencing_platform} \n  ({total})")) %>% 
  left_join(unc13a_data_fisher)

unc13a_sequencing_platform_pct %>%
  ggplot(aes(x = plot_name, y = detection )) + 
  geom_col(fill = "firebrick") +
  labs(title = "UNC13A CE detection and sequencing platform", 
       y = "UNC13A CE detected",x = element_blank(), subtitle = "Fisher exact test") + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free_x') +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_text(aes(x = 1.5, y = 1.05, label = paste0("p = ", signif(fisher_pvalue,digits = 2) ) )) +
  theme(legend.text = element_text(size = 15)) +
  ggpubr::theme_pubr()
 # scale_y_continuous(expand = c(0,0) ) 
```

## Effect of sequencing machine (read length) on CE discovery

```{r unc detection by sequencer  }
# library size 
#llumina HiSeq 2500 (125 bp paired end) or an Illumina NovaSeq (100 bp paired end). 
unc13a_sequencing_platform <- clean_data_table %>%
  filter(tissue_clean %in% c("Frontal_Cortex", 
                             "Lumbar_Spinal_Cord", 
                             "Cervical_Spinal_Cord", 
                             "Motor_Cortex", 
                             "Temporal_Cortex")) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
  mutate(tissue = as.character(tissue)) %>%
  filter(sequencing_platform != "") %>% 
  filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP"))

quick_fisher = function(tbl){

    if(length(unique(tbl$sequencing_platform)) < 2){
      return(NA)
    }
    comparison_table = table(unc13a = tbl$unc13a_detected, 
                             sequencing_platform = tbl$sequencing_platform)
    p_value = fisher.test(comparison_table)$p.value

    return(p_value)
}

unc13a_data_fisher = unc13a_sequencing_platform %>% 
  filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
  group_by(disease_group2,tissue_clean) %>% 
  nest() %>% 
  mutate(fisher_pvalue = map_dbl(data, quick_fisher))

unc13a_data_fisher$data = NULL

unc13a_sequencing_platform_pct = unc13a_sequencing_platform %>%
  group_by(disease_group2,tissue_clean, sequencing_platform, unc13a_detected) %>% tally() %>%
  spread(key = unc13a_detected, value = n,fill = 0) %>%
  mutate(detection = `+` / ( `-` + `+`) ) %>% 
  mutate(total =  ( `-` + `+`) ) %>% 
  mutate(plot_name = glue::glue("{sequencing_platform} \n  ({total})")) %>% 
  left_join(unc13a_data_fisher)

unc13a_sequencing_platform_pct %>%
  ggplot(aes(x = plot_name, y = detection )) + 
  geom_col(fill = "firebrick") +
  labs(title = "UNC13A CE detection and sequencing platform", 
       y = "UNC13A CE detected",x = element_blank(), subtitle = "Fisher exact test") + 
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free_x') +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_text(aes(x = 1.5, y = 1.05, label = paste0("p = ", signif(fisher_pvalue,digits = 2) ) )) +
  theme(legend.text = element_text(size = 15)) +
  ggpubr::theme_pubr()
 # scale_y_continuous(expand = c(0,0) ) 
```
## Effect of library size on CE discovery

```{r unc detection by library size  }
# library size 
force_colors = c("#F8766D", "#A3A500", "#00BF7D", "#00B0F6", "#E76BF3")
names(force_colors) = c("Control","ALS \n non-TDP","ALS-TDP","FTLD \n non-TDP","FTLD-TDP")

plot_list = list()

lib_comps = c("Cervical_Spinal_Cord", "Frontal_Cortex", 
              "Lumbar_Spinal_Cord", "Temporal_Cortex",
              "Motor_Cortex", "Cerebellum")

plot_list = list()
for(t in 1:length(lib_comps)){
  plt = clean_data_table %>%
     filter(tissue_clean == lib_comps[t]) %>% 
    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
    mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
    ggplot(aes(x = disease_group2,y = log10(library_size),fill = disease_group2)) + 
    geom_boxplot(colour = "black", outlier.colour = NA) +
    geom_point(pch = 21, size = 2, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
    facet_wrap(~ tissue_clean,scales = 'free_x') + 
    # ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
    ggpubr::stat_compare_means(label = "p.signif",
                               label.x.npc = "center",
                               method = "wilcox",
                               comparisons = list(
                                                  c("Control","ALS-TDP"),
                                                  c("ALS \n non-TDP","ALS-TDP"))) +
      ggpubr::stat_compare_means(label = "p.signif", 
                               label.x.npc = "center", 
                               method = "wilcox",
                               comparisons = list(
                                                  c("Control","FTLD-TDP"),
                                                  c("FTLD-TDP","FTLD \n non-TDP"))) + 
    labs(y = "log10 library size", fill= "UNC13A CE detected", x = element_blank()) +
    theme(axis.text.x = element_text(size = 15)) + 
    ggpubr::theme_pubr() + 
    scale_fill_manual(values = force_colors) +
            theme(axis.ticks.x = element_blank(),
              axis.text.x = element_blank(),
              legend.position = "none") + 
    ylab(element_blank())

  plot_list[[t]] <- plt
}

cowplot::plot_grid(plotlist = plot_list,nrow = 3)
```

## Effect of RIN and UNC13A discovery

```{r rin discovery}
force_colors = c("#F8766D", "#A3A500", "#00BF7D", "#00B0F6", "#E76BF3")
names(force_colors) = c("Control","ALS \n non-TDP","ALS-TDP","FTLD \n non-TDP","FTLD-TDP")

plot_list = list()

lib_comps = c("Cervical_Spinal_Cord", "Frontal_Cortex", 
              "Lumbar_Spinal_Cord", "Temporal_Cortex",
              "Motor_Cortex", "Cerebellum")

plot_list = list()
for(t in 1:length(lib_comps)){

        groups = c("ALS-TDP", "Control", "ALS \n non-TDP", "FTLD-TDP", "FTLD \n non-TDP")

    plt = clean_data_table %>%
        filter(tissue_clean == lib_comps[t]) %>% 
        mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
        filter(disease_group2 %in% groups) %>%

        ggplot(aes(x = disease_group2,y = rin,fill =disease_group2 )) + 
        geom_boxplot(colour = "black", outlier.colour = NA) +
        geom_point(pch = 21, size = 2, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
        facet_wrap(~ tissue_clean,scales = 'free_x') + 
        # ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center", method = "wilcox" ) +
        ggpubr::stat_compare_means(label = "p.signif",
                                   label.x.npc = "center",
                                   method = "wilcox",
                                   comparisons = list(
                                       c("Control","ALS-TDP"),
                                       c("ALS \n non-TDP","ALS-TDP"))) +
        ggpubr::stat_compare_means( label = "p.signif",
                                   label.x.npc = "center", 
                                   method = "wilcox",
                                   comparisons = list(
                                       c("Control","FTLD-TDP"),
                                       c("FTLD-TDP","FTLD \n non-TDP"))) + 
        labs(y = "RIN", x = element_blank()) +
        theme(axis.text.x = element_text(size = 15)) + 
        ggpubr::theme_pubr() + 
        theme(axis.ticks.x = element_blank(),
              axis.text.x = element_blank(),
              legend.position = "none") +
      scale_color_manual(values = force_colors) + 
      scale_fill_manual(values = force_colors)

    plot_list[[t]] <- plt
}

cowplot::plot_grid(plotlist = plot_list,ncol = 2)    
```

## Effect of UNC13A discovery and UNC13B TPM

```{r unc13a ce and UNC13B_TPM}
## effect of RIN?
clean_data_table %>%
  filter(disease_tissue == T) %>% 
  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  add_count(tissue_clean,disease_group2) %>% 
  mutate(plot_name = glue::glue("{disease_group2} \n ({n})")) %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-") ) %>% 
  ggplot(aes(x = plot_name, fill = unc13a_detected,y = UNC13B_TPM )) + 
  geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
  geom_boxplot(outlier.colour = NA) +
  labs(title = "UNC13A CE detection and UNC13B TPM",subtitle = "Wilcox test") + 
  facet_wrap(~ tissue_clean,scales = 'free') + 
  ggpubr::stat_compare_means(label = "p.signif") +
  labs(y = "UNC13B TPM", x = "UNC13A CE detection") +
  theme(axis.text.x = element_text(size = 15)) +
  ggpubr::theme_pubr() +
  ggpubr::stat_compare_means(label = "p.format", label.x.npc = "center",label.y.npc = 0.9)
```

## Celltype composition between cases and controls

```{r}
force_colors = c("#F8766D", "#A3A500", "#00BF7D", "#00B0F6", "#E76BF3")
names(force_colors) = c("Control","ALS \n non-TDP","ALS-TDP","FTLD \n non-TDP","FTLD-TDP")

plot_list = list()

lib_comps = c("Cervical_Spinal_Cord", "Frontal_Cortex", 
              "Lumbar_Spinal_Cord", "Temporal_Cortex",
              "Motor_Cortex", "Cerebellum")

cell_types = fread(paste0(here::here(),"/data/All_1917_samples_Darmanis_dtangle_deconv.tsv"))
for(t in 1:length(lib_comps)){

        groups = c("ALS-TDP", "Control", "ALS \n non-TDP", "FTLD-TDP", "FTLD \n non-TDP")

    plt = cell_types %>% 
        left_join(clean_data_table[,.(sample,
                                      UNC13A_3prime_leaf,
                                      UNC13A_5prime_1_leaf,
                                      disease_group2,
                                      disease_tissue)], by = 'sample') %>%
        filter(!is.na(disease_group2)) %>% 
        unique() %>% 
        filter(tissue_clean == lib_comps[t]) %>% 
        filter(disease_group2 %in% groups) %>% 
        mutate(disease_group2 = fct_relevel(disease_group2, "Control")) %>% 
        ggplot(aes(x = disease_group2,y = deconv,fill =disease_group2 )) + 
        geom_boxplot(colour = "black", outlier.colour = NA) +
        facet_wrap(~cell,scales = 'free_x',nrow = 1) +
        ggpubr::stat_compare_means(label = "p.signif", 
                                   label.x.npc = "center", 
                                   method = "wilcox",
                                   comparisons = list(
                                       c("Control","FTLD-TDP"),
                                       c("FTLD-TDP","FTLD \n non-TDP"))) +
        ggpubr::stat_compare_means(label = "p.signif",
                                   label.x.npc = "center",
                                   method = "wilcox",
                                   comparisons = list(
                                       c("Control","ALS-TDP"),
                                       c("ALS \n non-TDP","ALS-TDP"))) +
        ggtitle(lib_comps[t]) + 
        labs(fill = element_blank(),x = element_blank()) + 
        geom_point(pch = 21, size = 2, alpha = 0.9, position = position_jitter(width = 0.2, height = 0) ) + 
        ggpubr::theme_pubr() +
        theme(axis.ticks.x = element_blank(),
              axis.text.x = element_blank(),
              legend.position = "none") + 
        scale_fill_manual(values = force_colors)
plot_list[[t]] <- plt

}

cowplot::plot_grid(plotlist = plot_list,ncol = 2)

```

## Effect of celltype composition on UNC13A CE detection

```{r celltype and detection}

cell_types = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/All_1917_samples_Darmanis_dtangle_deconv.tsv")

cell_types %>% 
  left_join(clean_data_table[,.(sample,
                               UNC13A_3prime_leaf,
                               UNC13A_5prime_1_leaf,
                               disease_group2,
                               disease_tissue)], by = 'sample') %>%
  filter(disease_group2 %in% c("FTLD-TDP", "ALS-TDP")) %>%
  filter(disease_tissue == T) %>% 
  filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
  mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
  mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
  add_count(disease_group2,tissue_clean) %>% 
  mutate(n = n / 5) %>% 
  mutate(plot_name = glue::glue("{disease_group2} \n {n}")) %>% 
  filter(!is.na(inclusion_reads)) %>% 
  unique() %>% 
  ggplot(aes(x = cell, y = deconv, fill = unc13a_detected)) + 
  geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
  geom_boxplot(outlier.color = NA) + 
  facet_wrap(~plot_name+tissue_clean,scales = 'free') +
  ggpubr::stat_compare_means(label = "p.signif") +
  ggpubr::stat_compare_means(label = "p.format",vjust = 1.2) +
  labs(title = "Cell type decomposition and UNC13A CE detection ",subtitle = "Wilcox test") + 
  ggpubr::theme_pubr()

```

## Pecent of samples UNC13A detected by tissue

```{r}
boxplot_table %>%
  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>%
  mutate(detected = inclusion_reads > 0) %>%
  dplyr::select(sample,disease_group2,detected,tissue_clean) %>%
  unique() %>%
  group_by(disease_group2,tissue_clean) %>%
  mutate(n_sample = n_distinct(sample)) %>%
  mutate(n_sample_detected = sum(detected)) %>%
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>%
  unique() %>%
  mutate(detection_rate = n_sample_detected / n_sample) %>%
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>%
  ungroup() %>%
  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>%
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Motor Cortex",
                                    "Thoracic Spinal Cord")) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>%
  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>%
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate,fill = detection_name),position = position_dodge2(width = 0.8, preserve = "single")) +
  ggpubr::theme_pubr() +
  ylab("Percent of Tissues \n UNC13A Cryptic Detected") +
  theme(text = element_text(size = 18)) +
  xlab("N tissues") +
  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) +
  theme(axis.text.x =  element_text(size = 14)) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size = 18))  +
  scale_fill_manual(values = c("#4C97B5","#4C97B5","#4C97B5","#4C97B5","#4C97B5","red","#C87156","#C87156")) +
  theme(legend.position = "none")
```

## Pecent of samples UNC13A detected by tissue and genotype

```{r}

boxplot_table %>%
  left_join(clean_data_table %>% dplyr::select(sample, rs12973192)) %>%
  unique() %>%
  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>%
  mutate(detected = inclusion_reads > 0) %>%
  dplyr::select(sample,disease_group2,detected,tissue_clean,rs12973192) %>%
  unique() %>%
  group_by(disease_group2,tissue_clean,rs12973192) %>%
  mutate(n_sample = n_distinct(sample)) %>%
  mutate(n_sample_detected = sum(detected)) %>%
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>%
  unique() %>%
  mutate(detection_rate = n_sample_detected / n_sample) %>%
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>%
  ungroup() %>%
  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>%
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Motor Cortex",
                                    "Thoracic Spinal Cord")) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>%
  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>%
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate,fill = rs12973192),position = position_dodge2(width = 0.8, preserve = "single")) +
  ggpubr::theme_pubr() +
  ylab("Percent of Tissues \n UNC13A Cryptic Detected") +
  theme(text = element_text(size = 18)) +
  xlab("N tissues") +
  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) +
  theme(axis.text.x =  element_text(size = 14)) +
      theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size = 18))  +
  theme(legend.position = "bottom")
```

## UNC13A CE level across tissue

This only plots the short novel acceptor and novel donor

```{r onlyshort}

star_maker = function (p.value) 
{
    unclass(symnum(p.value, corr = FALSE, na = FALSE, cutpoints = c(0, 0.0001,
        0.001, 0.01, 0.05, 0.1, 1), symbols = c("****","***", "**", 
        "*", ".", " ")))
}
only_short_plotter = boxplot_table %>%
    filter(tissue_clean %in%
               c("Frontal Cortex",
                 "Lumbar Spinal Cord",
                 "Cervical Spinal Cord",
                 "Motor Cortex",
                 "Temporal Cortex",
                 "Cerebellum") )%>%
    group_by(disease_group2,tissue_clean) %>%
    mutate(n_sample = n_distinct(sample)) %>%
    ungroup() %>%
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                      "Frontal Cortex",
                                      "Lumbar Spinal Cord",
                                      "Temporal Cortex",
                                      "Motor Cortex"))

only_short_plotter %>% 
  dplyr::select(sample,tissue_clean,inclusion_reads,disease_group2) %>% 
  mutate(disease_group2 = gsub("\n", "",disease_group2)) %>% 
  unique() %>% 
  group_by(tissue_clean) %>% 
  nest() %>%
  mutate(wilcox_res = map(data, ~pairwise.wilcox.test(.x$inclusion_reads,.x$disease_group2) %>% broom::tidy())) %>%
  dplyr::select(-data) %>% 
  unnest() %>% 
  mutate(star_plot = star_maker(p.value)) %>% fwrite("~/Desktop/stars_for_weaverly_fig3b.csv")

only_short_plotter %>% 
    ggplot(aes(x = detection_name,
               y = inclusion_reads * 10^6,
               fill ="red")) +
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single",width = 1),outlier.colour = NA) +
    geom_point(size = 2,show.legend = F,pch = 21, position = position_jitterdodge(dodge.width=1,jitter.width = 0.3))+
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) +
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) +
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(text = element_text(size = 18))  +
    theme(legend.position = "bottom") + 
    stat_compare_means(comparisons = comps)
```

## UNC13A CE level across tissue - split by junction type

```{r}
clean_data_table %>%
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>%
    filter(!tissue_clean %in% c("Choroid","Liver")) %>%
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>%
    melt() %>%
    filter(grepl("_3prime|_5prime_1|_5prime_2",variable)) %>%
      # filter(grepl("_5prime_2",variable)) %>%
    unique() %>%
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor",
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor",
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>%
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  dplyr::select(sample,disease_group2,
                tissue_clean,
                junction_name,
                value) %>%
    unique() %>%
    filter(tissue_clean %in%
               c("Frontal Cortex",
                 "Lumbar Spinal Cord",
                 "Cervical Spinal Cord",
                 "Motor Cortex",
                 "Temporal Cortex",
                 "Cerebellum"))%>%
    group_by(disease_group2,tissue_clean) %>%
    mutate(n_sample = n_distinct(sample)) %>%
    ungroup() %>%
    mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
    mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                      "Frontal Cortex",
                                      "Lumbar Spinal Cord",
                                      "Temporal Cortex",
                                      "Motor Cortex")) %>%
    ggplot(aes(x = detection_name,
               y = value * 10^6,
               fill = junction_name)) +
    geom_boxplot(show.legend = F,position = position_dodge(preserve = "single")) +
    geom_point(size = 2,pch = 21, alpha = 0.7, position = position_jitterdodge(jitter.width = 0.2))+
    scale_y_log10() +
    ggplot2::facet_wrap(vars(tissue_clean),scales = "free_x",nrow = 3) +
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 12)) +
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
        theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size = 18))  +
  theme(legend.position = "bottom")
```

## UNC13A CE level across disease types

```{r}

disease_comparisons = list( c("Control","ALS-TDP"),
                           c("Control","ALS \n non-TDP"),
                           c("Control","FTLD-TDP"),
                           c("Control","FTLD \n non-TDP" ))

table_to_test = boxplot_table %>%
    filter(disease_tissue == T) %>%
    group_by(disease_group2) %>%
    mutate(n_sample = n_distinct(sample)) %>%
    mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>%
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>%
    dplyr::select(detection_name,
                  participant_id,
                  inclusion_reads,
                  disease_group2,
                  tissue_clean,
                  disease_tissue) %>%
    unique()

test_pair = pairwise.wilcox.test(table_to_test$inclusion_reads, table_to_test$detection_name,
                     p.adjust.method = "BH") %>% broom::tidy()

test_pair = test_pair %>%
  mutate(p_value_draw = case_when(p.value < 0.0001~ "***",
                                  p.value < 0.01 ~ "**",
                                   p.value < 0.05 ~ "*",
                          TRUE ~ paste0("Adj. p-value \n",as.character(round(p.value,2))))) %>%
  mutate(y.position = seq(0.25,by = 0.1,length.out = 7))

table_to_test %>%
    ggplot(aes(x = detection_name, y = inclusion_reads * 10^6)) +
    geom_boxplot() +
    geom_jitter(height = 0) +
    scale_y_log10() +
    ggpubr::theme_pubr() +
    ylab("UNC13A cryptic inclusion \n reads per million") +
    theme(text = element_text(size = 24)) +
    xlab("N samples") +
    stat_pvalue_manual(test_pair %>% filter(p.value < 0.05),
                       label = "p_value_draw",size = 8) +
   stat_compare_means(size = 8)

```

# Detection rate as a linear model
```{r}

```

# UNC13A RNA expression (TPM)

## UNC13A TPM by disease and tissue

```{r}
comps = clean_data_table %>%
  group_by(disease_group2,tissue_clean) %>%
  mutate(n_sample = n_distinct(sample)) %>%
  ungroup() %>%
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>% pull(detection_name) %>%
  unique() %>%
  combn(.,2,simplify = F)

clean_data_table %>%
  group_by(disease_group2,tissue_clean) %>%
  mutate(n_sample = n_distinct(sample)) %>%
  ungroup() %>%
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
 filter(tissue_clean %in%
           c("Frontal Cortex",
             "Lumbar Spinal Cord",
             "Cervical Spinal Cord",
             "Motor Cortex",
             "Temporal Cortex",
             "Cerebellum") )%>%
mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                              "Frontal Cortex",
                              "Lumbar Spinal Cord",
                              "Temporal Cortex",
                              "Motor Cortex")) %>%
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) +
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") +
    xlab("") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none') +
  stat_compare_means(comparisons = comps)
```

```{r make_those_plots_function}
plot_by_tissue = function(tissue_name,dt){

only_tissue = dt %>%
  group_by(disease_group2,tissue_clean) %>%
  mutate(n_sample = n_distinct(sample)) %>%
  ungroup() %>%
  filter(tissue_clean == tissue_name) %>%
  mutate(disease_group2 = gsub("Control", " Control",disease_group2)) %>%
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )"))  %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean))

unique_x_values = only_tissue %>%
  pull(detection_name) %>%
  unique()

comparisons = combn(unique_x_values,2,simplify = F)

the_plot = only_tissue %>%
    ggplot(aes(x = detection_name, y = UNC13A_TPM)) +
    ggpubr::theme_pubr() +
    geom_boxplot(aes(fill = disease_group2)) +
    geom_jitter(pch = 21, color = 'black',width  = 0.2,height = 0,aes(fill =   disease_group2)) +
    ylab("UNC13A TPM") +
    xlab("") +
    guides(color = FALSE) +
    theme(text = element_text(size = 20)) +
  facet_wrap(~tissue_clean, scales = 'free',nrow = 3) +
  theme(legend.position = 'none') +
  stat_compare_means() +
  stat_compare_means(label = "p.signif",comparisons = comparisons,)

  return(the_plot)
}

dis_tis = clean_data_table[disease_tissue == T,unique(tissue_clean)]
dis_tis = c("Cerebellum",dis_tis)
my_plots = purrr::map(dis_tis,plot_by_tissue, clean_data_table)
ggarrange(plotlist=my_plots)

```

## Correlation STMN2 Cryptic PSI and UNC13A TPM

only in ALS/FTLD-TDP

```{r}
clean_data_table %>%
    filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) +
  geom_point() +
  stat_cor(size = 8) +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A TPM") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  ylim(0,50)
```

## Correlation STMN2 Cryptic PSI and UNC13A TPM - tissue separations

```{r}

clean_data_table %>%
    filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = UNC13A_TPM)) +
  geom_point() +
  stat_cor(size = 8) +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A TPM") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = "free_y")
```

## Correlation between UNC13A CE PSI and UNC13A TPM

```{r}
clean_data_table %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(UNC13A_annotated_leaf > 15) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(unc13a_cryptic_leaf_psi > 0) %>%
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13B_TPM)) +
  geom_point() +
  stat_cor(label.y = 20,size = 8,method = 's',cor.coef.name = 'rho') +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 22)) +
  scale_x_continuous(labels = scales::percent)

```

## Correlation between UNC13A CE PSI and UNC13A TPM - tissue separated

```{r}
clean_data_table %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(UNC13A_annotated_leaf > 15) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = UNC13A_TPM, color = disease_group2)) +
  geom_point() +
  stat_cor(size = 8) +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  ylab("UNC13A TPM") +
  xlab("UNC13A CE PSI") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  facet_wrap(~tissue_clean,scales = 'free') +
  scale_x_continuous(labels = scales::percent)

```

# Relationship between STMN2 and UNC13A CE PSI
## Scatter plot showing the correlation in non-log space UNC13A & STMN2 CE

```{r}
clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>%
  # filter(STMN2_cryptic_leaf >= 2) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(unc13a_cryptic_leaf_psi > 0) %>%
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
  geom_point() +
  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A CE PSI") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
```
## Scatter plot showing the correlation in non-log space RAP1GAP1 & STMN2 CE

```{r}
clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>%
  # filter(STMN2_cryptic_leaf >= 2) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(unc13a_cryptic_leaf_psi > 0) %>%
  ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
  geom_point() +
  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("STMN2 Cryptic PSI") +
  ylab("UNC13A CE PSI") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
```

# Relationship between STMN2 and UNC13B fsE PSI
## Scatter plot showing the correlation in non-log space in cortex UNC13B fsE & STMN2 CE

```{r}
clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  mutate(junction_reads_unc13b = UNC13B_nmd_junction_5prime + UNC13B_annotated + UNC13B_nmd_junction_3prime) %>% 
  filter(junction_reads_unc13b >= 30) %>%
  # filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 2) %>%
  # filter(STMN2_cryptic_leaf >= 2) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(unc13a_cryptic_leaf_psi > 0) %>%
  ggplot(aes(x = unc13a_cryptic_leaf_psi, y = unc13b_nmd_psi)) +
  geom_point() +
  stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
  geom_smooth(method = lm, se = FALSE,color = "black") +
  xlab("UNC13A CE PSI") +
  ylab("UNC13B fsE PSI") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 18)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1L))
```

## Side by side comparison UNC13A & STMN2 CE PSI
```{r}
clean_data_table %>%
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
    filter(tissue_clean %in%
               c("Frontal Cortex",
                 "Lumbar Spinal Cord",
                 "Cervical Spinal Cord",
                 "Motor Cortex",
                 "Temporal Cortex",
                 "Cerebellum") )%>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(disease_tissue == T) %>%
  filter(UNC13A_3prime_leaf + UNC13A_5prime_1_leaf >= 1) %>%
  filter(STMN2_cryptic_leaf >= 1) %>%
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  melt() %>%
    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Temporal Cortex")) %>%
  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(aes(color = variable),outlier.shape = NA) +
  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
  scale_fill_manual(values = c("#008574","#cd5582")) +
  scale_color_manual(values = c("#008574","#cd5582")) +
  facet_wrap(~tissue_clean,nrow = 3, scales = 'free') +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("CE PSI")  +
  theme(legend.position = 'none') + 
  stat_compare_means(label = 'p.signif')

```

## Side by side comparison UNC13A & STMN2 CE PSI v2
```{r}
clean_data_table %>%
  filter(UNC13A_annotated_leaf > 10) %>%
  filter(STMN2_annotated_leaf > 10) %>%
  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  # filter(unc13a_cryptic_leaf_psi > 0) %>%
  # filter(disease_tissue == T) %>%
    mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  filter(tissue_clean %in% c("Motor Cortex", "Cervical Spinal Cord", "Frontal Cortex", "Lumbar Spinal Cord", "Temporal Cortex")) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_rat = log2((unc13a_cryptic_leaf_psi + 0.1) / (stmn_2_cryptic_psi_leaf + 0.1))) %>%
  group_by(tissue_clean, disease_group2) %>%
  mutate(average_log2Fold = mean(log2_rat)) %>%
  ungroup() %>%
  dplyr::select(-log2_rat) %>%
  pivot_longer(cols = c("stmn_2_cryptic_psi_leaf","unc13a_cryptic_leaf_psi")) %>%
  mutate(tissue_clean = fct_reorder(tissue_clean,average_log2Fold)) %>%
  mutate(variable = ifelse(grepl("stmn_2",name), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = tissue_clean, y = value,fill = name)) +
  geom_boxplot() +
  geom_point(pch = 21,height = 0,width = 0.2,position = position_jitterdodge()) +
  scale_fill_manual(values = c("#004D40","#882255")) +
  facet_wrap(~tissue_clean+disease_group2,nrow = 1,scales = 'free')  +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position = 'none')
```

## Side by side comparison UNC13A & STMN2 CE PSI - v3
```{r}
clean_data_table %>%
  # filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  # filter(unc13a_cryptic_leaf_psi > 0) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  melt() %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
    mutate(tissue_clean = fct_relevel(tissue_clean,"Motor Cortex", "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Temporal Cortex")) %>%
  mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(aes(color = variable)) +
  geom_jitter(aes(fill = variable),pch = 21,height = 0,width = 0.2,size = 3) +
  scale_fill_manual(values = c("#004D40","#882255")) +
  scale_color_manual(values = c("#004D40","#882255")) +
  facet_wrap(~tissue_clean+disease_group2,nrow = 3,scales = 'free') +
    scale_y_continuous(labels = scales::percent)  +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position = 'none')

```

## Log2FC UNC13A & STMN2 CE PSI
```{r}
clean_data_table %>%
  filter(UNC13A_annotated_leaf > 10) %>%
  filter(STMN2_annotated_leaf > 10) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(unc13a_cryptic_leaf_psi > 0) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,disease_group2,tissue_clean,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf)))%>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%

  # mutate(variable = ifelse(grepl("stmn_2",variable), "STMN2", "UNC13A")) %>%
  ggplot(aes(x = tissue_clean,y = log2_unc, fill = tissue_clean)) +
  geom_boxplot(position = 'dodge') +
  # geom_jitter(height = 0,width = 0.2,pch = 21) +
  facet_wrap(~disease_group2,scales = "free_x") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size = 24))  +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
```

# Effect of UNC13A genotype

## UNC13A psi by genotype - rs12973192 - RE

```{r}
rs12973192_comparisons = clean_data_table %>%
  filter(junction_reads_unc13a >=30) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )"))

comp_for_gg = rs12973192_comparisons %>% pull(detection_name) %>%
  unique() %>%
  combn(.,2,simplify = F)

ce_psi = rs12973192_comparisons %>%
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  ylab("UNC13A PSI") +
  xlab(element_blank()) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means() +
  theme(legend.position = "none") +
  stat_compare_means(comparisons = comp_for_gg,label = "p.signif") +
  scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
  ggtitle("CE SNP (rs12973192)")

```

## UNC13A psi by genotype - rs12608932 - RI
```{r}
rs12608932_comparisons = clean_data_table %>%
  filter(junction_reads_unc13a >=30) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  group_by(rs12608932) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )"))

comp_for_gg = rs12608932_comparisons %>% pull(detection_name) %>%
  unique() %>%
  combn(.,2,simplify = F)

in_psi = rs12608932_comparisons %>%
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = detection_name, fill = rs12608932)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  ylab(element_blank()) +
  xlab(element_blank()) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(legend.position = "none") +
  theme(text = element_text(size = 24)) +
  stat_compare_means() +
  stat_compare_means(comparisons = comp_for_gg,label = "p.signif") +
    scale_y_continuous(trans='log10',labels = scales::percent_format(accuracy = 1L))  +
  ggtitle("intronic SNP (rs12608932)")
```

```{r}
ggpubr::ggarrange(ce_psi,in_psi)
```

## UNC13A psi by genotype - rs12973192
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## UNC13A psi by genotype - tissue split
```{r}
genotype_comparisons = list(c("C/C", "C/G"), c("C/C", "G/G"), c("C/G", "G/G"))

clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## UNC13A psi in detected tissues by genotype - tissue split
```{r}

clean_data_table %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  ggplot(aes(y = unc13a_cryptic_leaf_psi, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("UNC13A PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi by genotype
```{r}

clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi by genotype - tissue split
```{r}

clean_data_table %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## STMN2 psi in detected tissues by genotype - tissue split
```{r}

clean_data_table %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(tissue_clean = gsub("_"," ",tissue_clean)) %>%
  ggplot(aes(y = stmn_2_cryptic_psi_leaf, x = rs12973192, fill = rs12973192)) +
  geom_boxplot() +
  geom_jitter( height = 0) +
  facet_wrap(~disease_group2 + tissue_clean,scales = 'free') +
  ylab("STMN2 PSI") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  stat_compare_means(comparisons = genotype_comparisons,label = "p.signif")

```

## Correlation UNC13A PSI and STMN2 PSI with genotype effect - concordants samples

```{r}

clean_data_table %>%
  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
  mutate(snp_status = case_when(rs12973192 == "C/C" ~ "healthy/healthy",
                                rs12973192 == "C/G" ~ "healthy/risk",
                                rs12973192 == "G/G" ~ "risk/risk")) %>% 
  filter(disease_tissue == T) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
    mutate(rs12973192 = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>%
  filter(!grepl("Cord",tissue_clean)) %>%
  mutate(log10_stmn2 = log10(stmn_2_cryptic_psi_leaf)) %>%
    mutate(log10_unc13a = log10(unc13a_cryptic_leaf_psi)) %>%
    ggpubr::ggscatter(.,
                      x = "log10_stmn2",
                      y = "log10_unc13a",
                      fill = 'snp_status',
                      color = 'black', 
                      size = 3,shape = 21,
                      ) +
    ylab("UNC13A CE PSI ") +
    xlab("STMN2 Cryptic PSI ") +
  geom_smooth(method = 'lm',aes(color = snp_status),se = F) + 
  stat_cor(aes( x = stmn_2_cryptic_psi_leaf,
                      y = unc13a_cryptic_leaf_psi,
                      color = snp_status),
           method = 'spearman',
           cor.coef.name = 'rho',
           show.legend = FALSE,
           size = 14) +
    scale_color_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
      scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +

  theme(legend.title = element_blank()) +
  theme(text = element_text(size = 24),legend.text = element_text(size = 32)) +
    ylab(expression(paste("Lo", g[10], "UNC13A CE PSI"))) +
      xlab(expression(paste("Lo", g[10], "STMN2 CE PSI")))

```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex only

```{r}
detected_table_for_graph = clean_data_table %>%
  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3)) %>% as.data.table()

#get out that person who is discongruous

comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F)

detected_table_for_graph %>%
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) +
  geom_boxplot(show.legend = F) +
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
  stat_compare_means() +
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") +
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5)
```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12973192 - cortex and cord

```{r}
detected_table_for_graph = clean_data_table %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3))

comparisons = combn(unique(detected_table_for_graph$detection_name),2,simplify = F)

detected_table_for_graph %>%
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) +
  geom_boxplot(show.legend = F) +
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
  stat_compare_means( label = "p.format") +
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") +
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5)
```

```{r}
pretty_effect_plot = function(model, p_value_cutoff = 0.05, title = "Model 1"){
    p_value = Tmisc::lmp(model)
    r_squared = summary(model)$adj.r.squared
    pretty_plot = broom::tidy(model) %>% 
        filter(p.value < p_value_cutoff) %>%
        mutate(term = fct_reorder(term, -abs(estimate))) %>%
        ggplot(aes(x = term, y = estimate,fill = estimate > 0 )) +
        geom_col() +
        coord_flip() +
        ggtitle(title,subtitle = glue::glue("Model p-value: {p_value} \n Adj. R-squared: {r_squared}"))
}

detect2 = clean_data_table %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(disease_tissue == T) %>% 
  mutate(call = fct_relevel(rs12973192,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(call)) %>% 
  mutate(age_at_death = as.numeric(age_at_death))

m1 = lm(unc13a_cryptic_leaf_psi ~ stmn_2_cryptic_psi_leaf * 
          number_g_alleles + 
          tissue_clean + 
          age_at_death + 
          sex, 
              detect2)

k = pretty_effect_plot(m1,p_value_cutoff = 0.05)
print(k)

```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - rs12608932

```{r}
detected_table_for_graph_rs12608932 = clean_data_table %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
  group_by(rs12608932) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12608932} \n ( {n_sample} )"))

comparisons = combn(unique(detected_table_for_graph_rs12608932$detection_name),2,simplify = F)

detected_table_for_graph_rs12608932 %>%
  ggplot(aes(y = log2_unc, x = detection_name,fill = detection_name)) +
  geom_boxplot(show.legend = F) +
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
  stat_compare_means( label = "p.format") +
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") +
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12608932 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5)
```

## Log2 Fold UNC / STMN2 CE in TDP-path with both detected - STR - cortex and cord

```{r str and log2 graph}
str_genotype = janitor::clean_names(fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/rs56041637.genotypes.377samples.txt")) %>% 
  dplyr::rename(presumed_unique_id = number_sample_id)

detected_table_for_graph = clean_data_table %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(unc13a_cryptic_leaf_psi > 0 ) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter(junction_reads_unc13a >= 30) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,participant_id,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,unc13a_cryptic_leaf_psi) %>%
  mutate(log2_unc = log2((unc13a_cryptic_leaf_psi)/(stmn_2_cryptic_psi_leaf))) %>% 
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>% 
  left_join(str_genotype) %>% 
    filter(!is.na(genotype_ci)) %>% unique()

comparisons = combn(unique(detected_table_for_graph$genotype),2,simplify = F)

detected_table_for_graph %>%
  ggplot(aes(y = log2_unc, x = genotype,fill = genotype)) +
  geom_boxplot(show.legend = F) +
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
  stat_compare_means( label = "p.format") +
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab("Ratio UNC13A CE PSI / STMN2 PSI") +
  ylab(expression(paste("Lo", g[2], " fold UNC13A CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A STR Genotype") +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5)
```

## STR by SNP genotype

```{r}
clean_data_table %>% 
  mutate(presumed_unique_id = gsub("-b38","",participant_id)) %>% 
  select(participant_id,presumed_unique_id,rs12973192,rs12608932) %>% 
  unique() %>% 
  left_join(str_genotype) %>% 
  unique() %>% 
  mutate(genotype = fct_relevel(genotype,"6/6", "6/7", "6/8", "6/9",
                                "6/10","6/12","7/8","8/8","8/10","9/10")) %>% 
  ggplot(aes(x = genotype,fill = rs12973192)) + 
  geom_bar() + 
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ylab("N individuals") +
  xlab("STR genotype")

```

## Detection rate by genotype

```{r}
overall_fisher = clean_data_table %>%
    filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>%
  dplyr::select(participant_id,unc13a_detected,rs12973192) %>%
  unique() %>%
  group_by(rs12973192) %>%
  mutate(n_sample = n_distinct(participant_id)) %>%
  mutate(n_sample_detected = sum(unc13a_detected)) %>%
  dplyr::select(rs12973192,n_sample,n_sample_detected) %>%
  unique() %>%
  mutate(n_non_detected = n_sample - n_sample_detected) %>%
  dplyr::select(-n_sample)

over_p = overall_fisher %>%
  column_to_rownames('rs12973192') %>%
  chisq.test() %>%
  broom::tidy() %>%
  .$p.value

overall_fisher %>%
  mutate(n_sample = n_non_detected + n_sample_detected) %>%
  mutate(detection_rate = n_sample_detected / n_sample) %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
  geom_col(show.legend = F) +
  ggpubr::theme_pubr() +
  scale_y_continuous(labels = scales::percent) +
  ylab("% of TDP-43 Proteionopathy Patients \n UNC13A Cryptic Detected") +
  theme(text = element_text(size = 18)) +
  xlab("N individuals") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))

```

Although this difference is not significant, with the Fisher's exact giving
a p-value of `r round(over_p,2)`.

# Variant allele expression in TDP-43 KD experiments

Across our KD's, we observed a clear allelic bias in the SK-DZ-N cells

```{r}
kd_vafs = fread(file.path(here::here(),"data","all_kds_unc13.snp.out.tsv"))

kd_vafs %>%
    left_join(rel_rna_cryptic_amount) %>%
    filter(!is.na(condition)) %>%
    mutate(VAF = alt_reads / total) %>%
    filter(condition != "control" ) %>%
    filter(alt_reads > 5) %>%
    ggplot(aes(x = stmn_2_cryptic_psi,y = cryptic_psi, fill = -(VAF - 0.5),size = total)) +
    geom_point(pch = 21)  +
    coord_fixed() +
  ylim(0,0.8) +
  xlim(0,0.8) +
  geom_abline() +
    scale_fill_gradient2()

```

# Liu Facs Neurons

Also look at the coverage in the Liu nuclear facs sorted neurons

```{r}
liu_vafs = fread(file.path(here::here(),"data","liu_facs_neurons_unc13.snp.out.tsv"))
liu_stmn2 = fread(file.path(here::here(),"data","liu_stmn2_and_unc13aaggregated.clean.annotated.bed"))
liu_stmn2[,sample := gsub(".SJ.out","",V4)]
liu_stmn2 = liu_stmn2 %>%
  unique() %>%
    pivot_wider(id_cols = 'sample',
                names_from = "V7",
                values_from = 'V5',
                values_fill = 0)

liu_stmn2 <- liu_stmn2 %>%
   mutate(stmn2_psi = STMN2_cryptic / (STMN2_cryptic + STMN2_annotated)) %>%
   mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime + UNC13A_5prime_2))
```

```{r}
liu_vafs %>%
    mutate(VAF = alt_reads / total) %>%
    data.table::melt(id.vars = c("sample",'VAF','total')) %>%
    mutate(allele = ifelse(variable == "ref_reads", "C","G")) %>%
    mutate(sample_name = sub("_TDP_.*", "", sample)) %>%
    mutate(sample_name = gsub("_unsorted", "", sample_name)) %>%
    mutate(condition = sub(".*_", "", sample)) %>%
    ggplot(aes(x = condition,y = value, fill = allele)) +
    geom_col(pch = 21, size = 4) +
    facet_wrap(~sample_name,scales = "free")

liu_stmn2 %>%
  dplyr::select(stmn2_psi,sample,unc13a_psi) %>%
  data.table::melt() %>%
   mutate(sample_name = sub("_TDP_.*", "", sample)) %>%
    filter(!grepl("unsorted",sample_name)) %>%
    mutate(condition = sub(".*_", "", sample)) %>%
    mutate(condition = ifelse(grepl("negative",condition), "TDP-43 \nNegative", "TDP-43 \nPositive")) %>%
    mutate(variable = ifelse(grepl("stmn2",variable), "STMN2", "UNC13A")) %>%
    ggplot(aes(x = condition,y = value, fill = variable)) +
    geom_col(size = 4,position = 'dodge') +
    facet_wrap(~sample_name,scales = "free_x") +
    scale_fill_manual(values = c("#004D40","#882255")) +
    theme_bw() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    theme(text = element_text(size = 24))  +
      xlab("") +
    ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
```

```{r}
liu_stmn2 %>%
  dplyr::select(stmn2_psi,sample,unc13a_psi) %>%
  data.table::melt() %>%
   mutate(sample_name = sub("_TDP_.*", "", sample)) %>%
    filter(!grepl("unsorted",sample_name)) %>%
    mutate(condition = sub(".*_", "", sample)) %>%
    mutate(condition = ifelse(grepl("negative",condition), "TDP-43 \nNegative", "TDP-43 \nPositive")) %>%
    mutate(variable = ifelse(grepl("stmn2",variable), "STMN2", "UNC13A")) %>%
  ggbarplot(,
            x = "condition",
            add = c("mean_se",'jitter'),
            y = "value",
            fill = 'variable',
            color = 'variable',
            position = position_dodge(0.8),dot.size = 6) +
  ggpubr::theme_pubr() +
  scale_color_manual(values = c("#000000","#000000")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(text = element_text(size = 24))  +
  scale_y_continuous(labels = scales::percent) +
  xlab("") +
  ylab("PSI")  +
  theme(legend.position =  'top') +
  theme(legend.title=element_blank()) +
  stat_compare_means(comparisons = list(c("STMN2","UNC13A")))
```

```{r}
liu_stmn2 %>%
   mutate(stmn2_psi = STMN2_cryptic / (STMN2_cryptic + STMN2_annotated)) %>%
   mutate(unc13a_psi = (UNC13A_3prime + UNC13A_5prime)/ (UNC13A_annotated + UNC13A_3prime + UNC13A_5prime)) %>%
   left_join(liu_vafs) %>%
    mutate(alt_vaf = (alt_reads)/total) %>%
    mutate(condition = ifelse(grepl("negative",sample),"negative", "positive")) %>%
             filter(!grepl("unsorted",sample)) %>%

    ggplot(aes(y = unc13a_psi,x = stmn2_psi, fill = alt_vaf)) +
    geom_point(pch = 21,size = 5) +
    ggtitle("STMN2 PSI and the VAF of the G allele \n in the Liu Nuclei ") +
  facet_grid(~condition) +
  scale_fill_viridis_c(option = "plasma") +
  coord_fixed() +
  geom_abline()
```

## Variant allele expression in patients

```{r}
nycg_vafs = fread(file.path(here::here(),"data","NYGC_all_samples_UNC13A_snp_coverage.tsv"),fill = T)
nycg_vafs %>%
    left_join(clean_data_table) %>%
    filter(rs12973192 == "C/G") %>%
    mutate(fraction_risk = alt_reads / total) %>%
    filter(!is.na(fraction_risk)) %>%
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = unc13a_cryptic_leaf_psi)) +
    geom_point(size = 4, pch = 21, aes(fill = fraction_risk - 0.5)) +
    xlab("STMN2 Cryptic PSI") +
    ylab("UNC13A CE PSI") +
    geom_abline() +
    scale_fill_gradient2()

nycg_vafs %>%
    left_join(clean_data_table) %>%
    filter(rs12973192 == "C/G") %>%
    mutate(fraction_risk = alt_reads / total) %>%
    filter(!is.na(fraction_risk)) %>%
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
    filter(unc13a_cryptic_leaf_psi > 0 ) %>%
    mutate(stmn2_unc_fract = ((unc13a_cryptic_leaf_psi * 100) / (stmn_2_cryptic_psi_leaf * 100))) %>%
    ggplot(aes(x = stmn2_unc_fract, y = fraction_risk)) +
    geom_point(size = 4, pch = 21, aes(fill = fraction_risk - 0.5)) +
    xlab("STMN2 Cryptic PSI") +
    ylab("UNC13A CE PSI") +
    scale_fill_gradient2()
```

# Ribosome profiling
## Ribosome profiling quality control

TODO: convert all libraries to p_load.

Perform quality control to check that reads are periodic.

```{r message=FALSE, warning=FALSE}
read_length_filter <- 29  # length of reads of interest
gencode_v29_info <- fread(paste0(here(), "/data/longest_proteincoding_transcript_hs_details.txt"))
ribo_df <- fread(paste0(here(), "/data/riboseq/sh_tdpb.Aligned.toTranscriptome.out.bam.bed.gz")) 
names(ribo_df)  = c("transcript_id", "start0", "end", "strand")
ribo_df = ribo_df %>% 
    inner_join(gencode_v29_info, by = "transcript_id") %>%
    dplyr::filter(end-start0 == read_length_filter, strand == "+") %>%
    mutate(distance_from_start = start0-cds_start) %>%
    group_by(distance_from_start) %>%
    mutate(n_this_dist = n()) %>%
    dplyr::select(distance_from_start, n_this_dist) %>%
    unique()

```

```{r, warning=FALSE}
ggplot(ribo_df, aes(x=distance_from_start, y=n_this_dist)) +
    geom_line() +
    geom_point() +
    xlim(-50, 50) +
    ylab("Counts") +
    xlab("Distance of 5' end from annotated start codon") +
    ggtitle(paste0("Periodicity of ", read_length_filter, "nt reads from TDP-43 KD replicate B"))
```

## Ribosome profiling volcano plot

Use DESeq to detect differential ribosome footprints and create volcano plots

```{r message = FALSE, warning=FALSE}
counts_df <- read_csv(paste0(here(), "/data/riboseq/cds_feature_counts.csv"))
counts <- as.matrix(counts_df %>% column_to_rownames("gene_id"))
ipsc_splicing <- read_csv(paste0(here(), "/data/ipsc_splicing_results.csv"))
ipsc_de = read_csv(paste0(here(), "/data/ipsc_differential_expression.csv"))

col_data <- data.frame(sample = colnames(counts)) %>%
  mutate(condition = ifelse(str_detect(sample, "tdp"), "tdp_ko", "control"))

dds <- DESeqDataSetFromMatrix(countData = counts,
                              colData = col_data,
                              design= ~condition)
dds <- DESeq(dds)
results_name <- resultsNames(dds)[2]
res <- results(dds, name=results_name)

res_df <- data.frame(res) %>%
  rownames_to_column("gene_id") %>%
  left_join(gencode_v29_info, by = "gene_id")

cryptic_df <- ipsc_splicing %>%
  dplyr::select(gene_name, cryptic_junction) %>%
  unique() %>%
  filter(cryptic_junction)

res_df2 <- res_df %>%
  mutate(cryptic_junction = gene_name %in% cryptic_df$gene_name) %>%
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2","TARDBP") ~ gene_name)) %>%
  mutate(colour = ifelse(cryptic_junction, "#fe6101",  "#648FFF")) %>%
  arrange(colour)
```

```{r, warning=FALSE}
p1 <- ggplot(res_df2, aes(x = log2FoldChange, y = -log10(pvalue), label = label_junction)) +
  geom_point(data = res_df2 %>% filter(!cryptic_junction),
             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#648FFF") +
  geom_point(data = res_df2 %>% filter(cryptic_junction),
             aes(x = log2FoldChange, y= -log10(pvalue)), color = "#fe6101") +
  ggpubr::theme_pubr() +
  ylab("-Log10(p-value)") +
  xlab("Log2 fold change") +
  geom_text_repel(box.padding = 1.2, max.overlaps = Inf)

p2 <- p1 + ylim(0,10)
p1
p2
```

# Targeted RNA-seq

Read in data from processed bam files

TODO: add scripts which deduplicate UMIs for each.

```{r message=FALSE, warning=FALSE}
specific_rt <- read_csv(paste0(here(), "/data/targeted_RNA_seq/specific_RT_primer/for_plot.csv")) %>%
    select(risk, sample_name, n)

random_hexamer <- read_csv(paste0(here(), "/data/targeted_RNA_seq/random_hexamer/values_for_plot.csv")) %>%
    select(-sample_name) %>%
    select(sample_name = actual_sample_name, risk, n)

tg_rnaseq_df <- bind_rows(specific_rt, random_hexamer) %>%
  group_by(sample_name) %>%
  mutate(n_healthy = ifelse(risk == "Non-Risk", n, 0)) %>%
  mutate(combined_healthy = sum(n_healthy)) %>%
  select(-n_healthy) %>%
  mutate(combined = sum(n)) %>%
  group_by(sample_name, risk) %>%
  mutate(combined_n = sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  unique() %>%
  group_by(sample_name) %>%
  mutate(p_value = pbinom(combined_healthy, sum(combined_n), 0.5)) %>%  # single-tailed bionomial test
  ungroup() %>%
  mutate(sample_name_ordered = fct_reorder(sample_name, dplyr::desc(combined_n)))

ggplot(tg_rnaseq_df %>%
         filter(combined_n > 0), aes(x = sample_name_ordered, y = combined_n, fill = risk)) +
  geom_bar(stat="identity", position="dodge") +
  geom_text(aes(x=sample_name, y=-2.5, label=signif(p_value,2)), size = 3.5) +
  xlab("") +
  ylab("Unique cDNAs") +
  ggeasy::easy_add_legend_title("Allele") +
  theme_minimal() +
  ggpubr::theme_pubr() +
  ggeasy::easy_rotate_x_labels()
```

# iCLIP of minigenes

Read in the processed data (following alignment with Bowtie2, and filtering for high-confidence alignments), and generate
a plot of the overall distribution of crosslinks, and the rolling average difference between the 2x Risk and 2x Healthy.

Next, filter for peaks in the iCLIP signal, and see how they change in 2x Risk versus 2x Healthy. Highlight those which are close to the SNPs.

TODO: add script which does this filtering
TODO: add fasta files for the minigenes

```{r message=FALSE, warning=FALSE}
exon_snp <- 556
intron_snp <- 1106
intron_start <- 141
intron_end <- 1429
acceptor1 <- 394
acceptor2 <- 444
donor <- 572
mg_l <- 1624
cpoor_start <- 490
cpoor_end <- 1325
risk_colour <- "#00BFC4"
healthy_colour <- "#F8766D"
confidence_interval <- 0.75  # fracion, i.e. 0.75 = 75%.
peak_threshold <- 5 # how many times above the mean to be a "peak"
region_length <- 50

filtered_df <- read_csv(paste0(here(), "/data/iCLIP/bowtie2_aligned_2_filtered_df.csv"))

rolling_window <- 20
peak_t <- peak_threshold*1000/mg_l  # by definition...

df2 <- filtered_df %>%
  ungroup() %>%
  group_by(condition, pos) %>%
  mutate(mean_counts = sum(normalised_counts)/2) %>%
  select(pos, condition, mean_counts) %>%
  unique() %>%
  pivot_wider(names_from = "condition", values_from ="mean_counts")

df2[is.na(df2)] <- 0

df2 <- df2 %>%
  mutate(diff = R-H) %>%
  mutate(mean_4 = (H+R)/2) %>%
  arrange(pos)

top <- ggplot(df2, aes(x=pos, y=diff)) +
  geom_area(aes(x=pos, y = mean_4), colour="grey50", fill = "grey50", alpha=0.2) +
  geom_vline(xintercept = exon_snp, linetype="dashed") +
  geom_vline(xintercept = intron_snp, linetype="dashed") +
  xlim(1,1600) +
  theme_minimal() +
  ylab("XLs per 1000") +
  xlab("") +
  geom_vline(xintercept = cpoor_start) +
  geom_vline(xintercept = cpoor_end)

bottom <- ggplot(df2, aes(x=pos, y=zoo::rollmean(diff, k = rolling_window, na.pad=T))) +
  geom_area(colour = "grey50", fill="grey50", alpha = 1) +
  geom_vline(xintercept =  exon_snp, linetype="dashed") +
  geom_vline(xintercept =  intron_snp, linetype="dashed") +
  xlim(1,1600) +
  ylim(-0.5,0.75) +
  theme_minimal() +
  ylab("Change in XL density") +
  xlab("Distance from start of minigene") +
  geom_vline(xintercept =  intron_start) +
  geom_vline(xintercept =  intron_end) +
  geom_vline(xintercept =  acceptor1) +
  geom_vline(xintercept =  acceptor2) +
  geom_vline(xintercept =  donor)

combined <- top/bottom
combined

find_ratio_bound <- function(e_x, sem_x, e_y, sem_y, quantile_value, randoms = 10000){
  # monte carlo confidence interval estimate for a ratio. Assumes the two variables
  # x and y are normally distributed. Finds confidence interval for (x-y)/y

  # randomly generate possible global means for x and y

  xs = rnorm(randoms, mean = e_x, sd=sem_x)  # risk
  ys = rnorm(randoms, mean = e_y, sd=sem_y)  # healthy

  ratios = (xs-ys)/ys

  return(quantile(ratios, quantile_value))
}

peak_df <- filtered_df %>%
  ungroup() %>%
  group_by(pos) %>%
  mutate(max_at_pos = max(normalised_counts)) %>%  # at least 1 of the 4 samples above the threshold
  ungroup() %>%
  filter(max_at_pos >= peak_t) %>%  # at least 1 of the 4 samples above the threshold
  group_by(condition, pos) %>%
  mutate(peak_mean = mean(normalised_counts),
         peak_sd = sd(normalised_counts)) %>%
  ungroup() %>%
  select(condition, pos, peak_mean, peak_sd) %>%
  unique() %>%
  pivot_wider(names_from = "condition", values_from = c("peak_mean", "peak_sd")) %>%
  mutate(frac_increase = (peak_mean_R-peak_mean_H)/peak_mean_H) %>%
  mutate(peak_rank = rank(frac_increase)) %>%
  mutate(colour = ifelse(abs(pos-exon_snp)<region_length, "gold",
                         ifelse(abs(pos-intron_snp)<region_length, "purple", "grey50")))

peak_df$upper_bound <- mapply(find_ratio_bound,
                        peak_df$peak_mean_R,
                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
                        peak_df$peak_mean_H,
                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
                        0.5*(1+confidence_interval))

peak_df$lower_bound <- mapply(find_ratio_bound,
                        peak_df$peak_mean_R,
                        peak_df$peak_sd_R/sqrt(2), #convert to SEM
                        peak_df$peak_mean_H,
                        peak_df$peak_sd_H/sqrt(2), #convert to SEM
                        0.5*(1-confidence_interval))

ggplot(peak_df, aes(x = peak_rank, y = 100*frac_increase, fill=colour)) +
  geom_bar(stat="identity") +
  scale_fill_identity() +
  theme_minimal() +
  ggpubr::theme_pubr()+
  ylab("% change in 2x Risk") +
  xlab("Rank") +
  geom_errorbar(aes(ymin = 100*lower_bound, ymax = 100*upper_bound, x = peak_rank),
                width = 0.3)

```

Supplementary iCLIP Figures

```{r message=FALSE, warning=FALSE}

bars <- filtered_df %>%
  ungroup() %>%
  group_by(condition, pos) %>%
  mutate(mean_normalised = mean(normalised_counts),
         sd_normalised = sd(normalised_counts)) %>%
  mutate(colour = ifelse(condition == "H", healthy_colour, risk_colour))

bar_top <- ggplot(bars %>% filter(abs(pos-exon_snp) < 35), aes(x = pos, y = mean_normalised,
                                                       fill = colour)) +
  geom_bar(stat="identity", position="dodge") +
  scale_fill_identity() +
  geom_errorbar(aes(ymin = mean_normalised-sd_normalised, ymax=mean_normalised+sd_normalised),
                position="dodge") +
  geom_vline(xintercept = exon_snp, linetype="dashed") +
  ylab("XLs per 1000") +
  ggpubr::theme_pubr() +
  ggtitle("iCLIP signal around CE SNP") +
  xlab("")

bar_bottom <- ggplot(bars %>% filter(abs(pos-intron_snp) < 35), aes(x = pos, y = mean_normalised,
                                                       fill = colour)) +
  geom_bar(stat="identity", position="dodge") +
  scale_fill_identity() +
  geom_errorbar(aes(ymin = mean_normalised-sd_normalised, ymax=mean_normalised+sd_normalised),
                position="dodge") +
  geom_vline(xintercept = intron_snp, linetype="dashed") +
  ggpubr::theme_pubr() +
  ylab("XLs per 1000") +
  ggtitle("iCLIP signal around intronic SNP") +
  xlab("")

bar_combined <- bar_top / bar_bottom
bar_combined

```

# Heptameric analysis

Analyse E-values (higher E value = stronger binding enrichment) for the motifs present at the exon SNP.

```{r message=FALSE, warning=FALSE}

# Read in supplementary 4 from Ray et al 2013

data <- read_tsv(paste0(here(), "/data/heptamers/41586_2013_BFnature12311_MOESM334_ESM.txt"))

for(type in c("exon", "intron")){

    if(type == "exon"){
        snp <-     "TAAAAGCATGGAT"
        healthy <- "TAAAAGGATGGAT"
    } else {
        snp <-     "GGATGGGTGGATA"
        healthy <- "GGATGGTTGGATA"
    }

    all_healthy <- c()
    all_risk <- c()

    # Make dataframe containing E scores for each relevant 7mer

    for(i in 1:(str_length(snp)-6)){
        this_snp <- str_sub(snp, i, i+6)
        this_healthy <- str_sub(healthy, i, i+6)

        all_healthy <- c(all_healthy, str_replace_all(this_healthy, "T", "U"))
        all_risk <- c(all_risk, str_replace_all(this_snp, "T", "U"))

        row1 <- data %>% filter(`7mer` == str_replace_all(this_snp, "T", "U")) %>%
            mutate(snp = T, pos = i)

        row2 <- data %>% filter(`7mer` == str_replace_all(this_healthy, "T", "U")) %>%
            mutate(snp = F, pos = i)

        if(i == 1){
            full <- bind_rows(row1, row2)
        } else {
            full <- bind_rows(full, row1, row2)
        }
    }

    # Find average change in E score for the two groups of heptamers

    full2 <- full %>%
        group_by(pos) %>%
        filter(n() == 2) %>%
        ungroup() %>%
        pivot_longer(cols = contains("set")) %>%
        select(-`7mer`) %>%
        pivot_wider(names_from = "snp", values_from = "value") %>%
        mutate(abs_change = `TRUE`-`FALSE`) %>%
        filter(!is.na(`TRUE`) & !is.na(`FALSE`)) %>%
        group_by(name) %>%
        mutate(mean_abs_change = mean(abs_change)) %>%
        select(name, mean_abs_change) %>%
        unique() %>%
        arrange(mean_abs_change) %>%
        mutate(is_tdp43 = ifelse(str_detect(name, "RNCMPT00076"), "red", "grey80")) %>%
        mutate(short_name = str_sub(name, 1, 11)) %>%
        ungroup() %>% group_by(short_name) %>%
        mutate(average_delta = mean(mean_abs_change)) %>%
        select(short_name, is_tdp43, average_delta) %>%
        unique()

    # Plot ranks

    plot <- ggplot(full2, aes(x=rank(average_delta), y = average_delta,
                      fill = is_tdp43)) +
        geom_bar(stat="identity") +
        scale_fill_identity() +
        theme_minimal() +
        ggtitle(paste0("Average change in binding E value for each dataset with SNP for ", type, " SNP")) +
        xlab("") +
        ylab("Average change in E score") +
        ggpubr::theme_pubr() +
        xlab("Rank")

    print(plot)

    full_just_tdp <- full %>%
        group_by(pos) %>%
        filter(n() == 2) %>%
        ungroup() %>%
        select(`7mer`, contains("RNCMPT00076"), snp, pos) %>%
        mutate(mean_E = (RNCMPT00076_e_setA + RNCMPT00076_e_setB)/2) %>%
        pivot_longer(cols = contains("RNCMP")) %>%
        mutate(ypos = ifelse(snp, -0.35, -0.27))

    plot2 <- ggplot(full_just_tdp, aes(x = pos, y = value, colour = snp,
                              label = `7mer`)) +
        geom_point(size = 3) +
        xlab("7mer") +
        ylab("E value (similar to rank?)") +
        ggeasy::easy_add_legend_title("Contains Risk SNP") +
        ggtitle(paste0("Binding E value for TDP43 for all 7x 7mers with or without risk SNP for ", type, " SNP")) +
        geom_label(aes(y = ypos), size = 2.7) +
        xlim(0.5, 7.5) +
        theme_bw() +
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())  +
        ggeasy::easy_remove_x_axis()

    print(plot2)

}

```

# Isothermal Titration Calorimetry

Read in values, plot and perform a t-test

```{r warning=FALSE, message=FALSE}
df <- read_csv(paste0(here(),
                      "/data/itc/itc final values.csv")) %>%
    group_by(RNA) %>%
    mutate(sd =sd(kd)) %>%
    mutate(mean_kd = mean(kd))

ggplot(df, aes(x = RNA, y = kd)) +
    geom_bar(aes(x = RNA, y = mean_kd), stat="identity", position="dodge") +
    geom_point() +
    geom_errorbar(aes(x=RNA, ymin = mean_kd-sd, ymax = mean_kd+sd),
                  width = 0.3) +
    ylim(0, NA) +
    ggpubr::theme_pubr() +
    ylab("Kd / nM") +
    xlab("")

t.test(df$kd[which(df$RNA=="H")], df$kd[which(df$RNA=="R")])
```

```{r als_ftd_genes_venn}
als_ftd_genomics = c("ALS2", "ANG", "APP", "AR", "ATN1", "ATXN1", "ATXN10", "ATXN2", 
"ATXN3", "C21orf2", "C9orf72", "CHMP2B", "CSF1R", "DCTN1", "DNAJC5", 
"DNMT1", "EPM2A", "FIG4", "FUS", "GRN", "HNRNPA1", "HTT", "ITM2B", 
"JPH3", "KIF5A", "LOC101927815", "MAPT", "NHLRC1", "NOP56", "NOTCH3", 
"OPTN", "PFN1", "PMP22", "PRNP", "PSEN1", "PSEN2", "SCFD1", "SETX", 
"SIGMAR1", "SLC52A2", "SLC52A3", "SOD1", "TARDBP", "TBK1", "TBP", 
"TNIP1", "TYROBP", "UBQLN2", "UNC13A", "VAPB", "VCP")

ipsc_de = as.data.table(ipsc_de)
ipsc_splicing = as.data.table(ipsc_splicing)

ipsc_de[gene_name %in% als_ftd_genomics][order(log2FoldChange)][,.(log2FoldChange,padj,gene_name,gene_expresion)] %>% mutate(has_cryptic = gene_name %in% ipsc_splicing[cryptic_junction == T,gene_name]) %>% unique() %>% arrange(log2FoldChange) %>% clipr::write_clip()
```

## Linear model on UNC13A CE PSI with covariates
```{r the_silly_linear}
# library size 
#llumina HiSeq 2500 (125 bp paired end) or an Illumina NovaSeq (100 bp paired end). 
full_cov = clean_data_table %>%
    filter(tissue_clean %in% c("Frontal_Cortex", 
                               "Lumbar_Spinal_Cord", 
                               "Cervical_Spinal_Cord", 
                               "Motor_Cortex", 
                               "Temporal_Cortex")) %>% 
    mutate(inclusion_reads = UNC13A_3prime_leaf + UNC13A_5prime_1_leaf) %>% 
    mutate(unc13a_detected = ifelse(inclusion_reads > 0, yes = "+", no = "-")) %>% 
    filter(sequencing_platform != "") %>% 
    left_join(cell_types %>% 
                  dplyr::select(cell,deconv,sample) %>% 
                  unique() %>% 
                  pivot_wider(id_cols = "sample",names_from = "cell",values_from = "deconv"), by = 'sample') %>%
    filter(!is.na(disease_group2)) %>% 
    unique() %>% 
    filter(disease_group2 %in% c("ALS-TDP","FTLD-TDP")) %>% 
    dplyr::select(c("unc13a_cryptic_leaf_psi",
                    "sample",  "tissue_clean",  
                    "rin",  "sex", 
                    "age",  "mutations", 
                    "sequencing_platform", 
                    "number_g_alleles", 
                    "stmn_2_cryptic_psi_leaf", "oligodendrocytes", 
                    "astrocytes", "neurons", "endothelial", "microglia")) %>% 
    column_to_rownames('sample')

fit <- lm(unc13a_cryptic_leaf_psi ~ ., data = full_cov)

fit %>% broom::tidy() %>% fwrite(paste0(here::here(),"/covariate_unc13a_ce.csv"))
full_cov %>% 
    group_by(tissue_clean) %>% 
    nest() %>% 
    mutate(lm_obj = map(data, ~lm(unc13a_cryptic_leaf_psi ~., data = .x))) %>% 
    mutate(broomed = map(lm_obj, broom::tidy)) %>% 
    dplyr::select(tissue_clean,broomed) %>% unnest() %>% filter(p.value < 0.05) %>% View()

nested %>% 
    pluck("lm_obj", 1) %>% broom::tidy()
unc13a_data_fisher = unc13a_sequencing_platform %>% 
    filter(tissue_clean != "Thoracic_Spinal_Cord") %>% 
    group_by(disease_group2,tissue_clean) %>% 
    nest() %>% 
    mutate(fisher_pvalue = map_dbl(data, quick_fisher))

unc13a_data_fisher$data = NULL

unc13a_sequencing_platform_pct = unc13a_sequencing_platform %>%
    group_by(disease_group2,tissue_clean, sequencing_platform, unc13a_detected) %>% tally() %>%
    spread(key = unc13a_detected, value = n,fill = 0) %>%
    mutate(detection = `+` / ( `-` + `+`) ) %>% 
    mutate(total =  ( `-` + `+`) ) %>% 
    mutate(plot_name = glue::glue("{sequencing_platform} \n  ({total})")) %>% 
    left_join(unc13a_data_fisher)

unc13a_sequencing_platform_pct %>%
    ggplot(aes(x = plot_name, y = detection )) + 
    geom_col(fill = "firebrick") +
    labs(title = "UNC13A CE detection and sequencing platform", 
         y = "UNC13A CE detected",x = element_blank(), subtitle = "Fisher exact test") + 
    facet_wrap(~disease_group2 + tissue_clean,scales = 'free_x') +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    geom_text(aes(x = 1.5, y = 1.05, label = paste0("p = ", signif(fisher_pvalue,digits = 2) ) )) +
    theme(legend.text = element_text(size = 15)) +
    ggpubr::theme_pubr()
# scale_y_continuous(expand = c(0,0) ) 
```

# Correlation between STMN2 CE PSI and RAP1GAP1/PFKP

```{r}
patient_others = fread(file.path(here::here(),"data","nygc_pfkp_unc13b_rap1gap.bed"))

patient_others[,sample := gsub(".SJ.out","",V4)]
clean_data_table = clean_data_table %>% 
    left_join(patient_others %>% 
                  pivot_wider(id_cols = "sample",names_from = "V7",values_from = "V5",values_fill = 0))

clean_data_table[, rap1gap_cryptic_psi := (RAP1GAP_noveldonor + RAP1GAP_novelacceptor)/ (RAP1GAP_noveldonor + RAP1GAP_novelacceptor + RAP1GAP_annotated)]
clean_data_table[, pfkp_cryptic_psi := (PFKP_novelacceptor + PFKP_noveldonor)/ (PFKP_novelacceptor + PFKP_noveldonor + PFKP_annotated)]

pfkp_plot = clean_data_table %>%
    filter(grepl("Cortex",tissue_clean)) %>%
    filter(disease_tissue == T) %>%
    filter((PFKP_novelacceptor + PFKP_noveldonor + PFKP_annotated) >=30) %>% 
    filter(junction_reads_stmn2 >= 30) %>%
    filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
    filter(pfkp_cryptic_psi > 0) %>%
    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = pfkp_cryptic_psi)) +
    geom_point() +
    stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
    geom_smooth(method = lm, se = FALSE,color = 'black') +
    xlab("STMN2 Cryptic PSI") +
    ylab("PFKP Cryptic PSI") +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 18)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1L))

rap1gap_plot = clean_data_table %>%
    filter(disease_tissue == T) %>%
    filter(grepl("Cortex",tissue_clean)) %>%
    filter((RAP1GAP_noveldonor + RAP1GAP_novelacceptor + RAP1GAP_annotated) >=30) %>% 
    filter(junction_reads_stmn2 >= 30) %>%
    filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
    filter(rap1gap_cryptic_psi > 0) %>%
    # ggplot(aes(x = rap1gap_cryptic_psi, y = unc13a_cryptic_leaf_psi)) +
    ggplot(aes(x = stmn_2_cryptic_psi_leaf, y = rap1gap_cryptic_psi)) +
    geom_point() +
    stat_cor(size = 8,method = "spearman",cor.coef.name = "rho") +
    geom_smooth(method = lm, se = FALSE,color = 'black') +
    xlab("STMN2 Cryptic PSI") +
    ylab("RAP1GAP Cryptic PSI") +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 18)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1L))  +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1L))

cowplot::plot_grid(rap1gap_plot,pfkp_plot)
```

```{r reproduce_4a_in_rap1gap}
detected_table_for_graph_rap1 = clean_data_table %>%
  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(rap1gap_cryptic_psi > 0) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
  filter((RAP1GAP_noveldonor + RAP1GAP_novelacceptor + RAP1GAP_annotated) >=30) %>% 
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,rap1gap_cryptic_psi) %>%
  mutate(log2_rap1gap = log2((rap1gap_cryptic_psi)/(stmn_2_cryptic_psi_leaf))) %>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
 as.data.table()

#get out that person who is discongruous

comparisons = combn(unique(detected_table_for_graph_rap1$detection_name),2,simplify = F)

rap1_fold_plot = detected_table_for_graph_rap1 %>%
  ggplot(aes(y = log2_rap1gap, x = detection_name,fill = detection_name)) +
  geom_boxplot(show.legend = F) +
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  ylab(expression(paste("Lo", g[2], " fold CE PSI / STMN2 CE PSI "))) +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5) + 
  ggtitle("RAP1GAP")

```

```{r reproduce_4a_in_pfkp}
detected_table_for_graph_pfkp = clean_data_table %>%
  filter(!(rs12973192 == "C/G" & rs12608932 == "C/C")) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  filter(pfkp_cryptic_psi > 0) %>%
  filter(stmn_2_cryptic_psi_leaf > 0 ) %>%
  filter(junction_reads_stmn2 >= 30) %>%
    filter((PFKP_novelacceptor + PFKP_noveldonor + PFKP_annotated) >=30) %>% 
  filter(grepl("Cortex",tissue_clean)) %>%
  filter(disease_tissue == T) %>%
  filter(disease_group2 %in% c("FTLD-TDP","ALS-TDP")) %>%
  dplyr::select(sample,tissue_clean,disease_group2,rs12973192,rs12608932,stmn_2_cryptic_psi_leaf,pfkp_cryptic_psi) %>%
  mutate(log2_pfkp = log2((pfkp_cryptic_psi)/(stmn_2_cryptic_psi_leaf))) %>%
  mutate(tissue_clean = gsub("_","\n",tissue_clean)) %>%
  group_by(rs12973192) %>%
  mutate(n_sample = length(unique(sample))) %>%
  ungroup() %>%
  mutate(detection_name = glue::glue("{rs12973192} \n ( {n_sample} )")) %>%
  mutate(no_cong = ifelse(rs12608932 == "C/C" & rs12973192 == "C/G",8,3)) %>% as.data.table()

comparisons = combn(unique(detected_table_for_graph_pfkp$detection_name),2,simplify = F)

pfkp_fold_plot = detected_table_for_graph_pfkp %>%
  ggplot(aes(y = log2_pfkp, x = detection_name,fill = detection_name)) +
  geom_boxplot(show.legend = F) +
  geom_jitter(show.legend = F, alpha = 0.8, height = 0,pch = 21) +
  stat_compare_means(aes(group = detection_name),comparisons = comparisons,label = "p.signif") +
  xlab("UNC13A rs12973192 Genotype") +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) +
  ggpubr::theme_pubr() +
  theme(text = element_text(size = 24)) +
  geom_hline(yintercept = 0,size = 1.5) + 
  ggtitle("PFKP") + 
  ylab(element_blank())

cowplot::plot_grid(rap1_fold_plot,pfkp_fold_plot)
```
# Nanopore

Read in pre-processed data (see separate script) and plot
```{r warning=FALSE, message=FALSE}
df <- read_csv(paste0(here(),
                      "/data/nanopore/combined_nanopore_dataframe.csv"))

df[is.na(df)] <- 0

p2_patient <- ggplot(df %>% filter(classification != "Neither",
                                   type=="Patient"), aes(x = name, y = 100*fraction, fill = class2)) +
    geom_bar(stat="identity", position="dodge") +
    ggpubr::theme_pubr() +
    ggeasy::easy_rotate_x_labels(side = "right") +
    ylab("Percentage of reads") +
    ggtitle("Patient RNA") +
    xlab("") +
    ggeasy::easy_add_legend_title("")

p2_shsy <- ggplot(df %>% filter(classification != "Neither",
                                type=="SHSY5Y"), aes(x = name, y = 100*fraction, fill = class2)) +
    geom_bar(stat="identity", position="dodge") +
    ggpubr::theme_pubr() +
    ggeasy::easy_rotate_x_labels(side = "right") +
    ylab("Percentage of reads") +
    ggtitle("SH-SY5Y") +
    xlab("") +
    ggeasy::easy_add_legend_title("")

p2_shsy | p2_patient

# without neither removed

p2_patient <- ggplot(df %>% filter(classification != "",
                                   type=="Patient"), aes(x = name, y = 100*fraction, fill = class2,
                                                         colour = class2)) +
    geom_bar(stat="identity", position="dodge") +
    ggpubr::theme_pubr() +
    ggeasy::easy_rotate_x_labels(side = "right") +
    ylab("Percentage of reads") +
    ggtitle("Patient RNA") +
    xlab("") +
    ggeasy::easy_add_legend_title("")

p2_shsy <- ggplot(df %>% filter(classification != "",
                                type=="SHSY5Y"), aes(x = name, y = 100*fraction, fill = class2,
                                                     colour = class2)) +
    geom_bar(stat="identity", position="dodge") +
    ggpubr::theme_pubr() +
    ggeasy::easy_rotate_x_labels(side = "right") +
    ylab("Percentage of reads") +
    ggtitle("SH-SY5Y") +
    xlab("") +
    ggeasy::easy_add_legend_title("") 
p2_shsy | p2_patient

# fractional

df2 <- df %>%
    filter(classification != "Neither") %>%
    group_by(name) %>%
    mutate(f = n/sum(n))

ggplot(df2, aes(x = name, y = 100*f, fill = classification)) +
    geom_bar(stat="identity") +
    ggpubr::theme_pubr() +
    ggeasy::easy_rotate_x_labels(side = "right") +
    ylab("Percentage of reads") +
    xlab("") +
    ggeasy::easy_add_legend_title("")

ggplot(df2 %>%
           filter(type == "Patient"), aes(x = name, y = 100*f, fill = classification)) +
    geom_bar(stat="identity") +
    ggpubr::theme_pubr() +
    ggeasy::easy_rotate_x_labels(side = "right") +
    ylab("Percentage of reads") +
    xlab("") +
    ggeasy::easy_add_legend_title("") |
    ggplot(df2 %>%
           filter(type != "Patient"), aes(x = name, y = 100*f, fill = classification)) +
    geom_bar(stat="identity") +
    ggpubr::theme_pubr() +
    ggeasy::easy_rotate_x_labels(side = "right") +
    ylab("Percentage of reads") +
    xlab("") +
    ggeasy::easy_add_legend_title("")

```